---
title: "Squirrel CMR: Squirrel Data Preparation" ## name of your project and analysis step
description: |
    The aim of this study is...
author:
    ## author information; single arguemnts can be deleted if not needed
    - name: "John Smith"
      url: https://izw-berlin.de/en/john-smith-en.html  
      affiliation: Leibniz Institute for Zoo and Wildlife Research
      affiliation_url: https://izw-berlin.de/en/
      orcid_id: 0000-0000-0000-0000
    - name: "Jane Doe"
      url: https://janedoe.com
      affiliation: Leibniz Institute for Zoo and Wildlife Research
      affiliation_url: https://izw-berlin.de/en/
    ## add more authors with the same intendation and styling if needed 
    ## a new author always starts with a hyphen!
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
        highlight: kate      ## styling of code
        code_folding: false  ## if `true` you can expand and shrink code chunks
        toc: true            ## if `true` adds a table of content
        toc_depth: 2         ## level to be displayed in the table of content
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      dev = "ragg_png", fig.width = 9, fig.height = 6, dpi = 600, retina = 1)
Sys.setlocale("LC_TIME", "C")
```

* **Research question:**
* **Study area:**
* **Data:** 


# Setup

```{r packages}
## for non-CRAN packages please keep install instruction
## but commented so it is not run each time, e.g.
# devtools::install_github("EcoDynIZW/template")

## libraries used in this script
## please add ALL LIBRARIES NEEDED HERE
## please remove libraries from the list that are not needed anymore 
## at a later stage
d6::simple_load(c("tidyverse", 
                  #"gridExtra", 
                  "here", 
                  "lubridate"))
```


# Data

```{r data}
# Individual data of captured squirrels
squirrels_raw <- vroom::vroom(file = here::here("data-raw/cmr-raw/squirrels_raw_data_20251126.csv"), 
                             delim = ";", 
                             locale = locale(encoding = "ISO-8859-1", 
                                             decimal_mark = ".")
)

# Trapping effort data
trapping_effort_trep <- vroom::vroom(file = here::here("data-raw/cmr-raw/Trapping_Effort_20250713_SD_Treptower_Park.csv"),
                                delim = ";")
trapping_effort_lilie <- vroom::vroom(file = here::here("data-raw/cmr-raw/Trapping_Effort_20250713_SD_Lilienthal.csv"),
                                delim = ";")
trapping_effort_rahn <- vroom::vroom(file = here::here("data-raw/cmr-raw/Trapping_Effort_20250713_SD_Rahnsdorf.csv"),
                                delim = ";")

# Human activity
hum_act <- vroom::vroom(file = here::here("data-raw/cmr-raw/human_activity_all_20251126.csv"),
                                delim = ";")
```

# CMR Data
## Data manipulation
Looking for strange values, typos etc. 
```{r}
table(squirrels_raw$Project_phase, useNA = "always") # FINE
table(squirrels_raw$Season, useNA = "always") # FINE
table(squirrels_raw$Study_site, useNA = "always") # FINE
table(squirrels_raw$Trap_nr, useNA = "always") # FINE
table(squirrels_raw$Squirrel_ID, useNA = "always") # Seems fine at a first glance, 9 times no ID - Fine
table(squirrels_raw$Capture_or_recapture, useNA = "always") # 6 times not indicated if capture or recapture - is from 6 of the individuals without ID (other individuals without ID were captures for sure) - FINE
table(squirrels_raw$Sex, useNA = "always") # 9 times no sex - 6x individuals without ID, 3 x captures where we did not determine the sex - FINE
table(squirrels_raw$age_class, useNA = "always") # 4 times NA - individuals without ID, for the others with ID we could determine the age class - FINE
table(squirrels_raw$docility_score, useNA = "always") # FINE
```



### New column "study_site_specific"

Here I divide the study site Treptower Park into two parts: Treptower Park north
and Treptower Park south. We want to test this differentiation, as possible differences between the areas of the park were identified in a master's thesis

Trap locations in...
Trep_north: B_Trep_11, B_Trep_16, B_Trep_05, B_Trep_02_mobil, B_Trep_07_mobil, 
            B_Trep_09, B_Trep_07, B_Trep_19, B_Trep_04, B_Trep_18
            
Trep_south: B_Trep_13, B_Trep_03, B_Trep_14, B_Trep_01, B_Trep_02, B_Trep_06, 
            B_Trep_15, B_Trep_17, B_Trep_08
  

```{r}
## adding new column to table
## first defining the list of the specific study sites 

treptower_park_south <- c("B_Trep_13", "B_Trep_03", "B_Trep_14", "B_Trep_01",
                          "B_Trep_02", "B_Trep_06", "B_Trep_15", "B_Trep_17", 
                          "B_Trep_08")

treptower_park_north <- c("B_Trep_11", "B_Trep_16", "B_Trep_05", "B_Trep_02_mobil", "B_Trep_07_mobil", "B_Trep_09", "B_Trep_07", "B_Trep_19", "B_Trep_04", "B_Trep_18","B_Trep_10")


## using case_when to create the 'study_site_specific' column
squirrels_raw <- squirrels_raw %>%
  dplyr::mutate(study_site_specific = dplyr::case_when(
    Trap_nr %in% treptower_park_south ~ "Treptow_south",
    Trap_nr %in% treptower_park_north ~ "Treptow_north",
    Study_site == "Lilienthal" ~ "Lilienthal",
    Study_site == "Rahnsdorf" ~ "Rahnsdorf",
    TRUE ~ NA_character_  # For all other cases, set to NA
  ))
```

Check values
```{r}
table(squirrels_raw$Study_site, useNA = "always")
table(squirrels_raw$study_site_specific, useNA = "always") # values fit together
140 + 116
```

### New column "area_type"
This column differs between "urban" and "rural" area types. This is needed for following analyses with the marked package. Treptower Park and Lilienthal count as urban sites, Rahnsdorf as rural. 

```{r}
## create new column 
squirrels_raw <- squirrels_raw %>%
  mutate(area_type = case_when(
    Study_site %in% c("Treptower_Park", "Lilienthal") ~ "urban",
    Study_site == "Rahnsdorf" ~ "rural",
    TRUE ~ "unknown" # for any other values
  ))
```

Check values
```{r}
table(squirrels_raw$Study_site, useNA = "always")
205+256
table(squirrels_raw$area_type, useNA = "always") # rural 88 urban 461
```


### New columns "year" and "season_year"
```{r}
## creating a column for the year that the squirrel was captured in 
squirrels_raw <- squirrels_raw %>%
  mutate(Year = year(dmy(Date)))  # dmy() converts "DD.MM.YYYY" to a date format

## create the new column 'season_year'
squirrels_raw <- squirrels_raw %>%
  mutate(season_year = paste0(Season, "_", Year))
```

Check values
```{r}
table(squirrels_raw$Year, useNA = "always")
table(squirrels_raw$season_year, useNA = "always")

table(squirrels_raw$season_year, squirrels_raw$Study_site, useNA = "always")
```


## Cleanup of raw data
```{r}
names(squirrels_raw)
```

Change column type to factor or numeric
```{r}
squirrels_raw <- 
  squirrels_raw %>%
  mutate(across(.cols = c(Season, Project_phase, Study_site, Trap_nr, Squirrel_ID, Name,
                          Capture_or_recapture, Sex, repro_status, age_class, physical_status,
                          nutritional_status, blood_smear, full_blood, plasma, faeces_sample,
                          after_sample, EHI_feaces, EHI_after, fur_sample, hind_foot_sample,
                          oral_sample, tissue_sample, EHI_tissue, score_sheet_cat, Weather,
                          study_site_specific, area_type, season_year),
                .fns =  as.factor)) %>%
  mutate(Date = as.Date(Date, format = "%d.%m.%Y")) %>%
  droplevels()

squirrels_raw$Study_site <- fct_relevel(squirrels_raw$Study_site, "Rahnsdorf") #%>% 
squirrels_raw$study_site_specific <- fct_relevel(squirrels_raw$study_site_specific, "Rahnsdorf")
```

Create a new data frame that only includes columns of importance for analysis and remove all rows without an ID
```{r}
squirrels <- squirrels_raw %>% 
  dplyr::select(Date, Year, Season, season_year, Project_phase, 
                Study_site, study_site_specific, area_type, Trap_nr, 
                Squirrel_ID, Name, Capture_or_recapture, Sex, age_class, 
                body_length_mm, tail_length_mm, hind_foot_length_mm, 
                Weight_squirrel, physical_status, nutritional_status, 
                tissue_sample, Temperature_Celsius, Weather, 
                trap_score, docility_score, release_score, protocol_nr, protocol_nr_old) %>%
  filter(!is.na(Squirrel_ID))
# protocol_nr_old is important for the genetic data
```

# Trapping Effort
## Data manipulation
```{r}
## prepare the data frames for further adjustments
## NAs need to be replaced with 0 and characters (geklaut = stolen), too 
# Treptow
trapping_effort_trep <- trapping_effort_trep %>%
  filter(if_any(everything(), ~ !is.na(.))) %>% # remove empty rows
  mutate(across(where(is.character), ~recode(.x, "GEKLAUT?" = "0", .default = .x))) %>%
  mutate(across(where(is.character), ~recode(.x, "09:48" = "0.5", .default = .x))) %>% # add a value of 0.5 because we forgot to activate it, but we did it ~ 2 hours before we stopped
  mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>%
  mutate(across(7:26, as.numeric)) 

# Lilie
trapping_effort_lilie <- trapping_effort_lilie %>%
  filter(if_any(everything(), ~ !is.na(.))) %>%
  mutate(across(where(is.character), ~recode(.x, "stolen" = "0", .default = .x))) %>%
  mutate(across(where(is.character), ~recode(.x, "forgot to activate" = "0", .default = .x))) %>%
  mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>%
  mutate(across(7:17, as.numeric))

# Rahnsdorf
trapping_effort_rahn <- trapping_effort_rahn %>%
  filter(if_any(everything(), ~ !is.na(.))) %>%
  mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>%
  mutate(across(7:20, as.numeric))
```

Change the format of the table: The columns per trap should become single rows 
```{r}
# Treptow
effort_traps_trep <- trapping_effort_trep %>%
  select(7:26, Project_phase, Study_site, Trapping_Date, Trapping_Duration) %>%  
  pivot_longer(cols = c(-Project_phase, -Study_site, -Trapping_Date, -Trapping_Duration), 
               names_to = "Trap_nr", 
               values_to = "opened") %>%
  mutate(hours_day = round(x = as.numeric(x = Trapping_Duration * opened, 
                                          units = "hours"), 
                           digits = 2))

# Lilie
effort_traps_lilie <- trapping_effort_lilie %>%
  select(7:17, Project_phase, Study_site, Trapping_Date, Trapping_Duration) %>%  
  pivot_longer(cols = c(-Project_phase, -Study_site, -Trapping_Date, -Trapping_Duration), 
               names_to = "Trap_nr", 
               values_to = "opened") %>%
  mutate(hours_day = round(x = as.numeric(x = Trapping_Duration * opened, 
                                          units = "hours"), 
                           digits = 2))

# Rahnsdorf
effort_traps_rahn <- trapping_effort_rahn %>%
  select(7:20, Project_phase, Study_site, Trapping_Date, Trapping_Duration) %>%  
  pivot_longer(cols = c(-Project_phase, -Study_site, -Trapping_Date, -Trapping_Duration), 
               names_to = "Trap_nr", 
               values_to = "opened") %>%
  mutate(hours_day = round(x = as.numeric(x = Trapping_Duration * opened, 
                                          units = "hours"), 
                           digits = 2))
```


Combine the different tables into one
```{r}
trapping_effort <- bind_rows(effort_traps_trep, effort_traps_lilie, effort_traps_rahn)
363 + 518 + 840
# change column types
trapping_effort <- trapping_effort %>%
  mutate(across(.cols = c(Project_phase, Study_site, Trap_nr),
                .fns =  as.factor)) %>%
  mutate(Date = as.Date(Trapping_Date, format = "%d.%m.%Y")) %>%
  mutate(Trapping_Duration = hms::as_hms(as.numeric(Trapping_Duration * opened))) # remove the hours for traps that were not activated
```

## Summaries
Absolute trapping_duration per day per study site for all traps that were open
```{r}
effort_site_phase_day <- trapping_effort %>%
  group_by(Project_phase, Study_site, Trapping_Date) %>%
  summarise(hours_day_site = sum(hours_day))
```

Absolute trapping_duration per project phase per study site for all traps that were open
```{r}
effort_site_phase <- trapping_effort %>%
  group_by(Project_phase, Study_site) %>%
  summarise(hours_day_site = sum(hours_day))
```

# Human activity
## Data manipulation
### New column "study_site_specific"
Explanation see above

```{r}
hum_act <- hum_act %>%
  dplyr::mutate(study_site_specific = dplyr::case_when(
    Trap_nr %in% treptower_park_south ~ "Treptow_south",
    Trap_nr %in% treptower_park_north ~ "Treptow_north",
    Study_site == "Lilienthal" ~ "Lilienthal",
    Study_site == "Rahnsdorf" ~ "Rahnsdorf",
    TRUE ~ NA_character_  # For all other cases, set to NA
  ))
```

Check values
```{r}
table(hum_act$Study_site, useNA = "always")
table(hum_act$study_site_specific, useNA = "always") # values fit together
109 + 109
```

### New column "area_type"
This column differs between "urban" and "rural" area types. This is needed for following analyses with the marked package. Treptower Park and Lilienthal count as urban sites, Rahnsdorf as rural. 

```{r}
## create new column 
hum_act <- hum_act %>%
  mutate(area_type = case_when(
    Study_site %in% c("Treptower_Park", "Lilienthal") ~ "urban",
    Study_site == "Rahnsdorf" ~ "rural",
    TRUE ~ "unknown" # for any other values
  ))
```

Check values
```{r}
table(hum_act$Study_site, useNA = "always")
13+218
table(hum_act$area_type, useNA = "always") # rural 88 urban 461
```


### New column "year"

```{r}
## creating a column for the year that the squirrel was captured in 
hum_act <- hum_act %>%
  mutate(Year = year(dmy(Date)))  # dmy() converts "DD.MM.YYYY" to a date format
```

Check values
```{r}
table(hum_act$Year, useNA = "always")
```
### New column "timeblock"
There are too many different starting times (start_obs).
We will standardise them and create 30 min time blocks 
```{r}
hum_act <- hum_act %>%
  mutate(
    timeblock = case_when(
      # Start_obs %in% c("07:00:00", "07:05:00") ~ "07:00",
      # Start_obs %in% c("07:30:00", "07:31:00", "07:32:00", "07:33:00", "07:34:00", 
      #                  "07:35:00") ~ "07:30",
      Start_obs %in% c("07:50:00", "07:55:00", "08:00:00", "08:01:00", "08:02:00", 
                       "08:03:00", "08:04:00", "08:05:00", "08:10:00", "08:15:00") ~ "08:00",
      Start_obs %in% c("08:18:00", "08:20:00", "08:24:00", "08:25:00", "08:30:00", "08:31:00", "08:32:00", 
                       "08:33:00", "08:34:00", "08:35:00","08:40:00", "08:45:00" ) ~ "08:30",
      Start_obs %in% c("08:55:00", "09:00:00", "09:03:00", "09:04:00", "09:05:00", "09:06:00", 
                       "09:10:00", "09:15:00") ~ "09:00",
      Start_obs %in% c( "09:16:00", "09:30:00", "09:33:00", "09:34:00", "09:35:00", "09:40:00", 
                        "09:45:00") ~ "09:30",
      Start_obs %in% c("09:55:00", "09:57:00", "10:00:00", "10:01:00", "10:03:00", "10:04:00", 
                       "10:05:00", "10:10:00", "10:15:00") ~ "10:00",
      Start_obs %in% c("10:20:00","10:25:00", "10:30:00", "10:31:00", "10:33:00", "10:35:00", 
                       "10:37:00", "10:40:00") ~ "10:30",
      Start_obs %in% c("11:00:00","11:03:00", "11:05:00", "11:07:00") ~ "11:00",
      Start_obs %in% c("11:30:00") ~ "11:30"
    )
  )

unique(hum_act$timeblock)

# hourly time block
hum_act <- hum_act %>%
  mutate(
    hourblock = case_when(
      #timeblock %in% c("07:00", "07:30") ~ "07:00",
      timeblock %in% c("08:00", "08:30") ~ "08:00",
      timeblock %in% c("09:00", "09:30") ~ "09:00",
      timeblock %in% c("10:00", "10:30") ~ "10:00",
      timeblock %in% c("11:00", "11:30") ~ "11:00"
    )
  )

unique(hum_act$hourblock)
```

Quick exploration
```{r}
table(hum_act$Study_site, hum_act$timeblock)
table(hum_act$Trap_nr, hum_act$timeblock)
table(hum_act$Study_site, hum_act$hourblock)
table(hum_act$Trap_nr, hum_act$hourblock)
```
## Count days per study site
```{r}
days <- hum_act %>% 
  group_by(Study_site, Date) %>% 
  summarise(n_obs_day = n())

days

days2 <- days %>% 
  group_by(Study_site) %>%
  summarise(n_days = n())

days2
table(days$Study_site)
```

## Calculate direct human activity 

total human activity -> adding up the number of people, dogs, and extraordinary 
events (e.g. cars).
```{r}
hum_act <- hum_act %>%
  mutate(
    huac = People + Dogs + Bicycles + Ex_events)
```


## Calculating mean human activity
We need values for the human acitivty
In 2024 and 2025 students measured the human activity at 2 specific spots per study site, in total 6 spots.
For the following analysis we calculate the mean value of the total human activity per study site or the specific study sites where Treptower park is divided in two parts

Means (not used, because not accounting for the number of observations)
```{r}
huac_means <- hum_act %>%
  group_by(Study_site) %>%
  summarise(sum_huac = sum(huac)) %>%
  mutate(mean_30 = sum_huac/8,
         mean_60 = sum_huac/4)
```

Means per timeblock (not used)
```{r}
huac_means_site_block <- hum_act %>%
  group_by(Study_site, timeblock, hourblock) %>% 
  summarise(mean_huac = mean(huac),
            mean_people = mean(People),
            mean_dogs = mean(Dogs),
            mean_bikes = mean(Bicycles),
            mean_exevents = mean(Ex_events),
            n_obs = n(),
            .groups = "drop")


huac_means_specific_block <- hum_act %>%
  group_by(study_site_specific, timeblock, hourblock) %>% 
  summarise(mean_huac_spec = mean(huac),
            mean_people = mean(People),
            mean_dogs = mean(Dogs),
            mean_bikes = mean(Bicycles),
            mean_exevents = mean(Ex_events),
            n_obs = n(),
            .groups = "drop")
```

Means per study site = total sum divided by interval (used)
```{r}
huac_means <- hum_act %>%
  group_by(Study_site) %>% 
  summarise(sum_all_huac = sum(huac),
            n_obs = n(),
            sum_huac = sum_all_huac/n_obs, 
            .groups = "drop") %>%
  mutate(mean_30 = sum_huac/8,
         mean_60 = sum_huac/4)

huac_means_specific <- hum_act %>%
  group_by(Study_site, study_site_specific) %>% 
  summarise(sum_all_huac = sum(huac),
            n_obs = n(),
            sum_huac = sum_all_huac/n_obs, 
            .groups = "drop") %>%
  mutate(spec_mean_30 = sum_huac/8,
         spec_mean_60 = sum_huac/4) %>%
  left_join(y = huac_means[, c(1,5,6)], by = "Study_site") # add the values per study site
```



# Save data
```{r}
# Squirrel CMR
## raw data
vroom::vroom_write(x = squirrels_raw, file = here::here("output", "data-proc", "cmr-proc", "squirrels_raw_20251127.csv"))
saveRDS(object = squirrels_raw, file = here::here("output", "data-proc", "cmr-proc", "squirrels_raw_20251127.RDS"))

## data for analysis
vroom::vroom_write(x = squirrels, file = here::here("output", "data-proc", "cmr-proc", "squirrels_20251127.csv"))
saveRDS(object = squirrels, file = here::here("output", "data-proc", "cmr-proc", "squirrels_20251127.RDS"))

# Trapping effort
## total
vroom::vroom_write(x = trapping_effort, file = here::here("output", "data-proc", "cmr-proc", "trapping_effort_20251122.csv"))
saveRDS(object = trapping_effort, file = here::here("output", "data-proc", "cmr-proc", "trapping_effort_20251122.RDS"))

## per study site, per project phase, per day
saveRDS(object = effort_site_phase_day, file = here::here("output", "data-proc", "cmr-proc", "trapping_effort_sitephaseday_20251122.RDS"))

## per study site, per project phase
saveRDS(object = effort_site_phase, file = here::here("output", "data-proc", "cmr-proc", "trapping_effort_sitephase_20251122.RDS"))

# Human activity
## total, detailed table
vroom::vroom_write(x = hum_act, file = here::here("output", "data-proc", "cmr-proc", "human_activity_20251127.csv"))
saveRDS(object = hum_act, file = here::here("output", "data-proc", "cmr-proc", "human_activity_20251127.RDS"))

## means
saveRDS(object = huac_means_specific, 
        file = here::here("output", "data-proc", "cmr-proc", "human_activity_means_20251127.RDS"))
```


***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>
