---
title: "Squirrel CMR: Exploration individual data" ## name of your project and analysis step
description: |
    The aim of this study is...
author:
    ## author information; single arguemnts can be deleted if not needed
    - name: "John Smith"
      url: https://izw-berlin.de/en/john-smith-en.html  
      affiliation: Leibniz Institute for Zoo and Wildlife Research
      affiliation_url: https://izw-berlin.de/en/
      orcid_id: 0000-0000-0000-0000
    - name: "Jane Doe"
      url: https://janedoe.com
      affiliation: Leibniz Institute for Zoo and Wildlife Research
      affiliation_url: https://izw-berlin.de/en/
    ## add more authors with the same intendation and styling if needed 
    ## a new author always starts with a hyphen!
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
        highlight: kate      ## styling of code
        code_folding: false  ## if `true` you can expand and shrink code chunks
        toc: true            ## if `true` adds a table of content
        toc_depth: 2         ## level to be displayed in the table of content
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      dev = "ragg_png", fig.width = 9, fig.height = 6, dpi = 600, retina = 1)
Sys.setlocale("LC_TIME", "C")
```

-   **Research question:**
-   **Study area:**
-   **Data:**

# Setup

```{r packages}
## for non-CRAN packages please keep install instruction
## but commented so it is not run each time, e.g.
# devtools::install_github("EcoDynIZW/template")

## libraries used in this script
## please add ALL LIBRARIES NEEDED HERE
## please remove libraries from the list that are not needed anymore 
## at a later stage
library("ggplot2")
library("lubridate")
library("tidyverse")
```

## Plot setup

Set theme for ggplots

```{r theme}
theme_set(
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 2),
    plot.margin = margin(6, 10, 4, 4),
    axis.ticks = element_line(colour = "black", linewidth = 1.5),
    axis.ticks.length = unit(0.2, "cm"),
    title = element_text(size = 20), # axis family = "Zilla Slab", family = "Roboto", 
    axis.text = element_text(size = rel(1.4)))) # axis ticks, family = "Roboto", 
```

Labeller for study sites

```{r}
# Remove the underscore in the name of the study site without changing the factor levels
Study_site_names <- as_labeller(c("Rahnsdorf" = "woodland area", "Treptower_Park" = "urban park", "Lilienthal" = "cemetery"))

Study_site_names_specific <- as_labeller(c("Rahnsdorf" = "Rahnsdorf", "Lilienthal" = "Lilienthal", "Treptow_north" = "Treptower Park north", "Treptow_south" = "Treptower Park south"))
```

```{r}
site_cols <- c(
  "Rahnsdorf" = "#1b9e77",
  "Treptower_Park" = "#d95f02",
  "Lilienthal" = "#7570b3"
)
```

# Data

```{r data}
#squirrels_raw <- readRDS(file = here::here("output", "data-proc", "cmr-proc", "squirrels_raw_20250819.RDS"))

squirrels <- readRDS(file = here::here("output", "data-proc", "cmr-proc", "squirrels_20251127.RDS"))
squirrels <- squirrels %>% 
  mutate(
    Date = as.Date(Date, format = "%d.%m.%Y"),
    #Sex = ifelse(Sex %in% c("", NA), "unknown", Sex),
    #Sex = factor(Sex, levels = c("f", "m", "unknown")), 
    Project_phase = as.factor(Project_phase)
  )

trapping_effort <- readRDS(file = here::here("output", "data-proc", "cmr-proc", "trapping_effort_20251122.RDS"))
trapping_effort <- trapping_effort %>%
  mutate(
    Duration_hours = as.numeric(Trapping_Duration, units = "hours")
  )

hum_act_means <- readRDS(file = here::here("output", "data-proc", "cmr-proc", "human_activity_means_20251127.RDS"))
```

Basic information

```{r}
str(squirrels)
```

# Capturing events

## CE per ID

```{r}
captures_per_individual <- squirrels %>%
  count(Squirrel_ID, Study_site)

ggplot(captures_per_individual, aes(n)) +
  geom_histogram(binwidth = 1, boundary = 0) +
  labs(x = "Number of captures per individual",
       y = "Count of individuals",
       title = "Capture frequency distribution")

ggplot(captures_per_individual, aes(n, fill = Study_site)) +
  geom_histogram(position = "dodge") +
  scale_fill_manual(values = c("Treptower_Park" = "tomato3", 
                               "Lilienthal" = "orange2", 
                               "Rahnsdorf" = "palegreen4")) +
  labs(x = "Study site", y = "captures per individual",
       title = "Capture frequency per site") +
  theme(legend.position = "right")

# ggsave(filename = here::here("plots", "cmr", "Captures_ID_site.png"),
#        height = 4, width = 6, units = "in",
#        dpi=800)
```

```{r}
mean(captures_per_individual$n)
sd(captures_per_individual$n)
```

The most captured individuals are 1: 8041A5D Treptower_Park 19 - Scar 2: 806170D Treptower_Park 17 - Valentine 3: 804C8B8 Lilienthal 16 - Alio 4: 80618A8 Lilienthal 16 - Gandalf 5: 8041A02 Rahnsdorf 15 - Ratchet

## CE study sites

```{r}
captures_site <- squirrels %>%
  count(Study_site)
captures_site
```

```{r}
ggplot(captures_site, aes(Study_site, n, fill = Study_site)) +
  geom_col() +
  scale_fill_manual(values = c("Treptower_Park" = "tomato3", 
                               "Lilienthal" = "orange2", 
                               "Rahnsdorf" = "palegreen4")) +
  scale_x_discrete(labels = Study_site_names) +
  labs(x = "Study site", y = "Number of capturing events",
       title = "") # Capturing events per site

# ggsave(filename = here::here("plots", "cmr", "CE_study_sites.png"),
#        height = 4, width = 5, units = "in",
#        dpi=800)
```

### CE study sites + sex

```{r}
captures_site_sex <- squirrels %>%
  group_by(Study_site, Sex) %>%
  count()

ggplot(captures_site_sex, aes(Study_site, n, fill = Sex)) +
  geom_col() +
  # scale_fill_manual(values = c("Treptower_Park" = "tomato3", 
  #                              "Lilienthal" = "orange2", 
  #                              "Rahnsdorf" = "palegreen4")) +
  labs(x = "Study site", y = "Number of captures",
       title = "Capture frequency per site")
```

## CE per project phase

```{r}
captures_phase <- squirrels %>%
  count(Project_phase)

ggplot(captures_phase, aes(Project_phase, n)) +
  geom_point() +
  scale_y_continuous(
    limits = c(0, 110),
    breaks = seq(0, 110, by = 25),
    expand = c(.004, .001)) +
  labs(x = "Project phase", y = "Number of capturing events",
       title = "Capturing events per project phase")
```

number phases per ID

```{r}
phases_ID <- squirrels %>%
  group_by(Squirrel_ID, Project_phase) %>%
  count()

phases_ID

phases_ID2 <- phases_ID %>%
  group_by(Squirrel_ID) %>%
  count()

phases_ID2
```

```{r}
mean(phases_ID2$n)
sd(phases_ID2$n)
summary(phases_ID2$n)
```

## CE per season year

```{r}
captures_season_year <- squirrels %>%
  count(Project_phase, season_year, sort = F)
captures_season_year$season_year <- fct_reorder(.f = captures_season_year$season_year, as.numeric(captures_season_year$Project_phase))

ggplot(captures_season_year, aes(season_year, n)) +
  geom_point() +
  scale_y_continuous(
    limits = c(0, 110),
    breaks = seq(0, 110, by = 25),
    expand = c(.004, .001)) +
  labs(x = "season_year", y = "Number of capturing events",
       title = "Capturing events per year and season")
```

## CE and temperature

```{r}
ggplot(squirrels, aes(x = Temperature_Celsius)) +
  geom_histogram(binwidth = 1, color = "black", fill = "grey70") +
  scale_x_continuous(
    limits = c(-1, 30),
    breaks = seq(0, 30, by = 10),
    expand = c(.004, .001)) +
  scale_y_continuous(
    limits = c(0, 50),
    breaks = seq(0, 50, by = 25),
    expand = c(.004, .001)) +
  labs(x = "Temperature (°C)", y = "Number of capturing events",
       title = "Capture frequency across temperatures")

ggplot(squirrels, aes(x = Temperature_Celsius, fill = Study_site)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(
    limits = c(-1, 30),
    breaks = seq(0, 30, by = 10),
    expand = c(.004, .001)) +
  scale_y_continuous(
    limits = c(0, 30),
    breaks = seq(0, 30, by = 10),
    expand = c(.004, .001)) +
  facet_wrap(~ Study_site) +
  labs(x = "Temperature (°C)", y = "Captures")
```

```{r}
daily <- squirrels %>%
  group_by(Date) %>%
  summarise(
    mean_temp = mean(Temperature_Celsius, na.rm = TRUE),
    captures = n()
  )

ggplot(daily, aes(x = mean_temp, y = captures)) +
  geom_point() +
  geom_smooth(method = "loess") +
  scale_x_continuous(
    limits = c(0, 25),
    breaks = seq(0, 25, by = 5),
    expand = c(.004, .001)) +
  scale_y_continuous(
    limits = c(0, 15),
    breaks = seq(0, 15, by = 5),
    expand = c(.004, .001)) +
  labs(x = "Mean daily temperature (°C)",
       y = "Number of captures",
       title = "Daily capture frequency vs. temperature")
```

# Capture recapture

## CR per project phase

```{r}
squirrels %>%
  ggplot(aes(x = Project_phase, fill = Capture_or_recapture)) +
  geom_bar(stat = "count", position = "dodge") +
  scale_fill_manual(values = c("capture" = "grey28", "recapture" = "grey")) +
  scale_y_continuous(
    limits = c(0, 100),
    breaks = seq(0, 100, by = 50),
    expand = c(.004, .001)) +
  labs(title = "Captures and recaptures at all locations by project phase",
       x = "Project phase",
       y = "Number of individuals",
       fill = "Capture or recapture")
```

## CR per study site

```{r}
squirrels %>%
  ggplot(aes(x = Study_site, fill = Study_site)) +
  geom_bar(stat = "count", position = "stack", color = "white") +
   scale_y_continuous(
    limits = c(0, 200),
    breaks = seq(0, 200, by = 50),
    expand = c(.004, .001)) +
  scale_fill_manual(values = c("Treptower_Park" = "tomato3", 
                               "Lilienthal" = "orange2", 
                               "Rahnsdorf" = "palegreen4")) +
  labs(title = "",
       x = "Study sites",
       y = "Number of individuals",
       fill = "Study site") +
  facet_wrap(~Capture_or_recapture) 

# ggsave(filename = here::here("plots", "cmr", "CR_study_sites.png"), 
#        height = 4, width = 5, units = "in",
#        dpi=800)
```

## CR per study site + sex

```{r}
captures <- squirrels %>%
  filter(Capture_or_recapture == "capture")
```

```{r}
table(captures$Sex, useNA = "always")
table(captures$Study_site, useNA = "always")
table(captures$Sex, captures$Study_site, useNA = "always")
```

```{r}
recaptures <- squirrels %>%
  filter(Capture_or_recapture == "recapture") %>%
  group_by(Squirrel_ID, Study_site) %>%
  count()

recaptures
```

```{r}
table(recaptures$Study_site, useNA = "always")

# recapture rates
93/141 # 0.6595745
43/67 # 0.641791
35/53 # 0.6603774
15/21 # 0.7142857
```

## CR per study site and project phase

```{r}
## differentiation between study sites
squirrels %>%
  ggplot(aes(x = Project_phase, fill = Study_site)) +
  geom_bar(stat = "count", position = "stack", color = "white") +
  scale_y_continuous(
    limits = c(0, 50),
    breaks = seq(0, 50, by = 25),
    expand = c(.004, .001)) +
  scale_fill_manual(values = c("Treptower_Park" = "tomato3", 
                               "Lilienthal" = "orange2", 
                               "Rahnsdorf" = "palegreen4")) +
  labs(title = "Captures and recaptures per study site in each project phase",
       x = "Project phase",
       y = "Number of individuals",
       fill = "Study site") +
  facet_wrap(~Study_site, labeller = labeller(facet = Study_site_names))   # Create facets for each study site
```

## Recapture rate year

```{r}
recapture_rate_year <- squirrels %>%
  count(Year, Capture_or_recapture) %>%
  tidyr::pivot_wider(names_from = Capture_or_recapture,
                     values_from = n,
                     values_fill = 0) %>%
  mutate(recapture_rate = recapture / (capture + recapture))

ggplot(recapture_rate_year, aes(x = Year, y = recapture_rate)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0,1)) +
  labs(y = "Recapture rate", title = "Observed Recapture Rate per Year")
```

## Recapture rate study sites proj phase

```{r}
recapture_rate_site <- squirrels %>%
  count(Project_phase, Study_site, Capture_or_recapture) %>%
  tidyr::pivot_wider(names_from = Capture_or_recapture,
                     values_from = n,
                     values_fill = 0) %>%
  mutate(recapture_rate = recapture / (capture + recapture))

ggplot(recapture_rate_site, aes(Project_phase, recapture_rate)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0,1)) +
  labs(y = "Recapture rate",
       title = "Observed Recapture Rate per Study Site") + 
  facet_wrap(~Study_site, nrow = 3)
```

## Capture history rasterplot

```{r}
capture_history <- squirrels %>%
  arrange(Squirrel_ID, Project_phase, Study_site) %>%
  mutate(capture = 1) %>%
  select(Squirrel_ID, Study_site, Project_phase, capture)

ggplot(capture_history, aes(Project_phase, Squirrel_ID)) +
  geom_tile(aes(fill = capture)) +
  scale_fill_gradient(low = "white", high = "black", guide = "none") +
  labs(x = "Project_phase", y = "Individual ID",
       title = "Capture history (presence/absence)")

ggplot(capture_history, aes(Project_phase, Squirrel_ID)) +
  geom_tile(aes(fill = capture)) +
  scale_fill_gradient(low = "white", high = "black", guide = "none") +
  labs(x = "Project_phase", y = "Individual ID",
       title = "Capture history (presence/absence)") +
  facet_wrap(~ Study_site, scales = "free_y", nrow = 3) 
```

## Captures per study site

```{r}
ggplot(data = captures, mapping = aes(x = Study_site, fill = Study_site)) +
  geom_bar(stat = "count", position = "stack", color = "white") +
   scale_y_continuous(
    limits = c(0, 80),
    breaks = seq(0, 80, by = 40),
    expand = c(.004, .001)) +
  scale_fill_manual(values = c("Treptower_Park" = "tomato3", 
                               "Lilienthal" = "orange2", 
                               "Rahnsdorf" = "palegreen4")) +
  scale_x_discrete(labels = Study_site_names) +
  labs(title = "",
       x = "Study sites",
       y = "Number of individuals",
       fill = "Study site") 

# ggsave(filename = here::here("plots", "cmr", "Caps_study_sites.png"),
#        height = 4, width = 5, units = "in",
#        dpi=800)
```


```{r}
ggplot(data = captures, mapping = aes(x = Study_site, fill = Study_site)) +
  geom_bar(stat = "count", position = "stack", color = "white") +
   scale_y_continuous(
    limits = c(0, 40),
    breaks = seq(0, 40, by = 20),
    expand = c(.004, .001)) +
  scale_fill_manual(values = c("Treptower_Park" = "tomato3", 
                               "Lilienthal" = "orange2", 
                               "Rahnsdorf" = "palegreen4")) +
  scale_x_discrete(labels = Study_site_names) +
  labs(title = "",
       x = "Study sites",
       y = "Number of individuals",
       fill = "Study site") + 
  facet_wrap(~Sex, nrow = 3)

# ggsave(filename = here::here("plots", "cmr", "Caps_study_sites_sex.png"),
#        height = 4, width = 5, units = "in",
#        dpi=800)
```

# Sex

## Sex ratio study sites

```{r}
sex_site <- squirrels %>%
  count(Study_site, Sex)

ggplot(sex_site, aes(Study_site, n, fill = Sex)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = Study_site_names) +
  labs(y = "Proportion",
       x = "Study site",
       title = "") + #Sex ratio per study site
  theme(legend.position = "right")

# ggsave(filename = here::here("plots", "cmr", "Sex_ratio_site.png"),
#        height = 4, width = 6, units = "in",
#        dpi=800)
```

## Sex ratio study site, phase

```{r}
sex_phase_site <- squirrels %>%
  count(Project_phase, Study_site, Sex)

ggplot(sex_phase_site, aes(Project_phase, n, fill = Sex)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(y = "Proportion",
       title = "Sex ratio per project phase per study site") + 
  facet_wrap(~Study_site, nrow = 3) + 
  theme(legend.position = "top")
```

# Body condition

## Weight

```{r}
ggplot(squirrels, aes(Study_site, Weight_squirrel)) +
  geom_boxplot() +
  labs(x = "Study site", y = "Weight (g)",
       title = "Body weight distribution per study site")

ggplot(squirrels, aes(Project_phase, Weight_squirrel)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Project phase", y = "Weight (g)",
       title = "Body weight per project phase")

ggplot(squirrels, aes(Sex, Weight_squirrel)) +
  geom_boxplot() +
  labs(x = "Sex", y = "Weight (g)",
       title = "Body weight distribution per sex and study site") +
  facet_wrap(~Study_site, nrow = 3)
```

### Weight change of squirrels per project phase

Calculate the mean weight per project phase

```{r}
weight_phase <- squirrels %>%
  group_by(Study_site, Project_phase, Squirrel_ID) %>%
  summarize(
    mean_weight = mean(Weight_squirrel, na.rm = TRUE),
    sd_weight = sd(Weight_squirrel, na.rm = TRUE),
    n = n()
  ) 

weight_year <- squirrels %>%
  group_by(Squirrel_ID, Year) %>%
  summarise(mean_weight = mean(Weight_squirrel, na.rm = TRUE)) %>%
  ungroup()

diff_weight <- weight_phase %>%
  group_by(Squirrel_ID) %>%
  mutate(diff_weight = mean_weight - lag(mean_weight, default = mean_weight[1]))
```

```{r}
ggplot(data = squirrels, aes(x = Project_phase, y = Weight_squirrel, colour = Study_site, group = Squirrel_ID)) + 
  geom_line() + 
  geom_point() +
  facet_wrap(~Study_site)

# Mean weight per phase
ggplot(data = weight_phase, aes(x = Project_phase, y = mean_weight, color = Study_site, group = Squirrel_ID)) +
  geom_point(size = 3) +
  geom_line() +
  # geom_errorbar(aes(
  #   ymin = mean_weight - sd_weight,
  #   ymax = mean_weight + sd_weight
  # ), width = 0.2) +
  facet_wrap(~Study_site) +
  labs(
    x = "Project phase",
    y = "Mean weight (g)",
    color = "Study site"#,
    #title = "Mean squirrel weight per project phase"
  )
```

```{r}
ggplot(data = diff_weight, aes(x = Project_phase, y = diff_weight, colour = Squirrel_ID, group = Squirrel_ID)) + 
  geom_line() + 
  geom_point() +
  facet_wrap(~Study_site)
```

## Body length

```{r}
ggplot(squirrels, aes(Study_site, body_length_mm)) +
  geom_boxplot() +
  labs(x = "Study site", y = "Weight (g)",
       title = "Body weight distribution per study site")

ggplot(squirrels, aes(Project_phase, body_length_mm)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Season-Year", y = "Weight (g)",
       title = "Body weight per Season-Year")

ggplot(squirrels, aes(Sex, body_length_mm)) +
  geom_boxplot() +
  labs(x = "Study site", y = "Weight (g)",
       title = "Body weight distribution per study site") +
  facet_wrap(~Study_site, nrow = 3)
```

## Weight and body length

```{r}
ggplot(squirrels, aes(body_length_mm, Weight_squirrel,
                      color = Study_site)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_colour_manual(values = c("Treptower_Park" = "tomato3", 
                               "Lilienthal" = "orange2", 
                               "Rahnsdorf" = "palegreen4"),
                      labels = Study_site_names) +
  labs(x = "Body length (mm)", y = "Weight (g)", colour = "Study site", 
       title = "") + # Relationship between body length and weight
  theme(legend.position = "right")

# ggsave(filename = here::here("plots", "cmr", "weight_length.png"),
#        height = 4, width = 6, units = "in",
#        dpi=800)
```


```{r}
ggplot(squirrels, aes(body_length_mm, Weight_squirrel,
                      color = Study_site)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Body length (mm)", y = "Weight (g)",
       title = "Relationship between body length and weight") + 
  theme(legend.position = "top") +
  facet_wrap(~Study_site)
```

# Samples

```{r}
genetics_site <- squirrels %>%
  count(tissue_sample, Study_site) %>%
  filter(tissue_sample == "yes")
```


```{r}
ggplot(genetics_site, aes(x = Study_site, y = n, fill = Study_site)) +
  geom_col() +
  scale_fill_manual(values = c("Treptower_Park" = "tomato3", 
                               "Lilienthal" = "orange2", 
                               "Rahnsdorf" = "palegreen4")) +
  scale_x_discrete(labels = Study_site_names) +
  labs(x = "Study site", y = "Number of tissue samples",
       title = "") #Tissue samples per site

ggsave(filename = here::here("plots", "cmr", "Tissue_study_sites.png"),
       height = 4, width = 5, units = "in",
       dpi=800)
```

## Summary table

```{r}
table(squirrels$Sex, useNA = "always")
```

Helper function to count individuals per sex (f, m , NA)

```{r}
count_sex <- function(df) {
  df %>%
    mutate(Sex = as.character(Sex),
           Sex = ifelse(is.na(Sex) | Sex == "", "unknown", Sex)) %>%
    count(Sex) %>%
    pivot_wider(names_from = Sex, values_from = n, values_fill = 0)
}
```

Test if the function works

```{r}
count_sex(df = squirrels)
```

```{r}
# 1) Unique captured individuals per phase × site × sex
captured_ind <- squirrels %>%
  distinct(Project_phase, Study_site, Squirrel_ID, Sex) %>%
  group_by(Project_phase, Study_site) %>%
  group_modify(~ count_sex(.x)) %>%
  ungroup() %>%
  rename_with(~ paste0("captured_", .x), c("f", "m", "unknown"))

# 2) Unique recaptured individuals (≥ 2 captures)
recaptured_ind <- squirrels %>%
  count(Squirrel_ID) %>%
  filter(n >= 2) %>%
  select(Squirrel_ID) %>%
  left_join(squirrels %>% select(Squirrel_ID, Sex, Project_phase, Study_site),
            by = "Squirrel_ID") %>%
  distinct() %>%
  group_by(Project_phase, Study_site) %>%
  group_modify(~ count_sex(.x)) %>%
  ungroup() %>%
  rename_with(~ paste0("recaptured_", .x), c("f", "m"))

# 3) Capture events (all rows)
capture_events <- squirrels %>%
  group_by(Project_phase, Study_site) %>%
  group_modify(~ count_sex(.x)) %>%
  ungroup() %>%
  rename_with(~ paste0("events_", .x), c("f", "m", "unknown"))

# 4) Tissue samples (yes only)
tissue_samples <- squirrels %>%
  filter(tissue_sample == "yes") %>%
  group_by(Project_phase, Study_site) %>%
  group_modify(~ count_sex(.x)) %>%
  ungroup() %>%
  rename_with(~ paste0("tissue_", .x), c("f", "m", "unknown"))

# --- Combine everything ---
summary_table <- captured_ind %>%
  full_join(recaptured_ind,
            by = c("Project_phase", "Study_site")) %>%
  full_join(capture_events,
            by = c("Project_phase", "Study_site")) %>%
  full_join(tissue_samples,
            by = c("Project_phase", "Study_site")) %>%
  arrange(Project_phase, Study_site) %>%
  replace(is.na(.), 0)

# Add totals row
totals <- summary_table %>%
  summarise(across(where(is.numeric), sum)) %>%
  mutate(Project_phase = "TOTAL",
         Study_site = "TOTAL")

final_table <- bind_rows(summary_table, totals)

final_table
```

# Docility

```{r}
docility_squirrels <- squirrels %>%
  group_by(Squirrel_ID) %>%
  summarise(n = n(), mean_doc = mean(docility_score))

ggplot(docility_squirrels, aes(mean_doc)) +
  geom_histogram(binwidth = 0.5, boundary = 0) +
  labs(x = "Docility",
       y = "Frequency",
       title = "Docility distribution")
```

```{r}
table(docility_squirrels$mean_doc, useNA = "always")
```

# Nr traps per individual

Into how many different traps did the squirrel go?

## total

```{r}
p2_total <- squirrels %>%
  distinct(Squirrel_ID, Trap_nr, Study_site) %>%
  count(Study_site, Squirrel_ID) %>%
  ggplot(aes(x = n, fill = Study_site)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(
    limits = c(0, 8),
    breaks = seq(1, 7, by = 2),
    expand = c(.004, .001)) +
  scale_y_continuous(
    limits = c(0, 45),
    breaks = seq(0, 40, by = 20),
    expand = c(.004, .001)) +
  facet_wrap(~Study_site) +
  scale_fill_manual(values = site_cols) +
  labs(x = "Number of traps visited", y = "Individuals")

p2_total
```

## total per sex

```{r}
p2_sex <- squirrels %>%
  distinct(Squirrel_ID, Trap_nr, Sex, Study_site) %>%
  count(Study_site, Squirrel_ID, Sex) %>%
  ggplot(aes(x = n, fill = Sex)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(
    limits = c(0, 8),
    breaks = seq(1, 7, by = 2),
    expand = c(.004, .001)) +
  scale_y_continuous(
    limits = c(0, 45),
    breaks = seq(0, 40, by = 20),
    expand = c(.004, .001)) +
  facet_wrap(~Study_site) +
  scale_fill_manual(values = c("f"="#e7298a","m"="#1f78b4","unknown"="grey70")) +
  labs(x = "Number of traps visited", y = "Individuals")

p2_sex
```

## per project phase

```{r}
p2_phase <- squirrels %>%
  distinct(Squirrel_ID, Trap_nr, Study_site, Project_phase) %>%
  count(Project_phase, Study_site, Squirrel_ID) %>%
  ggplot(aes(x = n, fill = Study_site)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(
    limits = c(0, 5),
    breaks = seq(1, 4, by = 1),
    expand = c(.004, .001)) +
  scale_y_continuous(
    limits = c(0, 20),
    breaks = seq(0, 20, by = 10),
    expand = c(.004, .001)) +
  facet_grid(Project_phase ~ Study_site) +
  scale_fill_manual(values = site_cols) +
  labs(x = "Number of traps visited", y = "Individuals")

p2_phase
```

# CE per trap

## total

```{r}
CE_trap <- squirrels %>%
  count(Study_site, Trap_nr)
```

```{r}
 ggplot(data = CE_trap, aes(x = Trap_nr, y = n, fill = Study_site)) +
  geom_col() +
  scale_y_continuous(
    limits = c(0, 35),
    breaks = seq(0, 30, by = 10),
    expand = c(.004, .001)) +
  facet_wrap(~Study_site, scales = "free_x") +
  scale_fill_manual(values = site_cols) +
  labs(y = "Capturing events", x = "Trap") 
```

## total per sex

```{r}
CE_trap_sex <- squirrels %>%
  count(Study_site, Trap_nr, Sex) 
```

```{r}
ggplot(data = CE_trap_sex, aes(x = Trap_nr, y = n, fill = Sex)) +
  geom_col(position = "stack") +
  scale_y_continuous(
    limits = c(0, 35),
    breaks = seq(0, 30, by = 10),
    expand = c(.004, .001)) +
  facet_wrap(~Study_site, scales = "free_x") +
  scale_fill_manual(values = c("f" = "#e7298a","m" = "#1f78b4","unknown" = "grey60")) +
  labs(y = "Capturing events", x = "Trap") 
```

## per project phase

```{r}
p1_phase <- squirrels %>%
  count(Project_phase, Study_site, Trap_nr) %>%
  ggplot(aes(x = Trap_nr, y = n, fill = Study_site)) +
  geom_col() +
  scale_y_continuous(
    limits = c(0, 15),
    breaks = seq(5, 15, by = 10),
    expand = c(.004, .001)) +
  facet_grid(Project_phase ~ Study_site, scales = "free_x") +
  scale_fill_manual(values = site_cols) +
  labs(y = "Capture events", x = "Trap") 

p1_phase
```

# Nr individuals per trap

How many individuals are there per trap

## total

```{r}
p3_total <- squirrels %>%
  distinct(Squirrel_ID, Trap_nr, Study_site) %>%
  count(Study_site, Trap_nr) %>%
  ggplot(aes(x = Trap_nr, y = n, fill = Study_site)) +
  geom_col() +
  scale_y_continuous(
    limits = c(0, 15),
    breaks = seq(0, 15, by = 5),
    expand = c(.004, .001)) +
  facet_wrap(~Study_site, scales = "free_x") +
  scale_fill_manual(values = site_cols) +
  labs(y = "Unique individuals", x = "Trap") 

p3_total
```

## total per sex

```{r}
p3_sex <- squirrels %>%
  distinct(Squirrel_ID, Trap_nr, Sex, Study_site) %>%
  count(Study_site, Trap_nr, Sex) %>%
  ggplot(aes(x = Trap_nr, y = n, fill = Sex)) +
  geom_col() +
  scale_y_continuous(
    limits = c(0, 15),
    breaks = seq(0, 15, by = 5),
    expand = c(.004, .001)) +
  facet_wrap(~Study_site, scales = "free_x") +
  scale_fill_manual(values = c("f"="#e7298a","m"="#1f78b4","unknown"="grey60")) +
  labs(y = "Unique individuals", x = "Trap") 

p3_sex
```

## per project phase

```{r}
p3_phase <- squirrels %>%
  distinct(Squirrel_ID, Trap_nr, Study_site, Project_phase) %>%
  count(Project_phase, Study_site, Trap_nr) %>%
  ggplot(aes(x = Trap_nr, y = n, fill = Study_site)) +
  geom_col() +
  scale_y_continuous(
    limits = c(0, 7),
    breaks = seq(0, 6, by = 3),
    expand = c(.004, .001)) +
  facet_grid(Project_phase ~ Study_site, scales = "free_x") +
  scale_fill_manual(values = site_cols) +
  labs(y = "Unique individuals", x = "Trap") 

p3_phase
```

# Trap nr x project phase

```{r}
p4_heatmap <- squirrels %>%
  count(Study_site, Project_phase, Trap_nr) %>%
  ggplot(aes(x = Trap_nr, y = factor(Project_phase), fill = n)) +
  geom_tile(color = "white") +
  facet_wrap(~Study_site, scales = "free_x") +
  scale_fill_viridis_c(option = "magma") +
  labs(x = "Trap", y = "Project phase", fill = "Captures") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

p4_heatmap
```

# Capture activity per trap per day

## per day per study site

```{r}
p5_daily <- squirrels %>%
  count(Study_site, Date) %>%
  ggplot(aes(x = Date, y = n, color = Study_site)) +
  geom_point() +
  scale_color_manual(values = site_cols) +
  facet_wrap(~Study_site) +
  labs(y = "Captures per day", x = "Date")

p5_daily
```

## per trap per day

```{r}
p5_daily_trap <- squirrels %>%
  count(Study_site, Date, Trap_nr) %>%
  ggplot(aes(x = Date, y = n, color = Study_site, group = Trap_nr)) +
  geom_point() +
  facet_wrap(~Study_site) +
  scale_color_manual(values = site_cols) +
  labs(y = "Captures per trap per day", x = "Date")

p5_daily_trap
```

# Trapping effort

```{r}
str(trapping_effort)
```

```{r}
table(trapping_effort$opened)
table(trapping_effort$opened > 0, trapping_effort$Trapping_Duration)
table(trapping_effort$opened > 0, trapping_effort$hours_day)
```

## Trap status over time

```{r}
ggplot(trapping_effort, aes(x = Project_phase, y = Trap_nr, fill = opened)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "steelblue") +
  facet_wrap(~ Study_site, scales = "free_y") +
  labs(
    title = "Trap status over time",
    x = "Date",
    y = "Trap number",
    fill = "Opened"
  ) 
```

## Hours activated per trap

### total

```{r}
effort_hours <- trapping_effort %>%
  group_by(Study_site, Trap_nr) %>%
  summarise(total_hours = sum(hours_day, na.rm = TRUE)) %>%
ggplot(aes(x = Trap_nr, y = total_hours)) +
  geom_col() +
  scale_y_continuous(
    limits = c(0, 180),
    breaks = seq(0, 180, by = 60),
    expand = c(.004, .001)) +
  facet_wrap(~ Study_site, scales = "free_x") +
  labs(
    title = "Total hours of trap activity per trap",
    x = "Trap number",
    y = "Total hours open"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

effort_hours
```

Heatmap per trap

```{r}
ggplot(trapping_effort, aes(x = Project_phase, y = Trap_nr, fill = Trapping_Duration)) +
  geom_tile() +
  scale_fill_viridis_c() +
  facet_wrap(~ Study_site, scales = "free_y") +
  labs(
    title = "Active trapping duration per trap and day",
    x = "Date",
    y = "Trap number",
    fill = "Hours"
  ) 
```

### per project phase

```{r}
effort_hours_phase <- trapping_effort %>%
  group_by(Study_site, Project_phase, Trap_nr) %>%
  summarise(total_hours = sum(hours_day, na.rm = TRUE)) %>%
ggplot(aes(x = Trap_nr, y = total_hours)) +
  geom_col() +
  scale_y_continuous(
    limits = c(0, 40),
    breaks = seq(10, 40, by = 30),
    expand = c(.004, .001)) +
  facet_grid(Project_phase ~ Study_site, scales = "free_x") +
  labs(
    title = "Total hours of trap activity per trap",
    x = "Trap number",
    y = "Total hours open"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

effort_hours_phase
```

## Hours per day

```{r}
effort_day <- trapping_effort %>%
  group_by(Study_site, Date) %>%
  summarise(total_hours_day = sum(hours_day, na.rm = TRUE))

ggplot(effort_day, aes(x = Date, y = total_hours_day)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ Study_site) +
  labs(
    title = "Total trapping hours per day",
    x = "Date",
    y = "Total hours"
  ) 
```

## Open traps per day

```{r}
opened_day <- trapping_effort %>%
  group_by(Study_site, Date) %>%
  summarise(
    traps_open = sum(opened == 1),
    traps_total = n(),
    #prop_open = traps_open / traps_total
  )

ggplot(opened_day, aes(x = Date, y = traps_open)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(
    limits = c(0, 10.3),
    breaks = seq(0, 10, by = 5),
    expand = c(.004, .001)) +
  facet_wrap(~ Study_site) +
  labs(
    title = "Proportion of traps open per day",
    y = "Proportion open"
  ) 
```

## Average daily hours per trap

```{r}
avg_hours <- trapping_effort %>%
  group_by(Study_site, Trap_nr) %>%
  summarise(avg_hours = mean(hours_day, na.rm = TRUE))

ggplot(avg_hours, aes(x = Trap_nr, y = avg_hours)) +
  geom_col() +
  scale_y_continuous(
    limits = c(0, 5),
    breaks = seq(0, 5, by = 2.5),
    expand = c(.004, .001)) +
  facet_wrap(~ Study_site, scales = "free_x") +
  labs(
    title = "Average daily trapping hours per trap",
    x = "Trap number",
    y = "Average hours"
  ) + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5))
```

Variation per trap

```{r}
ggplot(trapping_effort, aes(x = Trap_nr, y = hours_day)) +
  geom_boxplot() +
  scale_y_continuous(
    limits = c(-0.2, 7),
    breaks = seq(0, 7, by = 3.5),
    expand = c(.004, .001)) +
  facet_wrap(~ Study_site, scales = "free_x") +
  labs(
    title = "Variation in trapping duration per trap",
    x = "Trap number",
    y = "Duration (hours)"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = .5))
```

# Human activity
```{r}
hum_act_means <- hum_act_means %>%
  mutate(Study_site = factor(
    Study_site,
    levels = c("Rahnsdorf", "Lilienthal", "Treptower_Park")
  ))

Study_site_names <- c(
  "Rahnsdorf" = "woodland area",
  "Treptower_Park" = "urban park",
  "Lilienthal" = "cemetery"
)
```


```{r}
ggplot(hum_act_means, aes(x = Study_site, y = mean_30, , fill = Study_site)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(labels = Study_site_names) +
  labs(
    x = "Study site",
    y = "Human activity index"
  ) +
  scale_fill_manual(values = c("Treptower_Park" = "tomato3", 
                               "Lilienthal" = "orange2", 
                               "Rahnsdorf" = "palegreen4")) 

# ggsave(filename = here::here("plots", "cmr", "hum_act_20260209.png"),
#        height = 4, width = 6, units = "in",
#        dpi=800)
```



```{r}
ggplot(data = hum_act_means, mapping = aes(x = study_site_specific, y = spec_mean_30, group = study_site_specific, colour = study_site_specific)) +
  geom_point()
```

***

<details>

<summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>
