---
title: "Squirrel genetics and survival: Trap locations" 
description: |
    This script creates maps of the study area, including an overview of Berlin
    and inset maps of the three study sites.
author:
    - name: "Sinah Drenske"
      affiliation: "Leibniz Institute for Zoo and Wildlife Research (IZW); Technische Universität Berlin"
      affiliation_url: https://www.izw-berlin.de/en/
      orcid_id: 0000-0002-2247-6507
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
        highlight: kate      ## styling of code
        code_folding: false  ## if `true` you can expand and shrink code chunks
        toc: true            ## if `true` adds a table of content
        toc_depth: 2         ## level to be displayed in the table of content
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      dev = "ragg_png", fig.width = 9, fig.height = 6, dpi = 600, retina = 1)
Sys.setlocale("LC_TIME", "C")
```

* **Research questions:**
  * Do Eurasian red squirrel populations inhabiting sites with different levels of urbanisation in Berlin differ in genetic diversity, population structure, and connectivity?
  * Does apparent survival of Eurasian red squirrels differ among sites with different levels of urbanisation in Berlin?
* **Study area:** Three study sites in Berlin, Germany, representing different levels of urbanisation: a woodland site in Rahnsdorf, Lilienthal Cemetery, and Treptower Park.
  * Woodland = Rahnsdorf
  * Cemetery = Lilienthal Cemetery
  * Urban park = Treptower Park
* **Data:** Spatial data for Berlin and the three study sites, including trap location data and map data used to visualise the study area and site-specific insets.


# Setup

```{r packages}
## libraries used in this script
library("sf")
library("cowplot")
library("d6berlin")
library("extrafont")
library("ggplot2")
library("ggdist")
library("grid")
library("geodata")
library("dplyr")
library("osmdata")
library("patchwork")
library("ggspatial")
library("rnaturalearth")
library("svglite")
library("tmap")
library("viridis")
```

```{r message=FALSE, include=FALSE}
# font_import()
# loadfonts(device = "win")
# fonts()
```

# Data

```{r data}
## To import tabular data (.csv, .txt, .xls etc.)
traps_df <- vroom::vroom(file = here::here("data-raw", "cmr-raw", "Trap_locations_20250813.csv"), locale = vroom::locale(decimal_mark = ","))

# geo data
## impervious surface
imperv <- rast(here::here("data-raw/geo-raw/imperviousness_berlin_copernicus_raster_10m_2018_3035/imperviousness_berlin_copernicus_raster_10m_2018_3035.tif"))

## water bodies
water <- st_read(here::here("data-raw/geo-raw/waterbodies_Berlin_25833.gpkg")) %>% 
  st_transform(crs = 3035) %>% 
  st_union()

## berlin border
berlin <- st_read(here::here("data-raw/geo-raw/berlin_city_border_25833.gpkg")) %>% 
  st_transform(crs = 3035) %>% 
  st_union()
```

Labeller for study sites
```{r}
# Remove the underscore in the name of the study site without changing the factor levels
Study_site_names <- as_labeller(c("Rahnsdorf" = "woodland area", "Treptower_Park" = "urban park", "Lilienthal" = "cemetery"))
```

# Format data
Delete columns
```{r}
traps_df <- traps_df %>%
  select(Study_site, Trap_nr, X, Y)
```

```{r}
traps <- st_as_sf(traps_df,
                  coords = c("X", "Y"),
                  crs = 4326) # WGS84
traps_3035 <- st_transform(traps, 3035)

```

# Berlin Map
```{r}
imperv_250 <- aggregate(imperv, fact = 25, fun = mean, na.rm = TRUE)

# get in data frame format
imperv250_df <- as.data.frame(imperv_250, xy = TRUE)
colnames(imperv250_df)[3] <- "imperv"

imperv_df <- as.data.frame(imperv, xy = TRUE)
colnames(imperv_df)[3] <- "imperv"

no_arrow_style_b <- north_arrow_orienteering(
  line_width = 1,
  line_col = "black",
  fill = c("black", "black"),
  text_col = "black",
  text_family = "",
  text_face = NULL,
  text_size = 10,
  text_angle = 0
)
```

```{r}
# ggplot() +
#   geom_raster(data = imperv_df, aes(x = x, y = y, fill = imperv))
# 
# ggplot() +
#   geom_raster(data = imperv250_df, aes(x = x, y = y, fill = imperv))
```


```{r}
berlin_map <- 
  ggplot() +
  geom_raster(data = imperv_df, aes(x = x, y = y, fill = imperv)) +
  scale_fill_viridis(option = "rocket", na.value = NA, direction = -1,
                     #begin = 0.2, 
                     end = 0.9, 
                     alpha = 0.3,
                     name = "Imperviousness") +
  geom_sf(data = water, fill = "lightblue3", colour = "transparent") + 
  geom_sf(data = traps_3035, size = 4, color = "black") + # aes(shape = Study_site), 
  geom_sf(data = berlin, colour = "grey40", lwd = 0.7, fill = NA) +
  annotation_north_arrow(which_north = "true", location = "bl",
                         pad_x = unit(0.3, "cm"), pad_y = unit(0.7, "cm"),
                         height = unit(1, "cm"),
                         width = unit(0.6, "cm"), style = no_arrow_style_b) +
  annotation_scale(text_cex = 1.5) +
scale_x_continuous(breaks = seq(13.1, 13.7, by = 0.3)) +
scale_y_continuous(breaks = seq(52.35, 52.70, by = 0.15)) +
  theme(
    legend.position = "right",
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    plot.margin = margin(6, 10, 4, 4),
    axis.title = element_blank(),
    axis.ticks = element_line(colour = "black", linewidth = 1.5),
    axis.ticks.length = unit(0.2, "cm"),
    axis.text = element_text(size = 18),
    legend.text = element_text(size = 12), 
    legend.title = element_text(size = 14)
   ) #+ 
  # #scale_x_discrete(limits=c("2", "0.5", "1")) + 
  # scale_shape_discrete(name = "Study site", 
  #                      limits = c("Rahnsdorf", "Lilienthal", "Treptower_Park"), 
  #                      labels = c("woodland area", "cemetery", "urban park")) #

berlin_map
```

## Map of Europe

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
europe <- world %>% 
  filter(continent  == "Europe")

europe_crop <- europe %>% 
   st_make_valid() %>% 
    st_crop(xmin = -15, ymin = 25, xmax = 30, ymax = 55) # xmin = -25, ymin = 25, xmax = 50, ymax = 70

# europe_crop2 <- europe %>% 
#    st_make_valid() %>% 
#     st_crop(xmin = -25, ymin = 25, xmax = 50, ymax = 70) # xmin = -25, ymin = 25, xmax = 50, ymax = 70

centroid_berlin <- berlin %>% 
  st_transform(crs = 4326) %>% 
  st_centroid()
```


```{r }
# fig.width = 9, fig.height = 6
  
eu_map <- 
  ggplot() +
  geom_sf(data = europe_crop, col = "black", fill = "grey90") +
  geom_sf(data = centroid_berlin, size = 6, pch = 22, linewidth = 3, fill = "red", col = "red") +
  geom_sf(data = centroid_berlin, col = "red", size = 4, pch = 22, fill = "grey90") +
  coord_sf(xlim = c(-15,30.1), ylim = c(35,55.1), expand = FALSE) +
  scale_x_continuous(breaks = seq(-10, 25, by = 15)) + 
  scale_y_continuous(breaks = seq(40, 50, by = 10)) + 
  theme(
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "transparent", colour = NA),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    plot.margin = margin(6, 10, 4, 4),
    axis.title = element_blank(),
    axis.ticks = element_line(colour = "black", linewidth = 1.5),
    axis.ticks.length = unit(0.2, "cm"),
    #title = element_text(family = "Zilla Slab", size = 15), # axis
    axis.text = element_text(size = 13)
    ) # axis ticks

eu_map
#dev.off()
```

## both together
```{r}
#dev.off()

plot.with.inset <-
  ggdraw() +
  draw_plot(berlin_map) +
  draw_plot(eu_map, x = 0.512, y = .665, width = .29, height = .29)

plot.with.inset

# Save the plot
ggsave(filename = here::here("plots", "maps",
                             "Study_area_20250903.tiff"),
         # width = 85,
         # height = 85,
         # units = "mm",
       dpi = 300)
```

# Study sites

```{r}
tp_traps <- traps_3035 %>% filter(Study_site == "Treptower_Park")
lilie_traps <- traps_3035 %>% filter(Study_site == "Lilienthal")
rahn_traps <- traps_3035 %>% filter(Study_site == "Rahnsdorf")

no_arrow_style <- north_arrow_orienteering(
  line_width = 1,
  line_col = "black",
  fill = c("black", "black"),
  text_col = "black",
  text_family = "",
  text_face = NULL,
  text_size = 15,
  text_angle = 0
)
```

## Treptow
```{r}
# Prepare the map
#bbox <- st_bbox(st_buffer(st_union(tp_traps), 500))

center <- st_centroid(st_union(tp_traps))

# feste Seitenlänge (z. B. 1000 m = 1 km Kante)
half_size <- 800  # Hälfte der Seitenlänge

bbox <- st_bbox(c(
  xmin = st_coordinates(center)[1] - half_size,
  xmax = st_coordinates(center)[1] + half_size,
  ymin = st_coordinates(center)[2] - half_size,
  ymax = st_coordinates(center)[2] + half_size
), crs = st_crs(tp_traps))

imperv_crop <- crop(imperv, ext(bbox$xmin, bbox$xmax, bbox$ymin, bbox$ymax))
imperv_tp   <- as.data.frame(imperv_crop, xy = TRUE)
names(imperv_tp)[3] <- "imperv"
berlin_crop <- st_crop(berlin, bbox)
water_crop <- st_crop(water, bbox)
```


```{r}
# Plot it
p_tp <- ggplot() +
  geom_raster(data = imperv_tp, aes(x = x, y = y, fill = imperv)) +
  scale_fill_viridis(option = "rocket", na.value = NA,
                     direction = -1, end = 0.9,
                     alpha = 0.3,
                     name = "Imperviousness") +
  geom_sf(data = water_crop, fill = "lightblue3", colour = NA) +
  geom_sf(data = traps_3035 %>% filter(Study_site == "Treptower_Park"),
          aes(shape = Study_site), size = 5, color = "black") +
  geom_sf(data = berlin_crop, colour = "grey40", lwd = 0.7, fill = NA) +
  annotation_north_arrow(which_north = "true", location = "bl",
                         pad_x = unit(0.8, "cm"), pad_y = unit(1.6, "cm"),
                         height = unit(2.5, "cm"),
                         width = unit(1.7, "cm"), style = no_arrow_style) +
  annotation_scale(text_cex = 3, 
                   pad_x = unit(0.8, "cm"),
                   pad_y = unit(1.2, "cm"),) +
  scale_x_continuous(breaks = seq(13.46 , 13.48 , by = 0.02)) +
  scale_y_continuous(breaks = seq(52.48, 52.49, by = 0.01)) +
  theme(
    legend.position = "none", 
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    plot.margin = margin(6, 10, 4, 4),
    axis.title = element_blank(),
    axis.ticks = element_line(colour = "black", linewidth = 1.5),
    axis.ticks.length = unit(0.2, "cm"),
    axis.text = element_text(size = 30)
  )
p_tp

# # Save plot
# ggsave(filename = here::here("plots", "maps",
#                              "Treptow_20250903.tiff"),
#        # width = 85,
#        # height = 85,
#        # units = "mm",
#        dpi = 300)
```

## Lilie
```{r}
# Prepare the map
#bbox <- st_bbox(st_buffer(st_union(lilie_traps), 500))

center <- st_centroid(st_union(lilie_traps))

# feste Seitenlänge (z. B. 1000 m = 1 km Kante)
half_size <- 800  # Hälfte der Seitenlänge

bbox <- st_bbox(c(
  xmin = st_coordinates(center)[1] - half_size,
  xmax = st_coordinates(center)[1] + half_size,
  ymin = st_coordinates(center)[2] - half_size,
  ymax = st_coordinates(center)[2] + half_size
), crs = st_crs(lilie_traps))

imperv_crop <- crop(imperv, ext(bbox$xmin, bbox$xmax, bbox$ymin, bbox$ymax))
imperv_tp   <- as.data.frame(imperv_crop, xy = TRUE)
names(imperv_tp)[3] <- "imperv"
berlin_crop <- st_crop(berlin, bbox)
water_crop <- st_crop(water, bbox)
```


```{r}
# Plot it
p_lilie <- ggplot() +
  geom_raster(data = imperv_tp, aes(x = x, y = y, fill = imperv)) +
  scale_fill_viridis(option = "rocket", na.value = NA,
                     direction = -1, end = 0.9,
                     alpha = 0.3,
                     name = "Imperviousness") +
  geom_sf(data = water_crop, fill = "lightblue3", colour = NA) +
  geom_sf(data = traps_3035 %>% filter(Study_site == "Lilienthal"),
          aes(shape = Study_site), size = 5, color = "black") +
  geom_sf(data = berlin_crop, colour = "grey40", lwd = 0.7, fill = NA) +
  annotation_north_arrow(which_north = "true", location = "bl",
                         pad_x = unit(0.8, "cm"), pad_y = unit(1.6, "cm"),
                         height = unit(2.5, "cm"),
                         width = unit(1.7, "cm"), style = no_arrow_style) +
  annotation_scale(text_cex = 3, 
                   pad_x = unit(0.8, "cm"),
                   pad_y = unit(1.2, "cm"),) +
  scale_x_continuous(breaks = seq(13.40 , 13.42 , by = 0.02)) +
  scale_y_continuous(breaks = seq(52.48, 52.49, by = 0.01)) +
  theme(
    legend.position = "none", 
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    plot.margin = margin(6, 10, 4, 4),
    axis.title = element_blank(),
    axis.ticks = element_line(colour = "black", linewidth = 1.5),
    axis.ticks.length = unit(0.2, "cm"),
    axis.text = element_text(size = 30)
  )
p_lilie

# Save plot
ggsave(filename = here::here("plots", "maps",
                             "Lilie_20250814.tiff"),
         # width = 85,
         # height = 85,
         # units = "mm",
       dpi = 300)
```

## Rahnsdorf
```{r}
# Prepare the map
#bbox <- st_bbox(st_buffer(st_union(rahn_traps), 500))
center <- st_centroid(st_union(rahn_traps))

# feste Seitenlänge (z. B. 1000 m = 1 km Kante)
half_size <- 800  # Hälfte der Seitenlänge

bbox <- st_bbox(c(
  xmin = st_coordinates(center)[1] - half_size,
  xmax = st_coordinates(center)[1] + half_size,
  ymin = st_coordinates(center)[2] - half_size,
  ymax = st_coordinates(center)[2] + half_size
), crs = st_crs(rahn_traps))

imperv_crop <- crop(imperv, ext(bbox$xmin, bbox$xmax, bbox$ymin, bbox$ymax))
imperv_tp   <- as.data.frame(imperv_crop, xy = TRUE)
names(imperv_tp)[3] <- "imperv"
berlin_crop <- st_crop(berlin, bbox)
water_crop <- st_crop(water, bbox)

# Plot it
p_rahn <- ggplot() +
  geom_raster(data = imperv_tp, aes(x = x, y = y, fill = imperv)) +
  scale_fill_viridis(option = "rocket", na.value = NA,
                     direction = -1, end = 0.9,
                     alpha = 0.3,
                     name = "Imperviousness") +
  geom_sf(data = water_crop, fill = "lightblue3", colour = NA) +
  geom_sf(data = traps_3035 %>% filter(Study_site == "Rahnsdorf"),
          aes(shape = Study_site), size = 5, color = "black") +
  geom_sf(data = berlin_crop, colour = "grey40", lwd = 0.7, fill = NA) +
  annotation_north_arrow(which_north = "true", location = "bl",
                         pad_x = unit(0.8, "cm"), pad_y = unit(1.6, "cm"),
                         height = unit(2.5, "cm"),
                         width = unit(1.7, "cm"), style = no_arrow_style) +
  annotation_scale(text_cex = 3, 
                   pad_x = unit(0.8, "cm"),
                   pad_y = unit(1.2, "cm"),) +
  scale_x_continuous(breaks = seq(13.68 , 13.70 , by = 0.02)) +
  scale_y_continuous(breaks = seq(52.45, 52.46, by = 0.01)) +
  theme(
    legend.position = "none", 
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    plot.margin = margin(6, 10, 4, 4),
    axis.title = element_blank(),
    axis.ticks = element_line(colour = "black", linewidth = 1.5),
    axis.ticks.length = unit(0.2, "cm"),
    axis.text = element_text(size = 30)
  )
p_rahn

# Save plot
ggsave(filename = here::here("plots", "maps",
                             "Rahnsdorf_20250814.tiff"),
         # width = 85,
         # height = 85,
         # units = "mm",
       dpi = 300)
```


***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>
