---
title: "Squirrel genetics and survival: Kinship analysis" 
description: |
    This script calculates pairwise genetic relatedness among individual red
    squirrels using the Queller and Goodnight estimator and identifies closely
    related individuals for exclusion from subsequent population structure analyses.
author:
    - name: "Sinah Drenske"
      affiliation: "Leibniz Institute for Zoo and Wildlife Research (IZW); Technische Universität Berlin"
      affiliation_url: https://www.izw-berlin.de/en/
      orcid_id: 0000-0002-2247-6507
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
        highlight: kate      ## styling of code
        code_folding: false  ## if `true` you can expand and shrink code chunks
        toc: true            ## if `true` adds a table of content
        toc_depth: 2         ## level to be displayed in the table of content
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      dev = "ragg_png", fig.width = 9, fig.height = 6, dpi = 600, retina = 1)
Sys.setlocale("LC_TIME", "C")
```

* **Research question:** Do Eurasian red squirrel populations inhabiting sites with different levels of urbanisation in Berlin differ in genetic diversity, population structure, and connectivity?
* **Study area:** Three study sites in Berlin, Germany, representing different levels of urbanisation: a woodland site in Rahnsdorf, Lilienthal Cemetery, and Treptower Park.
  * Woodland = Rahnsdorf
  * Cemetery = Lilienthal Cemetery
  * Urban park = Treptower Park
* **Data:** Microsatellite genotypes and study-site assignments of individual Eurasian red squirrels from the three study sites.


# Setup

```{r packages}
## for non-CRAN packages please keep install instruction
# For the package related: please follow the instructions here: https://github.com/timothyfrasier/related 

## libraries used in this script
library("related")
library("tidyverse")
```


# Data

```{r data}
# Genetic and CMR data
gen_squirrels_135_IDs <- readRDS(file = here::here("output", "data-proc", "genetic-proc", "cluster_analysis",  
                                                   "gen_squirrels_135_IDs_20250719.RDS"))
gen_squirrels_135_IDs <- gen_squirrels_135_IDs %>%
  mutate(Study_site = factor(Study_site, levels = c("Rahnsdorf", "Lilienthal", "Treptower_Park")))

# Kinship data
# Txt files as tables
kinship_squirrels_raw <- vroom::vroom(here::here("data-raw", "genetic-raw",
                                                 "Squirrel_Genotypes_135_IDs_kinship_20250819.txt"),
                                      col_names = FALSE)
```

## data preparation with study sites
```{r}
# Protocol nr as column and transform to numeric
gen_squirrels_135_IDs <- gen_squirrels_135_IDs %>% 
  rownames_to_column(var = "protocol_nr_old") %>% 
  mutate(protocol_nr_old = as.numeric(protocol_nr_old))

# Compare the order of individuals in both datasets (X1 and protocol nr old are the sample ID)
kinship_squirrels_raw$X1 == gen_squirrels_135_IDs$protocol_nr_old
kinship_squirrels_raw$X1[c(1:30)]
gen_squirrels_135_IDs$protocol_nr_old[c(1:30)]
# Test object for a join of the sample ID and the study site
test <- left_join(x = kinship_squirrels_raw[,1], y = gen_squirrels_135_IDs[, c(1,7)], by = join_by("X1" == "protocol_nr_old"))
test

# Use the first 2 letters of the study sites per sample
prefix <- substr(gen_squirrels_135_IDs$Study_site, 1, 2)
prefix
# create a column for the combination of sample ID and study site
kinship_squirrels_raw$X34 <- paste0(prefix, "_", kinship_squirrels_raw$X1)
kinship_squirrels_raw[, c(1,34)]
kinship_squirrels_raw <- left_join(x = kinship_squirrels_raw[], y = gen_squirrels_135_IDs[, c(1,7)], by = join_by("X1" == "protocol_nr_old"))
# kinship_squirrels_raw$X1
# kinship_squirrels_raw$X1 <- paste0(prefix, "_", kinship_squirrels_raw$X1)
# kinship_squirrels_raw$X1
```
Split per study site

```{r}
kinship_wood_raw <- kinship_squirrels_raw %>% filter(Study_site == "Rahnsdorf") %>% select(-c(34,35))
kinship_cemetery_raw <- kinship_squirrels_raw %>% filter(Study_site == "Lilienthal") %>% select(-c(34,35))
kinship_park_raw <- kinship_squirrels_raw %>% filter(Study_site == "Treptower_Park") %>% select(-c(34,35))
```

## Save data
```{r}
# # Save this as data set with an ID including the study sites
# vroom::vroom_write(x = kinship_squirrels_raw,
#                    file = here::here("output", "data-proc", "genetic-proc", "individual_genetics",
#                                      "kinship_squirrels_study_site_20260210.txt"))
# 
# # woodland
# vroom::vroom_write(x = kinship_wood_raw,
#                    file = here::here("output", "data-proc", "genetic-proc", "individual_genetics",
#                                      "kinship_squirrels_woodland_20260210.txt"), 
#                    col_names = F)
# 
# # cemetery
# vroom::vroom_write(x = kinship_cemetery_raw,
#                    file = here::here("output", "data-proc", "genetic-proc", "individual_genetics",
#                                      "kinship_squirrels_cemetery_20260210.txt"), 
#                    col_names = F)
# 
# # urban park
# vroom::vroom_write(x = kinship_park_raw,
#                    file = here::here("output", "data-proc", "genetic-proc", "individual_genetics",
#                                      "kinship_squirrels_park_20260210.txt"), 
#                    col_names = F)
```

# Data for analysis
```{r}
# Data for analysis as genotype data
kinship_squirrels <- readgenotypedata(genotype.data = here::here("data-raw", "genetic-raw",
                                                                 "Squirrel_Genotypes_135_IDs_kinship_20250819.txt"))

# Woodland
kinship_wood <- readgenotypedata(genotype.data = here::here("output", "data-proc", "genetic-proc", "individual_genetics",
                                                            "kinship_squirrels_woodland_20260210.txt"))

# Cemetery
kinship_cemetery <- readgenotypedata(genotype.data = here::here("output", "data-proc", "genetic-proc", "individual_genetics",
                                                            "kinship_squirrels_cemetery_20260210.txt"))

# Urban park
kinship_park <- readgenotypedata(genotype.data = here::here("output", "data-proc", "genetic-proc", "individual_genetics",
                                                            "kinship_squirrels_park_20260210.txt"))

# kinship_site_squirrels <- readgenotypedata(genotype.data = here::here("output", "data-proc", "genetic-proc",
#                                                                       "individual_genetics",
#                                                                       "kinship_squirrels_study_site_20250819.txt"))

# kinship_site_squirrels <- vroom::vroom(here::here("output", "data-proc", "genetic-proc", "individual_genetics",
#                                                   "kinship_squirrels_study_site_20250819.txt"),
#                                        col_names = FALSE)
```


# Kinship analysis
## Test data
Load data
```{r}
# data(simrel)
# #---Read data into R---#
# data(GenotypeData)
# input <- readgenotypedata(GenotypeData)
```

Calculate relatedness 
```{r}
#---Calculate Relatedness---#
#output <- coancestry(input$gdata, lynchrd = 2, quellergt = 2, wang = 2)
```

Look at the results
```{r}
# #---View Point Estimates---#
# output$relatedness
# 
# #---View 95
# output$relatedness.ci95
```

## Squirrels: All sites together
Load data and check for NAs
```{r}
squirrel_gdata <- kinship_squirrels$gdata
squirrel_gdata
# rows = individuals, columns = ID, 2 alleles of 16 loci
# table(squirrel_gdata$V1, useNA = "always") # allel 1 locus 1
# table(squirrel_gdata$V33, useNA = "always") # allel 2 locus 16
```


Calculate relatedness
```{r}
#output <- coancestry(input$gdata, lynchrd = 2, quellergt = 2, wang = 2)
related_output <- coancestry(genotype.data = kinship_squirrels$gdata, 
                             error.rates = 0.01, 
                             #wang = 2, 
                             quellergt = 2,
                             ci95.num.bootstrap = 500, #1st version with 1000 repetitions but now the session aborts every time with 1000
                             rng.seed = 20200604
                             )
```

Look at output
```{r}
relatedness <- related_output$relatedness
# summary(relatedness$wang)
summary(relatedness$quellergt)
# table(relatedness$wang > 0.5)
table(relatedness$quellergt > 0.5)
# boxplot(relatedness$wang)
boxplot(relatedness$quellergt)

# wang_related <- relatedness %>% filter(wang >= 0.5)
qg_related <- relatedness %>% filter(quellergt > 0.5)
```


```{r}
# # Wang
# table(wang_related$ind1.id)
# table(wang_related$ind2.id)
# 
# all_ids_wang <- c(wang_related$ind1.id, wang_related$ind2.id)
# id_counts_wang <- as.data.frame(table(all_ids_wang))
# colnames(id_counts_wang) <- c("ID", "count")

# Queller & Goodnight
table(qg_related$ind1.id)
table(qg_related$ind2.id)

all_ids_qg <- c(qg_related$ind1.id, qg_related$ind2.id)
id_counts_qg <- as.data.frame(table(all_ids_qg))
colnames(id_counts_qg) <- c("ID", "count")
```


```{r}
# ## Wang
# wang_related$remove <- apply(wang_related, 1, function(row) {
#   id1 <- row["ind1.id"]
#   id2 <- row["ind2.id"]
#   
#   c1 <- id_counts_wang$count[id_counts_wang$ID == id1]
#   c2 <- id_counts_wang$count[id_counts_wang$ID == id2]
#   
#   if (c1 > c2) {
#     return(id1)
#   } else if (c2 > c1) {
#     return(id2)
#   } else {
#     # Falls beide gleich oft vorkommen -> alphabetisch die kleinere ID nehmen
#     return(min(id1, id2))
#   }
# })
# 
# # Alle rauszunehmenden IDs
# ids_to_remove_wang <- unique(wang_related$remove)
# #ids_to_remove <- order(ids_to_remove, decreasing = FALSE)
# ids_to_remove_wang
# 

## Queller & Goodnight
qg_related$remove <- apply(qg_related, 1, function(row) {
  id1 <- row["ind1.id"]
  id2 <- row["ind2.id"]
  
  c1 <- id_counts_qg$count[id_counts_qg$ID == id1]
  c2 <- id_counts_qg$count[id_counts_qg$ID == id2]
  
  if (c1 > c2) {
    return(id1)
  } else if (c2 > c1) {
    return(id2)
  } else {
    # Falls beide gleich oft vorkommen -> alphabetisch die kleinere ID nehmen
    return(min(id1, id2))
  }
})

# Alle rauszunehmenden IDs
ids_to_remove_qg <- unique(qg_related$remove)
#ids_to_remove <- order(ids_to_remove, decreasing = FALSE)
ids_to_remove_qg
```


### Remove IDs
Who are the indivudals we have to remove?
```{r}
#gen_squirrels_135_IDs <- rownames_to_column(.data = gen_squirrels_135_IDs, var = "protocol_nr2") # Why???
ids_remove_qg <- data.frame(protocol_nr_old = ids_to_remove_qg) %>% 
  mutate(protocol_nr_old = as.numeric(protocol_nr_old))
ids_remove <- left_join(x = ids_remove_qg, y = gen_squirrels_135_IDs, by = "protocol_nr_old")
remove <- data.frame(ID = as.numeric(ids_to_remove_qg))
```

Order by study site, sex, season
```{r}
table(ids_remove$Study_site, useNA = "always")
table(ids_remove$Study_site, ids_remove$Sex, useNA = "always")
table(ids_remove$Season, useNA = "always")
```

calculate the percentage of individuals that has to be removed
```{r}
# noch anders anordnen
table(gen_squirrels_135_IDs$Study_site, gen_squirrels_135_IDs$Sex)
#table(gen_squirrels_93_IDs$Study_site, gen_squirrels_93_IDs$Sex)
table(ids_remove$Study_site, ids_remove$Sex)
```


```{r}
## Treptow
# all together
25/65
# males
15/30
# females
10/34

## Lilienthal
# all together
15/50 
# males
8/25
# females
7/24

## Rahnsdorf
# all together
2/20 
# males
2/12
# no females were removed
```

Urban park (Treptow) 
38% of all individuals from Treptow
50% of the males
29% of females

Cemetery (Lilienthal)
30% of all individuals from Treptow
32% of the males
29% of females

Woodland area (Rahnsdorf)
10% of all individuals from Treptow
17% of the males
0% of females

## Woodland
Load data and check for NAs
```{r}
wood_gdata <- kinship_wood$gdata#[-1, ]
# rows = individuals, columns = ID, 2 alleles of 16 loci
# table(squirrel_gdata$V1, useNA = "always") # allel 1 locus 1
# table(squirrel_gdata$V33, useNA = "always") # allel 2 locus 16
wood_gdata
squirrel_gdata
```


Calculate relatedness
```{r}
#output <- coancestry(input$gdata, lynchrd = 2, quellergt = 2, wang = 2)
related_wood_output <- coancestry(genotype.data = kinship_wood$gdata, #[-1, ], 
                             error.rates = 0.01, 
                             #wang = 2, 
                             quellergt = 2,
                             ci95.num.bootstrap = 500, #1st version with 1000 repetitions but now the session aborts every time with 1000
                             rng.seed = 20200604
                             )
```

Look at output
```{r}
relatedness_wood <- related_wood_output$relatedness
# summary(relatedness$wang)
summary(relatedness_wood$quellergt)
# table(relatedness$wang > 0.5)
table(relatedness_wood$quellergt > 0.5)
# boxplot(relatedness$wang)
boxplot(relatedness_wood$quellergt)

# wang_related <- relatedness %>% filter(wang >= 0.5)
qg_related_wood <- relatedness_wood %>% filter(quellergt > 0.5)
qg_related_wood
qg_related
```


```{r}
# # Wang
# table(wang_related$ind1.id)
# table(wang_related$ind2.id)
# 
# all_ids_wang <- c(wang_related$ind1.id, wang_related$ind2.id)
# id_counts_wang <- as.data.frame(table(all_ids_wang))
# colnames(id_counts_wang) <- c("ID", "count")

# Queller & Goodnight
table(qg_related_wood$ind1.id)
table(qg_related_wood$ind2.id)

all_ids_qg_wood <- c(qg_related_wood$ind1.id, qg_related_wood$ind2.id)
id_counts_qg_wood <- as.data.frame(table(all_ids_qg_wood))
colnames(id_counts_qg_wood) <- c("ID", "count")
```


```{r}
## Queller & Goodnight
qg_related_wood$remove <- apply(qg_related_wood, 1, function(row) {
  id1 <- row["ind1.id"]
  id2 <- row["ind2.id"]
  
  c1 <- id_counts_qg_wood$count[id_counts_qg_wood$ID == id1]
  c2 <- id_counts_qg_wood$count[id_counts_qg_wood$ID == id2]
  
  if (c1 > c2) {
    return(id1)
  } else if (c2 > c1) {
    return(id2)
  } else {
    # Falls beide gleich oft vorkommen -> alphabetisch die kleinere ID nehmen
    return(min(id1, id2))
  }
})

# Alle rauszunehmenden IDs
ids_to_remove_qg_wood <- unique(qg_related_wood$remove)
#ids_to_remove <- order(ids_to_remove, decreasing = FALSE)
ids_to_remove_qg_wood
```
### Remove IDs
Who are the indivudals we have to remove?
```{r}
#gen_squirrels_135_IDs <- rownames_to_column(.data = gen_squirrels_135_IDs, var = "protocol_nr2") # Why???
ids_remove_qg_wood <- data.frame(protocol_nr_old = ids_to_remove_qg_wood) %>% 
  mutate(protocol_nr_old = as.numeric(protocol_nr_old))
ids_remove_wood <- left_join(x = ids_remove_qg_wood, y = gen_squirrels_135_IDs, by = "protocol_nr_old")
remove_wood <- data.frame(ID = as.numeric(ids_to_remove_qg_wood))
```

Order by study site, sex, season
```{r}
table(ids_remove_wood$Study_site, useNA = "always")
table(ids_remove_wood$Sex, useNA = "always")
table(ids_remove_wood$Season, useNA = "always")
```

calculate the percentage of individuals that has to be removed
```{r}
# noch anders anordnen
table(gen_squirrels_135_IDs$Study_site, gen_squirrels_135_IDs$Sex)
#table(gen_squirrels_93_IDs$Study_site, gen_squirrels_93_IDs$Sex)
table(ids_remove_wood$Study_site, ids_remove_wood$Sex)
```


```{r}
# all together
2/20 
# males
2/12
# females
0/8
```


## Cemetery
Load data and check for NAs
```{r}
cemetery_gdata <- kinship_cemetery$gdata#[-1, ]
# rows = individuals, columns = ID, 2 alleles of 16 loci
# table(squirrel_gdata$V1, useNA = "always") # allel 1 locus 1
# table(squirrel_gdata$V33, useNA = "always") # allel 2 locus 16
cemetery_gdata
wood_gdata
squirrel_gdata
```


Calculate relatedness
```{r}
#output <- coancestry(input$gdata, lynchrd = 2, quellergt = 2, wang = 2)
related_cemetery_output <- coancestry(genotype.data = kinship_cemetery$gdata, #[-1, ], 
                                      error.rates = 0.01, 
                                      #wang = 2, 
                                      quellergt = 2,
                                      ci95.num.bootstrap = 500, #1st version with 1000 repetitions but now 
                                      # the session aborts every time with 1000
                                      rng.seed = 20200604
)
```

Look at output
```{r}
relatedness_cemetery <- related_cemetery_output$relatedness
# summary(relatedness$wang)
summary(relatedness_cemetery$quellergt)
# table(relatedness$wang > 0.5)
table(relatedness_cemetery$quellergt > 0.5)
# boxplot(relatedness$wang)
boxplot(relatedness_cemetery$quellergt)

# wang_related <- relatedness %>% filter(wang >= 0.5)
qg_related_cemetery <- relatedness_cemetery %>% filter(quellergt > 0.5)
qg_related_cemetery
qg_related
```


```{r}
# # Wang
# table(wang_related$ind1.id)
# table(wang_related$ind2.id)
# 
# all_ids_wang <- c(wang_related$ind1.id, wang_related$ind2.id)
# id_counts_wang <- as.data.frame(table(all_ids_wang))
# colnames(id_counts_wang) <- c("ID", "count")

# Queller & Goodnight
table(qg_related_cemetery$ind1.id)
table(qg_related_cemetery$ind2.id)

all_ids_qg_cemetery <- c(qg_related_cemetery$ind1.id, qg_related_cemetery$ind2.id)
id_counts_qg_cemetery <- as.data.frame(table(all_ids_qg_cemetery))
colnames(id_counts_qg_cemetery) <- c("ID", "count")
```


```{r}
## Queller & Goodnight
qg_related_cemetery$remove <- apply(qg_related_cemetery, 1, function(row) {
  id1 <- row["ind1.id"]
  id2 <- row["ind2.id"]
  
  c1 <- id_counts_qg_cemetery$count[id_counts_qg_cemetery$ID == id1]
  c2 <- id_counts_qg_cemetery$count[id_counts_qg_cemetery$ID == id2]
  
  if (c1 > c2) {
    return(id1)
  } else if (c2 > c1) {
    return(id2)
  } else {
    # Falls beide gleich oft vorkommen -> alphabetisch die kleinere ID nehmen
    return(min(id1, id2))
  }
})

# Alle rauszunehmenden IDs
ids_to_remove_qg_cemetery <- unique(qg_related_cemetery$remove)
#ids_to_remove <- order(ids_to_remove, decreasing = FALSE)
ids_to_remove_qg_cemetery
```
### Remove IDs
Who are the indivudals we have to remove?
```{r}
#gen_squirrels_135_IDs <- rownames_to_column(.data = gen_squirrels_135_IDs, var = "protocol_nr2") # Why???
ids_remove_qg_cemetery <- data.frame(protocol_nr_old = ids_to_remove_qg_cemetery) %>% 
  mutate(protocol_nr_old = as.numeric(protocol_nr_old))
ids_remove_cemetery <- left_join(x = ids_remove_qg_cemetery, y = gen_squirrels_135_IDs, by = "protocol_nr_old")
remove_cemetery <- data.frame(ID = as.numeric(ids_to_remove_qg_cemetery))
```

Order by study site, sex, season
```{r}
table(ids_remove_cemetery$Study_site, useNA = "always")
table(ids_remove_cemetery$Sex, useNA = "always")
table(ids_remove_cemetery$Season, useNA = "always")
```

calculate the percentage of individuals that has to be removed
```{r}
# noch anders anordnen
table(gen_squirrels_135_IDs$Study_site, gen_squirrels_135_IDs$Sex)
#table(gen_squirrels_93_IDs$Study_site, gen_squirrels_93_IDs$Sex)
table(ids_remove_cemetery$Study_site, ids_remove_cemetery$Sex)
```


```{r}
# all together
12/50 
# males
7/25
# females
5/24
```

## Urban park
Load data and check for NAs
```{r}
park_gdata <- kinship_park$gdata#[-1,]
# rows = individuals, columns = ID, 2 alleles of 16 loci
# table(squirrel_gdata$V1, useNA = "always") # allel 1 locus 1
# table(squirrel_gdata$V33, useNA = "always") # allel 2 locus 16
park_gdata
squirrel_gdata
```


Calculate relatedness
```{r}
#output <- coancestry(input$gdata, lynchrd = 2, quellergt = 2, wang = 2)
related_park_output <- coancestry(genotype.data = kinship_park$gdata,#[-1, ], 
                             error.rates = 0.01, 
                             #wang = 2, 
                             quellergt = 2,
                             ci95.num.bootstrap = 500, #1st version with 1000 repetitions but now the session aborts every time with 1000
                             rng.seed = 20200604
                             )
```

Look at output
```{r}
relatedness_park <- related_park_output$relatedness
# summary(relatedness$wang)
summary(relatedness_park$quellergt)
# table(relatedness$wang > 0.5)
table(relatedness_park$quellergt > 0.5)
# boxplot(relatedness$wang)
boxplot(relatedness_park$quellergt)

# wang_related <- relatedness %>% filter(wang >= 0.5)
qg_related_park <- relatedness_park %>% filter(quellergt > 0.5)
qg_related_park
qg_related
```


```{r}
# # Wang
# table(wang_related$ind1.id)
# table(wang_related$ind2.id)
# 
# all_ids_wang <- c(wang_related$ind1.id, wang_related$ind2.id)
# id_counts_wang <- as.data.frame(table(all_ids_wang))
# colnames(id_counts_wang) <- c("ID", "count")

# Queller & Goodnight
table(qg_related_park$ind1.id)
table(qg_related_park$ind2.id)

all_ids_qg_park <- c(qg_related_park$ind1.id, qg_related_park$ind2.id)
id_counts_qg_park <- as.data.frame(table(all_ids_qg_park))
colnames(id_counts_qg_park) <- c("ID", "count")
```


```{r}
## Queller & Goodnight
qg_related_park$remove <- apply(qg_related_park, 1, function(row) {
  id1 <- row["ind1.id"]
  id2 <- row["ind2.id"]
  
  c1 <- id_counts_qg_park$count[id_counts_qg_park$ID == id1]
  c2 <- id_counts_qg_park$count[id_counts_qg_park$ID == id2]
  
  if (c1 > c2) {
    return(id1)
  } else if (c2 > c1) {
    return(id2)
  } else {
    # Falls beide gleich oft vorkommen -> alphabetisch die kleinere ID nehmen
    return(min(id1, id2))
  }
})

# Alle rauszunehmenden IDs
ids_to_remove_qg_park <- unique(qg_related_park$remove)
#ids_to_remove <- order(ids_to_remove, decreasing = FALSE)
ids_to_remove_qg_park
```
### Remove IDs
Who are the indivudals we have to remove?
```{r}
#gen_squirrels_135_IDs <- rownames_to_column(.data = gen_squirrels_135_IDs, var = "protocol_nr2") # Why???
ids_remove_qg_park <- data.frame(protocol_nr_old = ids_to_remove_qg_park) %>% 
  mutate(protocol_nr_old = as.numeric(protocol_nr_old))
ids_remove_park <- left_join(x = ids_remove_qg_park, y = gen_squirrels_135_IDs, by = "protocol_nr_old")
remove_park <- data.frame(ID = as.numeric(ids_to_remove_qg_park))
```

Order by study site, sex, season
```{r}
table(ids_remove_park$Study_site, useNA = "always")
table(ids_remove_park$Sex, useNA = "always")
table(ids_remove_park$Season, useNA = "always")
```

calculate the percentage of individuals that has to be removed
```{r}
# noch anders anordnen
table(gen_squirrels_135_IDs$Study_site, gen_squirrels_135_IDs$Sex)
#table(gen_squirrels_93_IDs$Study_site, gen_squirrels_93_IDs$Sex)
table(ids_remove_park$Study_site, ids_remove_park$Sex)
```


```{r}
# all together
21/65 
# males
12/30
# females
9/34
```

###All sites
```{r}
ids_remove_qg_wood
ids_remove_qg_cemetery
ids_remove_qg_park

2+12+21 # 35 IDs removed
135-35 # 100 Tiere verbleiben
```


## Simulations
```{r}
kinship_sim <- familysim(freqs = kinship_squirrels$freqs, ninds = 100)
output_sim <- coancestry(kinship_sim, quellergt = 2) #wang = 2
simrel <- cleanuprvals(output_sim$relatedness, 100)
relvalues <- simrel[, 6]

head(output_sim$relatedness)
head(simrel)
#relvalues
```


```{r}
label1 <- rep("PO", 100)
label2 <- rep("Full", 100)
label3 <- rep("Half", 100)
label4 <- rep("Unrelated", 100)
labels <- c( label1 , label2 , label3 , label4 )
Relationship <- labels
simdata <- as.data.frame(cbind(Relationship, relvalues))
simdata$relvalues <- as.numeric(as.character(simdata$relvalues))
```

Boxplot and density plot
```{r}
qplot(as.factor(labels), relvalues, geom = "boxplot", ylab = "Relatedness Values", xlab = "Relatedness")
qplot(relvalues, ..density.., data = simdata, geom = "density", colour = as.factor(Relationship), xlab = "Relatedness value", ylab = "Density") + 
  theme(legend.position = "top") 
```

## Comparison between study sites
```{r}
# grouprel ( genotypes = reldata$gdata , estimatorname = " wang ", usedgroups = " all ",
# iterations = 100)

grouprel(genotypes = kinship_site_squirrels$gdata, estimatorname = "wang", usedgroups = "all",
         iterations = 100)
```

# Next steps
Remove the 42 individuals from the raw data that, according to this analysis, show a high degree of relatedness to other animals.

This data set is used for analysis with STRUCTURE and all further analysis.


***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>
