---
title: "Squirrel genetics: Descriptive genetic per population" ## name of your project and analysis step
description: |
    The aim of this study is...
author:
    ## author information; single arguemnts can be deleted if not needed
    - name: "John Smith"
      url: https://izw-berlin.de/en/john-smith-en.html  
      affiliation: Leibniz Institute for Zoo and Wildlife Research
      affiliation_url: https://izw-berlin.de/en/
      orcid_id: 0000-0000-0000-0000
    - name: "Jane Doe"
      url: https://janedoe.com
      affiliation: Leibniz Institute for Zoo and Wildlife Research
      affiliation_url: https://izw-berlin.de/en/
    ## add more authors with the same intendation and styling if needed 
    ## a new author always starts with a hyphen!
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
        highlight: kate      ## styling of code
        code_folding: false  ## if `true` you can expand and shrink code chunks
        toc: true            ## if `true` adds a table of content
        toc_depth: 2         ## level to be displayed in the table of content
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      dev = "ragg_png", fig.width = 9, fig.height = 6, dpi = 600, retina = 1)
Sys.setlocale("LC_TIME", "C")
```

* **Research question:**
* **Study area:**
* **Data:** 


# Setup

```{r packages}
## for non-CRAN packages please keep install instruction
## but commented so it is not run each time, e.g.
# devtools::install_github("EcoDynIZW/template")

## libraries used in this script
## please add ALL LIBRARIES NEEDED HERE
## please remove libraries from the list that are not needed anymore 
## at a later stage
#library()
d6::simple_load(c("adegenet", 
                  "FSA", 
                  "hierfstat",
                 # "kkeenan02/diveRsity", 
                  "tidyverse", 
                  #"ggplot2", 
                  #"gridExtra", 
                  #"here", 
                  "pegas",
                 "inbreedR", 
                 # "royfrancis/pophelper",
                  "poppr",
                 "related",
                 "rstatix", 
                 "writexl"
                  #"ericarcher/strataG"
                  #"scatterpie"
))
# For the package related: please follow the instructions here: https://github.com/timothyfrasier/related 
```

```{r}
set.seed(20200604)
```


## Plot setup
Set theme for ggplots
```{r theme}
theme_set(
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 2),
    plot.margin = margin(6, 10, 4, 4),
    axis.ticks = element_line(colour = "black", linewidth = 1.5),
    axis.ticks.length = unit(0.2, "cm"),
    title = element_text(size = 20), # axis family = "Zilla Slab", family = "Roboto", 
    axis.text = element_text(size = rel(1.4)))) # axis ticks, family = "Roboto", 
```

Labeller for study sites
```{r}
# Remove the underscore in the name of the study site without changing the factor levels
Study_site_names <- as_labeller(c("Rahnsdorf" = "woodland area", "Treptower_Park" = "urban park", "Lilienthal" = "cemetery"))
```


E:/PopDynCloud/Projects/sciurus_vulgaris_b_cmr_drenske_s/plots/allel_richness.png

# Data

```{r data}
# Individual genetic data
# squirrel_genind <- read.genepop(file = here::here("data-raw", "genetic-raw", "Cervus",  "Squirrel_Genotypes_136_IDs_20250718-GENEPOP.gen"), ncode = 3, quiet = TRUE)
squirrel_genind <- read.genepop(file = here::here("data-raw", "genetic-raw",
                                                  "Squirrel_Genotypes_135_IDs_20250816-GENEPOP.gen"), 
                                ncode = 3, 
                                quiet = TRUE)


# Genetic and CMR data
gen_squirrels_135_IDs <- readRDS(file = here::here("output", "data-proc", "genetic-proc", "cluster_analysis",  
                                                   "gen_squirrels_135_IDs_20250719.RDS"))
gen_squirrels_135_IDs <- gen_squirrels_135_IDs %>%
  mutate(Study_site = factor(Study_site, levels = c("Rahnsdorf", "Lilienthal", "Treptower_Park")))

gen_squirrels_93_IDs <- readRDS(file = here::here("output", "data-proc", "genetic-proc", "cluster_analysis",  
                                                   "gen_squirrels_93_IDs_20250823.RDS"))

gen_squirrels_93_IDs <- gen_squirrels_93_IDs %>%
  mutate(Study_site = factor(Study_site, levels = c("Rahnsdorf", "Lilienthal", "Treptower_Park")))
```


## Add Strata
```{r}
# Add strata (grouping variables, individual characteristics) to the genind object
strata(squirrel_genind) <- gen_squirrels_135_IDs
```

## Prepare objects for genetic analyses
```{r}
# GENIND
# genind object with study sites as the populations
squirrel_genind_sites <- squirrel_genind
setPop(squirrel_genind_sites) <- ~Study_site 

# genind object with sex & study sites as the populations
squirrel_genind_sites_sex <- squirrel_genind
setPop(squirrel_genind_sites_sex) <- ~Study_site/Sex 

# Separate the genotypes per population (needed for some analyses)
study_sites_pop <- seppop(squirrel_genind_sites)
# And per study site (needed for inbreeding)
Rahn_pop <- seppop(squirrel_genind_sites)$Rahnsdorf
Lilie_pop <- seppop(squirrel_genind_sites)$Lilienthal
Treptow_pop <- seppop(squirrel_genind_sites)$Treptower_Park

# genpop object: coverts genotypes data into alleles counts per population
squirrel_pop <- genind2genpop(squirrel_genind_sites)
squirrel_pop

# HIERFSTAT
# hierfstat object all individuals together
squirrel_all_hierf <- genind2hierfstat(squirrel_genind)

# hierfstat object study sites
squirrel_sites_hierf <- genind2hierfstat(squirrel_genind_sites)
```

Transform genind objects to datasets for plotting
```{r}
squirrel_genind_df <- genind2df(squirrel_genind)
squirrel_genind_sites_df <- genind2df(squirrel_genind_sites)
levels(squirrel_genind_sites_df$pop)
squirrel_genind_sites_df <- squirrel_genind_sites_df %>%
  mutate(pop = fct_relevel(pop, "Rahnsdorf", "Lilienthal", "Treptower_Park"))
levels(squirrel_genind_sites_df$pop)
```

## Save data
```{r}
# # squirrel_genind_df
# squirrel_genind_sites_df <- squirrel_genind_sites_df %>%
#   rownames_to_column() %>%
#   rename(Squirrel_ID = rowname)
# vroom::vroom_write(x = squirrel_genind_sites_df,
#                    file = here::here("output", "data-proc", "genetic-proc", "squirrel_genind_sites_df_20251108.csv"))
# saveRDS(object = squirrel_genind_sites_df,
#         file = here::here("output", "data-proc", "genetic-proc", "squirrel_genind_sites_df_20251108.RDS"))
# write_xlsx(x = squirrel_genind_sites_df, path = here::here("output", "data-proc", "genetic-proc", "squirrel_genind_sites_df_20251108.xlsx"))
```


# Data exploration
## Genind objects
```{r}
squirrel_genind
```


```{r}
squirrel_genind_sites
```


```{r}
study_sites_pop
```


```{r}
popNames(squirrel_genind) # 506 - the ID of the last individual
nAll(squirrel_genind) # returns the number of alleles for each locus.
nPop(squirrel_genind) # returns the number of populations.
#tab(squirrel_genind)
#indNames(squirrel_genind)
locNames(squirrel_genind)
alleles(squirrel_genind)
ploidy(squirrel_genind) # returns/sets ploidy of the individuals (Anzahl vollständiger Chromosomensätze)
```


```{r}
popNames(squirrel_genind_sites) # 3 Populations, our study sites
nAll(squirrel_genind_sites) # returns the number of alleles for each locus.
nPop(squirrel_genind_sites) # returns the number of populations.
#tab(squirrel_genind_sites)
#indNames(squirrel_genind_sites)
#locNames(squirrel_genind_sites)
#alleles(squirrel_genind_sites)
#ploidy(squirrel_genind_sites) # r
```

```{r}
popNames(squirrel_pop)
nAll(squirrel_pop)
```


## Hierfstat objects
Quick look at the hierfstat objects
```{r}
head(squirrel_all_hierf) # pop = 1 = all together
head(squirrel_sites_hierf) # pop = Study sites
```


## Basic summary of the data
```{r}
# all together
gen_sum <- summary(squirrel_genind)
gen_sum
names(gen_sum)

# per study site
gen_sum_sites <- summary(squirrel_genind_sites)
gen_sum_sites

# per population (not really helpful)
gen_sum_pop <- summary(squirrel_pop)
gen_sum_pop
```

```{r}
table(gen_squirrels_135_IDs$Study_site)
table(gen_squirrels_135_IDs$Study_site, gen_squirrels_135_IDs$Sex)
table(gen_squirrels_93_IDs$Study_site, gen_squirrels_93_IDs$Sex)
# in total 135
```


```{r}
# ggplot(data = squirrel_genind_sites_df, mapping = aes(x = pop, fill = pop)) + 
#   geom_bar() +
#   scale_fill_manual(values = c("Treptower_Park" = "tomato3", 
#                                "Lilienthal" = "orange2", 
#                                "Rahnsdorf" = "palegreen4")) 

ggplot(data = gen_squirrels_135_IDs, mapping = aes(x = Study_site, fill = Study_site)) + 
  geom_bar() +
  scale_x_discrete(expand = c(0.001, 0.001), labels = Study_site_names) + 
  scale_y_continuous(
    limits = c(0, 75),
    breaks = seq(0, 75, by = 25),
    expand = c(.004, .001)) +
  scale_fill_manual(values = c("Treptower_Park" = "#8B2F1D", 
                               "Lilienthal" = "#F0A202", 
                               "Rahnsdorf" = "#006D6F")) +
  labs(x = "Study site", y = "Nr of individuals") 

# ggsave(filename = here::here("plots","genetics", "samples_study_site_135IDS_20250906.png"),
#        # width = 1650*1.25,
#        # height = 1200,
#        # units = "px",
#        dpi = 300)

ggplot(data = gen_squirrels_93_IDs, mapping = aes(x = Study_site, fill = Study_site)) + 
  geom_bar() +
  scale_x_discrete(expand = c(0.001, 0.001), labels = Study_site_names) + 
  scale_y_continuous(
    limits = c(0, 75),
    breaks = seq(0, 75, by = 25),
    expand = c(.004, .001)) +
  scale_fill_manual(values = c("Treptower_Park" = "#8B2F1D", 
                               "Lilienthal" = "#F0A202", 
                               "Rahnsdorf" = "#006D6F")) +
  labs(x = "Study site", y = "Nr of individuals") 

# ggsave(filename = here::here("plots","genetics", "samples_study_site_93IDS_20250906.png"),
#        # width = 1650*1.25,
#        # height = 1200,
#        # units = "px",
#        dpi = 300)
```




# Number of alleles per locus (NA)
For all individuals and all loci
```{r}
nr_all <- nAll(squirrel_genind_sites)
nr_all
```
Per study site:
Here we have to calculate the count of copies per allele
```{r}
AC <- allele.count(squirrel_sites_hierf)
#AC
AC$Rsu3
```
and then we calculate the number of different alleles per study site
```{r}
# Create a clean tibble with the number of alleles per population and locus
allele_counts_tbl <- lapply(names(AC), function(locus) {
  
  mat <- AC[[locus]]                  # Extract the allele count matrix for this locus
  n_alleles_pop <- apply(             # Count alleles per population:
    mat,                              # For each population (column), count how many alleles
    2,                                # appear at least once (value > 0)
    function(col) sum(col > 0)
  )
  
  # Build a temporary data.frame for this locus
  out <- data.frame(
    locus = locus,                    # Locus name
    population = names(n_alleles_pop),# Population names
    n_alleles = n_alleles_pop,        # Number of alleles observed in that population
    row.names = NULL                  # Remove row names here already
  )
  
  out
}) |> 
  bind_rows() |>                      # Combine all loci into one table
  mutate(row_id = row_number()) |>    # Reset and generate clean sequential row numbers
  relocate(row_id)                    # Move row_id to the first column

allele_counts_tbl
```

Sum, Mean, SD number of alleles per population
```{r}
pop_mean_N_all <- allele_counts_tbl %>%
  group_by(population) %>%
  summarize(sum_all_counts = sum(n_alleles),
            mean_all_counts = mean(n_alleles),
            sd_all_counts = sd(n_alleles))
pop_mean_N_all

# # Test
# TP_all <- allele_counts_tbl %>% filter(population == "Treptower_Park")
# mean(TP_all$n_alleles)
# sum(TP_all$n_alleles)
# Lilie_all <- allele_counts_tbl %>% filter(population == "Lilienthal")
# mean(Lilie_all$n_alleles)
# sum(Lilie_all$n_alleles)
# Rahn_all <- allele_counts_tbl %>% filter(population == "Rahnsdorf")
# mean(Rahn_all$n_alleles)
# sum(Rahn_all$n_alleles)
```
## test for significant differences
Kruskal Wallis Test + Wilcoxon 
```{r}
kruskal.test(n_alleles ~ as.factor(population), data = allele_counts_tbl)

pairwise.wilcox.test(
  x = allele_counts_tbl$n_alleles,
  g = allele_counts_tbl$population,
  p.adjust.method = "bonferroni"
)
```

# Allele frequency
Frequency of each allel in a population

## All together
```{r}
allele_freqs <- tab(squirrel_genind, freq = TRUE, NA.method = "asis")
allele_freqs
```


```{r}
allele_freq_summary <- apply(allele_freqs, 2, mean, na.rm = TRUE)
allele_freq_summary # sums up to 1 per locus
```


```{r}
# # overall number of alleles per individual
#allele_counts <- tab(squirrel_genind, freq = FALSE)
# allele_counts

# relative allele frequencies per Locus per individual 
allele_freq <- as.data.frame(tab(squirrel_genind, freq = TRUE))
allele_freq
```


```{r}
#par(mfrow=c(2,2))
#plot(gen_sum$n.by.pop, gen_sum$pop.n.all, xlab="Colonies sample size",
#ylab="Number of alleles",main="Alleles numbers and sample sizes",
#type="n")
#text(gen_sum$n.by.pop,gen_sum$pop.n.all,lab=names(gen_sum$n.by.pop))
barplot(gen_sum$loc.n.all, ylab = "Number of alleles",
main = "Number of alleles per locus")

```

## Per study site
short way
```{r}
#stats_sites$pop.freq
```

Calculate allele frequencies for each population
```{r}
allele_freq_per_pop <- lapply(study_sites_pop, function(pop) {
  as.data.frame(tab(pop, freq = TRUE))  # Frequencies per individual
})
allele_freq_per_pop
```

 Aggregate to get mean allele frequency per locus per population (sum over individuals / n_individuals)
```{r}
allele_freq_summary <- lapply(allele_freq_per_pop, function(df) {
  colMeans(df, na.rm = TRUE)
})
allele_freq_summary
```

 Combine into data.frame
```{r}
allele_freq_summary_df <- as.data.frame(do.call(rbind, allele_freq_summary))
allele_freq_summary_df
```

Add population names as column and reset rownames
```{r}
allele_freq_summary_df$population <- names(allele_freq_summary)
#allele_freq_summary_df
allele_freq_summary_df <- allele_freq_summary_df %>% 
  dplyr::select(population, dplyr::everything()) # population as first column
#allele_freq_summary_df
rownames(allele_freq_summary_df) <- NULL
allele_freq_summary_df
```

Round values and look at results
```{r}
allele_freq_summary_df <- allele_freq_summary_df %>% mutate(across(where(is.numeric), ~ round(., 3)))

allele_freq_summary_df
```

```{r}
allele_freq_long <- allele_freq_summary_df %>%
  pivot_longer(cols = -population,
               names_to = "allele",
               values_to = "frequency")
allele_freq_long
```

```{r}
ggplot(allele_freq_long,
       aes(x = allele, y = frequency, fill = population)) +
  geom_col(position = "dodge") +
  labs(
    x = "Allele",
    y = "Frequency",
    fill = "Population",
    title = "Allele Frequencies per Population"
  ) +
  theme_minimal() + 
  facet_wrap(~population, nrow = 3)

ggplot(allele_freq_long,
       aes(x = allele, y = population, fill = frequency)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(
    x = "Allele",
    y = "Population",
    fill = "Frequency",
    title = "Heatmap of Allele Frequencies"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```


# Allelic richness (Ar)
Number of different alleles per locus in a population. Considers only the number of existing alleles, regardless of their frequency. High allelic richness indicates extensive genetic diversity and is important for a populations ability to adapt to environmental changes

## All together
```{r}
# calculate Allelic Richness 
allelic_richness_all <- allelic.richness(squirrel_all_hierf, diploid = TRUE)
#allelic_richness_all

# look at the output
allelic_richness_all$Ar  # Matrix mit Richness-Werten pro Locus und Population
#allelic_richness_all$Pop
```
```{r}
#boxplot(t(allelic_richness_all$Ar)) 
barplot(t(allelic_richness_all$Ar))
```

## Study sites
```{r}
# calculate Allelic Richness 
allelic_richness_sites <- allelic.richness(squirrel_sites_hierf)
allelic_richness_sites$min.all # number of alleles used for rarefaction

# look at the output
allelic_richness_all$Ar
allelic_richness_sites$Ar  # Matrix mit Richness-Werten pro Locus und Population

# Mean allelic richness per population (does that calculation even make sense?)
round(apply(allelic_richness_sites$Ar, 2, mean, na.rm = TRUE), digits = 2)
round(apply(allelic_richness_sites$Ar, 2, sd, na.rm = TRUE), digits = 2)
```

```{r}
barplot(t(allelic_richness_all$Ar)) 
barplot(t(allelic_richness_sites$Ar), 
        beside = TRUE, 
        legend.text = TRUE,
        args.legend = list(x = "topleft"))
```
Transform data frame
```{r}
allelic_richness_tbl <- allelic_richness_sites$Ar %>%
  as.data.frame() %>%                     # Convert matrix to data.frame
  rownames_to_column(var = "locus") %>%   # Add loci as a column
  pivot_longer(
    cols = -locus,                        # Everything except 'locus' column
    names_to = "population",              # Column names become population
    values_to = "Ar"  )                    # Cell values become Ar
  # ) %>%
  # mutate(row_id = row_number()) %>%       # Add sequential row numbers
  # relocate(row_id)                         # Move row_id to the first column

# View result
head(allelic_richness_tbl)
```

Kruskal Wallis Test
```{r}
kruskal.test(Ar ~ as.factor(population), data = allelic_richness_tbl)
pairwise.wilcox.test(
  x = allelic_richness_tbl$Ar,
  g = allelic_richness_tbl$population,
  p.adjust.method = "bonferroni"
)
```




# Private alleles
A private allele is a genetic variation found exclusively in a single population and not in others. 

How to calculate the number of private alleles:
Tabulate alleles that occur in only one population.
```{r}
pa <- private_alleles(squirrel_genind_sites)
pa
```

Other way of visualisation, now with one row per private allel
```{r}
priv_allel <- as.data.frame(t(private_alleles(squirrel_genind_sites)))
priv_allel <- rownames_to_column(priv_allel, var = "Locus.Allel")
priv_allel <- priv_allel %>% 
  separate_wider_delim(cols = Locus.Allel, names = c("Locus", "Allel"), delim = ".", cols_remove = FALSE)

priv_allel


colSums(priv_allel != 0) # count the nr of rows with another value than 0
```

Rahnsdorf has 15 private alleles
Lilienthal 6
Treptow 12

# Basic statistics per loci and populations
## All together
```{r}
stats_all <- basic.stats(squirrel_all_hierf)
stats_all
# Ho = observed heterozygosity
# Hs = within population gene diversity (sometimes misleadingly called expected heterozygosity)
# perloc = basic statistics per loci
# overall = basic statistics averaged over loci

# Basic statistics
# Ho, Hs see above
# Ht = overall gene diversity, total heterozygosity
# Dst = amount of gene diversity (Ht - Hs)
# Htp = (maybe expected heterozygosity when sample sizes are even)
# Dstp = Proportional differentiation Dst/Ht
# Fst = Inbreeding between populations
# Fis = 1 - Ho/Hs = Inbreeding within a population
# Dest = Jost's D

```

## Study sites
```{r}
stats_sites <- basic.stats(squirrel_sites_hierf)
stats_sites
# Ho = observed heterozygosity
# Hs = within population gene diversity (sometimes misleadingly called expected heterozygosity)
# perloc = basic statistics per loci
# overall = basic statistics averaged over loci

# Basic statistics
# Ho, Hs see above
# Ht = overall gene diversity, total heterozygosity
# Dst = amount of gene diversity (Ht - Hs)
# Htp = (maybe expected heterozygosity when sample sizes are even)
# Dstp = Proportional differentiation Dst/Ht
# Fst = Inbreeding between populations
# Fis = 1 - Ho/Hs = Inbreeding within a population
# Dest = Jost's D
```

```{r}
stats_sites$Ho
colMeans(stats_sites$Ho, na.rm = TRUE)
```

```{r}
stats_sites$pop.freq
```


# Heterozygosity (Ho + He, distribution of indidivuals per Loci)
Heterozygosity refers to the presence of different alleles at a given locus in an individual's genome. In a population genetics context, it is a measure of genetic variation and represents the proportion of individuals in a population that are heterozygous (having two different alleles, e.g., AaAaAa) at a specific locus.
Heterozygosity is a key indicator of genetic diversity, which is important for a population's adaptability and long-term survival.

H_obs = H_exp = HWE, no inbreeding, no selection
H_obs < H_exp = Inbreeding, selection
H_obs > H_exp = hybridisation, mix of different groups

Observed Heterozygosity: H_obs
The percentage of loci in a population that is heterozygous is known as observed heterozygosity (HO), and it is determined for each locus by dividing the total number of heterozygotes by the sample size.

Expected Heterozygosity: H_exp
This is the proportion of heterozygotes expected in a population under Hardy-Weinberg Equilibrium (HWE), based on the allele frequencies at the locus. 
## All together
```{r}
#ChatGPT
# bs <- hierfstat::basic.stats(squirrel_genind)
# 
# He <- bs$Hs      # expected heterozygosity per locus
# Ho <- bs$Ho      # observed heterozygosity per locus
# Ae <- 1 / (1 - He)  # effective number of alleles


Hobs_all <- stats_all$Ho %>%
  as.data.frame() %>%
  rename(Hobs = 1) %>%
  select(Hobs)

Hexp_all <- stats_all$Hs %>%
  as.data.frame() %>%
  rename(Hexp = 1) %>%
  select(Hexp)

Hobs_all
Hexp_all
```


Is mean observed H significantly lower than mean expected H ? Hint about inbreeding
```{r}
mean(Hobs_all$Hobs)
mean(Hexp_all$Hexp)


t.test(x = Hobs_all$Hobs, 
       y = Hexp_all$Hexp, 
       pair = T, 
       var.equal = TRUE, 
       alternative = "less" 
       )

# Mean values are significantly different

```
## Study sites
```{r}
hobs_site <- as.data.frame(stats_sites$Ho)
hobs_site

hexp_site <- as.data.frame(stats_sites$Hs)
hexp_site

hobs_site_mean <- colMeans(hobs_site, na.rm = TRUE)
hexp_site_mean <- colMeans(hexp_site, na.rm = TRUE)
hobs_site_mean
hexp_site_mean

# Rahnsdorf Hobs = Hexp
# Lilie Hobs < Hexp
# Treptow Hobs < Hexp
```

```{r}
het_summary <- data.frame(
  population = colnames(hobs_site),
  Ho_mean = round(colMeans(hobs_site, na.rm = TRUE), 2),
  Ho_mean2 = apply(hobs_site, 2, mean, na.rm = TRUE),
  Ho_sd = round(apply(hobs_site, 2, sd, na.rm = TRUE), 2),
  Hs_mean = round(colMeans(hexp_site, na.rm = TRUE), 2),
  Hs_mean2 = apply(hexp_site, 2, mean, na.rm = TRUE),
  Hs_sd = round(apply(hexp_site, 2, sd, na.rm = TRUE), 2)
)

het_summary
```

```{r}
t_test_rahnsdorf <- t.test(
  x = hobs_site$Rahnsdorf,      # observed heterozygosity per locus
  y = hexp_site$Rahnsdorf,      # expected heterozygosity per locus
  paired = TRUE,             # paired because each locus is a pair
  alternative = "two.sided" # test if Ho is different from Hs (default)
)

t_test_rahnsdorf
# 15 df = 16 Loci -1
# p = 0.28 = not significant
# mean difference = -0.033 Ho is on average 3.3% lower than He

# t.test(x = hobs_site$Rahnsdorf, 
#        y = hexp_site$Rahnsdorf, 
#        paired = TRUE, 
#        alternative = "two.sided" 
#        )
# 
# # NS
# 
# t.test(x = hobs_site$Lilienthal, 
#        y = hexp_site$Lilienthal, 
#        paired = TRUE, 
#        alternative = "two.sided" 
#        )
# 
# # p 0.02 = Hint about inbreeding
# 
# t.test(x = hobs_site$Treptower_Park, 
#        y = hexp_site$Treptower_Park, 
#        paired = TRUE, 
#        alternative = "two.sided" 
#        )
# 
# # p 0.02 = Hint about inbreeding
```

Observed heterozygosity (Ho) did not significantly differ from expected heterozygosity (Hs) in Rahnsdorf (paired t-test, t = -1.12, df = 15, p = 0.28), indicating no deviation from Hardy-Weinberg expectations.

```{r}
# Get population names
populations <- colnames(hobs_site)

# Run paired t-test for each population and store results in a list
t_test_results <- lapply(populations, function(pop) {
  test <- t.test(
    hobs_site[[pop]],   # observed heterozygosity
    hexp_site[[pop]],   # expected heterozygosity
    paired = TRUE,
    alternative = "two.sided"
  )
  
  # Return a small dataframe with key results
  data.frame(
    population = pop,
    t_value = test$statistic,
    df = test$parameter,
    p_value = test$p.value,
    mean_diff = test$estimate
  )
})

# Combine list into one table
t_test_summary <- bind_rows(t_test_results)

# Round numbers for readability
t_test_summary <- t_test_summary %>%
  mutate(
    t_value = round(t_value, 2),
    df = round(df, 0),
    p_value = round(p_value, 3),
    mean_diff = round(mean_diff, 3)
  )

t_test_summary

```

Observed heterozygosity (Ho) varied among populations. In Rahnsdorf, Ho did not differ significantly from expected heterozygosity (Hs) (paired t-test, t = -1.12, df = 15, p = 0.28), indicating no deviation from Hardy-Weinberg expectations. In Treptower Park, Ho tended to be slightly lower than Hs, but this difference was not statistically significant (t = -1.97, df = 15, p = 0.067). In Lilienthal, however, Ho was significantly lower than Hs (t = -2.20, df = 15, p = 0.044), suggesting reduced heterozygosity possibly due to inbreeding or genetic isolation.

### comparison between study sites
```{r}
hobs_long <- hobs_site %>%
  mutate(Locus = rownames(hobs_site)) %>%
  pivot_longer(cols = -Locus, names_to = "Study_site", values_to = "Ho")

# He long
hexp_long <- hexp_site %>%
  mutate(Locus = rownames(hexp_site)) %>%
  pivot_longer(cols = -Locus, names_to = "Study_site", values_to = "He")
```

Kruskal Wallis für H obs
```{r}
kruskal.test(Ho ~ Study_site, data = hobs_long)
pairwise.wilcox.test(hobs_long$Ho, hobs_long$Study_site, p.adjust.method = "bonferroni")
```
No significant differences.

Kruskal Wallis für H exp
```{r}
kruskal.test(He ~ Study_site, data = hexp_long)
pairwise.wilcox.test(hexp_long$He, hexp_long$Study_site, p.adjust.method = "bonferroni")
```

Observed and expected heterozygosity did not differ significantly among the three study sites. Mean observed heterozygosity (Ho) was 0.60 in Rahnsdorf, 0.52 in Treptower Park, and 0.53 in Lilienthal (Kruskal-Wallis test: χ² = 1.06, df = 2, p = 0.59). Mean expected heterozygosity (He) was 0.64, 0.56, and 0.53, respectively (χ² = 0.54, df = 2, p = 0.76), indicating comparable levels of genetic diversity across populations.

## Multilocus heterozygosity (MLH)
```{r}
df <- squirrel_genind_df

# ID from rows
df$ID <- seq_len(nrow(df))

# locus names
loci <- setdiff(names(df), c("pop", "ID"))

# 2 columns per loci, 1 for each allel
split_locus <- function(x) {
  data.frame(allele1 = substr(x, 1, 3),
             allele2 = substr(x, 4, 6))
}

# use for all loci
allele_df_list <- lapply(loci, function(locus) {
  tmp <- split_locus(df[[locus]])
  colnames(tmp) <- paste0(locus, c("_1", "_2"))
  tmp
})

# create dataframe
genotypes_split <- bind_cols(df["ID"], df["pop"], do.call(bind_cols, allele_df_list))
genotypes_split <- genotypes_split %>%
  select(-ID, -pop)

# show result
str(genotypes_split)
genotypes_split
```
Convert the datafram into the desired format for inbreedR
```{r}
genotypes_squirrels <- convert_raw(genotypes_split)
check_data(genotypes_squirrels)

```
Calculate multilocus heterozygosity
```{r}
MLH_sq <- sMLH(genotypes_squirrels)
MLH_sq

MLH_sq <- as.data.frame(MLH_sq) %>%
  rownames_to_column(var = "ID")

# MLH_site <- MLH_sq %>%
#   left_join()
```

```{r}
summary(MLH_sq$MLH_sq)
```


## Counts of homo- and heterozygosity
How many individuals are homo- or heterozygous per Locus?

Transform rownames to column
```{r}
squirrel_sites_ms <- squirrel_sites_hierf %>% 
  rownames_to_column(var = "ID") #%>% 
  #mutate(ID = as.numeric(ID))
```

Put the table into a long format (transform)
```{r}
squirrel_geno_long <- squirrel_sites_ms %>%
  pivot_longer(
    cols = -c(ID, pop), # everything except ID and population (= Study site) are microsatellites
    names_to = "Locus", 
    values_to = "Alleles"
  )
squirrel_geno_long
```

Split Loci in 2 alleles, each with 3 digits
```{r}
squirrel_geno_long <- squirrel_geno_long %>%
  mutate(
    Allele1 = str_sub(Alleles, 1, 3),
    Allele2 = str_sub(Alleles, 4, 6)
  )
squirrel_geno_long
```

Determination of homozygosity and heterozygosity
```{r}
squirrel_geno_long <- squirrel_geno_long %>%
  mutate(HomoHetero = if_else(Allele1 == Allele2, "homozygous", "heterozygous"))
squirrel_geno_long
```
Look at the results
```{r}
squirrel_geno_counts <- squirrel_geno_long %>%
  count(Locus, HomoHetero) %>% drop_na()

squirrel_geno_site <- squirrel_geno_long %>%
  count(pop, Locus, HomoHetero) %>% drop_na() 

squirrel_geno_site$pop <- fct_relevel(squirrel_geno_site$pop, "Rahnsdorf", "Lilienthal", "Treptower_Park")
fct_relevel(squirrel_geno_site$pop)
```
```{r}
squirrel_geno_counts
squirrel_geno_site
```
```{r}
hetero_prop <- squirrel_geno_site %>%
  group_by(pop, Locus) %>%                    # pro Population und Locus
  mutate(total_individuals = sum(n),          # Gesamtzahl der untersuchten Individuen
         prop_hetero = n / total_individuals) #%>%  # proportion heterozygous
  #filter(HomoHetero == "heterozygous") %>%   # nur Heterozygote behalten
  #select(pop, Locus, prop_hetero)
```

```{r}
hetero_prop %>%
  group_by(pop) %>% 
  filter(HomoHetero == "heterozygous") %>% # nur Heterozygote behalten
  select(pop, Locus, prop_hetero) %>%
  summarise(mean_prop_hetero = mean(prop_hetero),
            sd_prop_hetero = sd(prop_hetero))
```


Plot it
```{r}
# # All together
# ggplot(squirrel_geno_counts, aes(x = Locus, y = n, fill = HomoHetero)) +
#   geom_col(position = "dodge") +
#   scale_y_continuous(
#     limits = c(0, 110),
#     breaks = seq(0, 110, by = 55),
#     expand = c(.004, .001)) +
#   labs(
#     x = "Locus",
#     y = "N individuals",
#     fill = ""
#   ) +
#   # theme_minimal(base_size = 13) +
#   theme(
#     axis.text.x = element_text(angle = 45, hjust = 1),
#      legend.position = "top"
#    )

# Per study site
ggplot(squirrel_geno_site, aes(x = Locus, y = n, fill = HomoHetero)) +
  geom_col(position = "dodge") +
  scale_y_continuous(
    limits = c(0, 60),
    breaks = seq(0, 60, by = 30),
    expand = c(.004, .001)) +
  facet_wrap(~pop, nrow = 3) +
  labs(
    x = "Locus",
    y = "N individuals",
    fill = ""
  ) +
  #theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top",
    strip.background = element_rect(fill = "grey95", color = NA)
  )

# Proportion
ggplot(hetero_prop, aes(x = Locus, y = prop_hetero, fill = HomoHetero)) +
  geom_col(position = "dodge") +
  # scale_y_continuous(
  #   limits = c(0, 60),
  #   breaks = seq(0, 60, by = 30),
  #   expand = c(.004, .001)) +
  facet_wrap(~pop, nrow = 3) +
  labs(
    x = "Locus",
    y = "N individuals",
    fill = ""
  ) +
  #theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top",
    strip.background = element_rect(fill = "grey95", color = NA)
  )
```


# Effective number of alleles (Ae)
```{r}
#Ae <- 1 / (1 - He) #(oder Hexp_all)

Ae_sites <- 1 / (1 - hexp_site)

# Optional: Mittelwert und SD pro Population
Ae_mean <- colMeans(Ae_sites, na.rm = TRUE)
Ae_sd <- apply(Ae_sites, 2, sd, na.rm = TRUE)

Ae_summary <- data.frame(
  population = colnames(Ae_sites),
  Ae_mean = round(Ae_mean, 2),
  Ae_sd = round(Ae_sd, 2)
)

Ae_sites
Ae_summary
```

Summary of effective number of alleles (Ae) per population
Rahnsdorf: moderate to high genetic diversity, some loci highly variable (Ae_mean = 3.33, sd = 1.77)
Treptower_Park: similar mean Ae (3.09), but higher variation among loci (sd = 2.06)
Lilienthal: lowest diversity (Ae_mean = 2.87), relatively uniform across loci (sd = 1.13)
Interpretation: Lilienthal might be more genetically isolated or show inbreeding, while Rahnsdorf and Treptower_Park maintain higher genetic diversity.

The effective number of alleles (Ae) differed slightly among populations. Rahnsdorf exhibited the highest mean Ae per locus (Ae = 3.33 ± 1.77), indicating relatively high genetic diversity with considerable variation among loci. Treptower Park had a similar mean Ae (Ae = 3.09 ± 2.06) but with higher locus-to-locus variation. Lilienthal showed the lowest mean Ae (Ae = 2.87 ± 1.13), reflecting lower overall genetic diversity that was relatively consistent across loci, suggesting possible genetic isolation or inbreeding.

## test for significant differences
change format of the table
```{r}
Ae_pivot <- Ae_sites %>%
  rownames_to_column(var = "Locus") %>%
  pivot_longer(cols = -Locus, names_to = "Study_site", values_to = "Ae")

# Ae_pivot <- Ae_pivot %>%
#   unite("Locus_site", c(Locus, Study_site), remove = F)
```

Kruskal Wallis 
```{r}
kruskal.test(Ae ~ Study_site, data = Ae_pivot)
pairwise.wilcox.test(Ae_pivot$Ae, Ae_pivot$Study_site, p.adjust.method = "bonferroni")
```
```{r}
modAe <- lme4::glmer(formula = Ae ~ Study_site + (1|Locus), 
            data = Ae_pivot, 
            family = "gaussian")
summary(modAe)
```

# Individual inbreeding coefficient F
Inbreeding refers to an excess of homozygosity in a given individual due to the mating of
genetically related parents.
Inbreeding can be associated to a loss of fitness leading to ”inbreeding depression”.

A lower HO than HE suggests inbreeding, which reduces heterozygosity

## All together
```{r}
# mean inbreeding for each individual
Fsamp_all <- inbreeding(squirrel_genind, res.type = "sample")
# Fsamp_all[[1]]
# class(Fsamp_all)
head(names(Fsamp_all))
head(Fsamp_all[[1]],20)
# Fsamp_all is a list of values sampled from the likelihood distribution of each individual; 
# compute means for all individuals
Fmean <- sapply(Fsamp_all, mean)
Fmean
```

```{r}
hist(Fmean, 
     col = "orange", 
     xlab = "mean value of F",
     main = "Distribution of mean F across individuals")
```


```{r}
## estimate inbreeding - return maximum likelihood estimates
Fest <- inbreeding(squirrel_genind, res.type = "estimate")
mostInbred <- which.max(Fest)
# plot(Fdens[[mostInbred]], ylab = "Density", xlab = "F",
#      main = paste("Probability density of F values\nfor", names(mostInbred)))
# abline(v = Fest[mostInbred], col = "red", lty = 2)
# legend("topright", legend = "MLE", col = "red", lty = 2)

## note that estimates and average samples are likely to be different.
plot(Fest, ylab = "F", col = "blue",
     main = "comparison of MLE and average sample estimates of F")
points(Fmean, pch = 2, col = "red")
arrows(x0 = 1:length(Fest), y0 = Fest, 
       y1 = Fmean, x1 = 1:length(Fest), length = 0.125)
legend("topleft", legend = c("estimate", "sample"), col = c("blue", "red"),
       pch = c(1, 2), title = "res.type")
```

Look into individuals with high inbreeding 
```{r}
which(Fmean>0.4)
```

## Study sites
```{r}
# mean inbreeding for each individual
Fsamp_Rahn <- inbreeding(Rahn_pop, res.type = "sample")
Fsamp_Lilie <- inbreeding(Lilie_pop, res.type = "sample")
Fsamp_Treptow <- inbreeding(Treptow_pop, res.type = "sample")
# Fsamp_Rahn[[1]]
# class(Fsamp_Rahn)
head(names(Fsamp_Rahn))
head(Fsamp_Rahn[[1]],20)
# Fsamp_Rahn is a list of values sampled from the likelihood distribution of each individual; 
# compute means for all individuals
Fmean_Rahn <- sapply(Fsamp_Rahn, mean) 
Fmean_Lilie <- sapply(Fsamp_Lilie, mean)
Fmean_Treptow <- sapply(Fsamp_Treptow, mean)
#Fmean_Rahn
```


Mean individual heteroyzygosities
```{r}
mean(Fmean_Rahn)
mean(Fmean_Lilie)
mean(Fmean_Treptow)
```

standard deviations
```{r}
sd(Fmean_Rahn)
sd(Fmean_Lilie)
sd(Fmean_Treptow)
```

```{r}
F_df <- bind_rows(
  tibble(ID = names(Fmean_Rahn), F = Fmean_Rahn, Site = "Rahnsdorf"),
  tibble(ID = names(Fmean_Lilie), F = Fmean_Lilie, Site = "Lilienthal"),
  tibble(ID = names(Fmean_Treptow), F = Fmean_Treptow, Site = "Treptower_Park")
) 

F_df <- F_df %>%
  mutate(Site = factor(Site, levels = c("Rahnsdorf", "Lilienthal", "Treptower_Park")))
levels(F_df$Site)
```

Plot
```{r}
hist(Fmean, 
     col = "orange", 
     xlab = "mean value of F",
     main = "Distribution of mean F across all individuals",
     ylim = c(0,30),
     xlim = c(0,1.0))

hist(Fmean_Rahn, 
     col = "orange", 
     xlab = "mean value of F",
     main = "Distribution of mean F across individuals in Rahnsdorf",
     ylim = c(0,30),
     xlim = c(0,1.0))

hist(Fmean_Lilie, 
     col = "orange", 
     xlab = "mean value of F",
     main = "Distribution of mean F across individuals in Lilienthal",
     ylim = c(0,30),
     xlim = c(0,1.0))

hist(Fmean_Treptow, 
     col = "orange", 
     xlab = "mean value of F",
     main = "Distribution of mean F across individuals in Treptower Park",
     ylim = c(0,30),
     xlim = c(0,1.0))
```

```{r}
ggplot(F_df, aes(x = F)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "black") 

ggplot(F_df, aes(x = F)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "black") +
  facet_wrap(~ Site)
```

```{r}
ggplot(F_df, aes(x = Site, y = F, fill = Site)) +
  geom_boxplot()
```
```{r}
ggplot(F_df, aes(x = ID, y = F)) +
  geom_bar(stat = 'identity', fill = "steelblue") + 
  #geom_histogram(binwidth = 0.05, fill = "steelblue", color = "black") +
  facet_wrap(~ Site, scales = "free_x")
```


# Inbreeding coefficient (FIS) per study site

Per population: 
FIS = Inbreeding coefficient, measures the deviation from the HWE within a subpopulation
F_IS > 0 = More inbreeding than expected, deficit of heterozgotes
F_IS < 0 = More heterozygotes than expected

Per individual: 
The inbreeding coefficient F is defined as the probability that at a given locus, two
identical alleles have been inherited from a common ancestor and is calculated after Garris et al. 2010.

## Study sites
### Per population
```{r}
# FIS per locus (rows = loci, cols = populations)
#fis_mat <- as.data.frame(stats_sites$Fis) 
fis_mat <- stats_sites$Fis %>% as.data.frame() %>% round(digits = 4)

# Look at it
head(fis_mat)
#head(fis_mat2)

# using your Ho and Hs data.frames
fis_from_H <- 1 - (hobs_site / hexp_site)
fis_from_H <- fis_from_H %>% round(digits = 4) 

# Vorsicht: Division durch 0 → NA
head(fis_from_H)
```

Comparison of the different ways to calculate FIS
```{r}
all.equal(target = fis_mat, current = fis_from_H, tolerance = 1e-6)  # evtl TRUE oder mit kleinen numerischen Unterschieden
```
Summary table 
```{r}
# Mean FIS per population
fis_mean <- colMeans(fis_mat, na.rm = TRUE)
fis_mean2 <- apply(fis_mat, 2, mean, na.rm = TRUE)

# SD per population (does not make sense)
fis_sd <- apply(fis_mat, 2, sd, na.rm = TRUE)
# Summary table
fis_summary <- data.frame(
  population = colnames(fis_mat),
  FIS_mean = round(fis_mean, 2),
  FIS_mean2 = round(fis_mean2, 2),
  FIS_sd = round(fis_sd, 2)
)
fis_summary
```

Bootstrap über Loci: Konfidenzintervalle für mittleres FIS (robuster)
```{r}
#set.seed(42)
nboot <- 1000
pop_names <- colnames(fis_mat)
boots <- matrix(NA, nrow = nboot, ncol = length(pop_names))
colnames(boots) <- pop_names
n_loci <- nrow(fis_mat)

for (i in 1:nboot) {
  # sample loci with replacement
  idx <- sample(1:n_loci, size = n_loci, replace = TRUE)
  boot_mean <- colMeans(fis_mat[idx, , drop = FALSE], na.rm = TRUE)
  boots[i, ] <- boot_mean
}

# compute 95% CI per population
ci_lower <- apply(boots, 2, function(x) quantile(x, 0.025, na.rm = TRUE))
ci_upper <- apply(boots, 2, function(x) quantile(x, 0.975, na.rm = TRUE))

fis_boot_summary <- data.frame(
  population = pop_names,
  FIS_mean = round(colMeans(boots, na.rm = TRUE), 2),
  CI_lower = round(ci_lower, 2),
  CI_upper = round(ci_upper, 2)
)
fis_boot_summary
```

Estimated FIS values indicated low to moderate heterozygote deficits across populations. In Rahnsdorf, mean FIS was 0.05 (95% CI: −0.03–0.13), with a confidence interval overlapping zero, suggesting no significant deviation from Hardy–Weinberg equilibrium. Treptower Park showed a higher mean FIS of 0.09 (95% CI: 0.02–0.18), and Lilienthal had the highest value (FIS = 0.11, 95% CI: 0.03–0.19); in both cases the confidence intervals did not include zero, indicating significant positive FIS and thus heterozygote deficits. Overall, FIS increased slightly along the gradient, which could reflect population substructure, reduced gene flow, or mild inbreeding in more isolated areas.

```{r}
fis_summary_plot <- fis_summary %>%
  mutate(
    population = recode(
      population,
      Rahnsdorf = "Woodland area",
      Lilienthal = "Cemetery",
      Treptower_Park = "Urban park"
    ),
    population = factor(
      population,
      levels = c("Woodland area", "Cemetery", "Urban park")
    )
  )

ggplot(fis_summary_plot, aes(x = population, y = FIS_mean, fill = population)) +
  geom_col(width = 0.65) +
  geom_text(
    aes(label = sprintf("%.02f", FIS_mean)),
    vjust = -0.6,
    size = 4
  ) +
  scale_y_continuous(
    limits = c(0, 0.15),
    expand = expansion(mult = c(0, 0.08))
  ) +
  scale_fill_manual(values = c("Urban park" = "#8B2F1D", 
                               "Cemetery" = "#F0A202", 
                               "Woodland area" = "#006D6F")) +
  labs(
    x = NULL,
    y = expression(F[IS]),
    title = "Low to moderate subpopulation inbreeding across sites"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    legend.position = "none"
  )



# ggsave(filename = here::here("plots","genetics", "FIS_20260204.png"),
#        # width = 1650*1.25,
#        # height = 1200,
#        # units = "px",
#        dpi = 300)

```


## test for significant differences
```{r}
FIS_pivot <- fis_mat %>%
  rownames_to_column(var = "Locus") %>%
  pivot_longer(cols = -Locus, names_to = "Study_site", values_to = "FIS")
```

Kruskal Wallis 
```{r}
kruskal.test(FIS ~ Study_site, data = FIS_pivot)
pairwise.wilcox.test(FIS_pivot$FIS, FIS_pivot$Study_site, p.adjust.method = "bonferroni")
```

# Combine inbreeding data and MLH data
```{r}
genetic_indices <- F_df %>%
  left_join(y = MLH_sq, by = "ID") %>%
  mutate(ID = str_replace_all(ID, pattern = "\\*", replacement = ""), 
         ID = as.numeric(ID))

gen_squirrels_135_IDs <- gen_squirrels_135_IDs %>%
  rownames_to_column(.data = , var = "ID") %>%
  mutate(ID = as.numeric(ID))


genetic_indices_ID <- genetic_indices %>% 
  left_join(y = gen_squirrels_135_IDs, by = "ID")
```

```{r}
table(genetic_indices_ID$Study_site, genetic_indices_ID$Site)
```
```{r}
genetic_indices_ID
```


## comparison between populations
### sMLH
Mean per group
```{r}
genetic_indices_ID %>%
  group_by(Study_site) %>%
  dplyr::summarise(mean_sMLH = mean(MLH_sq), 
                   sd_sMLH = sd(MLH_sq), 
                   mean_F = mean(F),
                   sd_F = sd(F))
```

how is it distributed
```{r}
qqnorm(genetic_indices_ID$MLH_sq)
qqline(genetic_indices_ID$MLH_sq)
```

Kruskal-Wallis-Test
```{r}
kruskal.test(MLH_sq ~ Study_site, data = genetic_indices_ID)
pairwise.wilcox.test(x = genetic_indices_ID$MLH_sq, g = genetic_indices_ID$Study_site, 
                     p.adjust = "bonferroni")
```
Post-Hoc-Test for comparison between the study sites
```{r}
#library(FSA)   # enthält dunnTest

dunnTest(MLH_sq ~ Study_site,
         data = genetic_indices_ID,
         method = "bonferroni")

#library(rstatix)

dunn_test(MLH_sq ~ Study_site,
         data = genetic_indices_ID, 
          p.adjust.method = "bonferroni")
```

We tested for differences in standardized multilocus heterozygosity (sMLH) among the three study sites using a Kruskal–Wallis test (χ² = 6.55, df = 2, p = 0.038).
Post-hoc Dunn tests with Benjamini–Hochberg correction revealed that squirrels from Treptower Park had significantly higher sMLH compared to Rahnsdorf (adj. p = 0.032), whereas comparisons involving Lilienthal were not significant (all adj. p > 0.05).

### inbreeding
how is it distributed
```{r}
qqnorm(genetic_indices_ID$F)
qqline(genetic_indices_ID$F)
```
```{r}
View(genetic_indices_ID)
```

```{r}
mean_F <- genetic_indices_ID %>%
  group_by(study_site_specific) %>%
  summarise(mean_F = mean(F),
            sd_F = sd(F))
```


Kruskal-Wallis-Test
```{r}
kruskal.test(F ~ Study_site, data = genetic_indices_ID)
pairwise.wilcox.test(x = genetic_indices_ID$F, g = genetic_indices_ID$Study_site, 
                     p.adjust = "bonferroni")
```

Post-Hoc-Test for comparison between the study sites
```{r}
#library(FSA) 

dunnTest(F ~ Study_site,
         data = genetic_indices_ID,
         method = "bonferroni")

#library(rstatix)

dunn_test(F ~ Study_site,
         data = genetic_indices_ID, 
          p.adjust.method = "bonferroni")
```

The inbreeding coefficient F did not differ significantly among populations (Rahnsdorf: 0.197, Lilienthal: 0.239, Treptower Park: 0.222; Kruskal-Wallis p = 0.215), indicating no evidence of elevated inbreeding in any population.

In contrast, standardized multilocus heterozygosity (sMLH) differed significantly between populations (Kruskal-Wallis p = 0.038), with a significant pairwise difference between Rahnsdorf and Treptower Park (Dunn test, Bonferroni-adjusted p = 0.032). Mean sMLH values were highest in Rahnsdorf (1.12), intermediate in Lilienthal (0.997), and lowest in Treptower Park (0.964).

These results suggest that while inbreeding levels are generally low and similar across populations, subtle differences in individual genetic diversity exist, with the urban population (Treptower Park) exhibiting slightly lower heterozygosity compared to the rural population (Rahnsdorf).

### Summary results
Ho/He: no significant population-level differences.

sMLH: significant differences at the individual level, Rahnsdorf highest.

F: no significant differences; individual heterozygosity differences not solely due to inbreeding.


Observed and expected heterozygosity (Ho and He) did not differ significantly among the three study sites (Ho: χ² = 1.06, p = 0.589; He: χ² = 0.54, p = 0.765), indicating similar genetic diversity at the population level. In contrast, standardized multilocus heterozygosity (sMLH) differed significantly between populations (Kruskal-Wallis χ² = 6.55, p = 0.038), with individuals from Rahnsdorf showing the highest average heterozygosity. Individual inbreeding coefficients (F) did not differ significantly among sites, suggesting that the observed differences in sMLH are not strongly driven by inbreeding.“


# Save data
```{r}
# # All together
# ## Genind
# saveRDS(squirrel_genind,
#         file = here::here("output", "data-proc", "genetic-proc", "individual_genetics",  "squirrel_genind_20251120.RDS"))
# ## data frame
# saveRDS(squirrel_genind_df,
#         file = here::here("output", "data-proc", "genetic-proc", "individual_genetics",  "squirrel_genind_df_20251120.RDS"))
# ## hierf
# saveRDS(squirrel_all_hierf, 
#         file = here::here("output", "data-proc", "genetic-proc", "individual_genetics",  "squirrel_hierf_20251120.RDS"))
# 
# # Per study site
# ## genind
# saveRDS(squirrel_genind_sites,
#         file = here::here("output", "data-proc", "genetic-proc", "individual_genetics",  "squirrel_genind_sites_20251120.RDS"))
# ## data frame
# saveRDS(squirrel_genind_sites_df,
#         file = here::here("output", "data-proc", "genetic-proc", "individual_genetics",  "squirrel_genind_sites_df_20251120.RDS"))
# ## hierf
# saveRDS(squirrel_sites_hierf, 
#         file = here::here("output", "data-proc", "genetic-proc", "individual_genetics",  "squirrel_sites_hierf_20251120.RDS"))
# # squirrel individual genetic indices
# saveRDS(genetic_indices_ID, 
#         file = here::here("output", "data-proc", "genetic-proc", "individual_genetics", 
#                           "squirrel_genetic_indices_20251127.RDS"))
```

***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>
