---
title: "Squirrel genetics and survival: Population differences" 
description: |
    This script calculates population genetic parameters, including pairwise FST
    and compares genetic differentiation among the three study sites.
author:
    - name: "Sinah Drenske"
      affiliation: "Leibniz Institute for Zoo and Wildlife Research (IZW); Technische Universität Berlin"
      affiliation_url: https://www.izw-berlin.de/en/
      orcid_id: 0000-0002-2247-6507
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
        highlight: kate      ## styling of code
        code_folding: false  ## if `true` you can expand and shrink code chunks
        toc: true            ## if `true` adds a table of content
        toc_depth: 2         ## level to be displayed in the table of content
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      dev = "ragg_png", fig.width = 9, fig.height = 6, dpi = 600, retina = 1)
Sys.setlocale("LC_TIME", "C")
```

* **Research question:** Do Eurasasian red squirrel populations inhabiting sites with different levels of urbanisation in Berlin differ in genetic diversity, population structure, and connectivity?
* **Study area:** Three study sites in Berlin, Germany, representing different levels of urbanisation: a woodland site in Rahnsdorf, Lilienthal Cemetery, and Treptower Park.
  * Woodland = Rahnsdorf
  * Cemetery = Lilienthal Cemetery
  * Urban park = Treptower Park
* **Data:** Microsatellite genotypes and study-site assignments of individual Eurasian red squirrels from the three study sites and genind object of the data.


# Setup

```{r packages}
## for non-CRAN packages please keep install instruction
# install.packages("devtools")
# devtools::install_github("EcoDynIZW/d6")
# For the package related: please follow the instructions here: https://github.com/timothyfrasier/related 

## libraries used in this script
d6::simple_load(c("adegenet", 
                  "hierfstat",
                  "tidyverse", 
                  "pegas",
                  "poppr",
                  "related",
                  "strataG",
                  "writexl"
))
```

```{r}
set.seed(20200604)
```


## Plot setup
Set theme for ggplots
```{r theme}
theme_set(
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 2),
    plot.margin = margin(6, 10, 4, 4),
    axis.ticks = element_line(colour = "black", linewidth = 1.5),
    axis.ticks.length = unit(0.2, "cm"),
    title = element_text(size = 20), # axis family = "Zilla Slab", family = "Roboto", 
    axis.text = element_text(size = rel(1.4)))) # axis ticks, family = "Roboto", 
```

Labeller for study sites
```{r}
# Remove the underscore in the name of the study site without changing the factor levels
Study_site_names <- as_labeller(c("Rahnsdorf" = "woodland area", "Treptower_Park" = "urban park", "Lilienthal" = "cemetery"))
```

# Data

```{r data}
# Genind
squirrel_genind_sites <- readRDS(file = here::here("output", "data-proc", "genetic-proc", "individual_genetics",  "squirrel_genind_sites_20251120.RDS"))
# Hierf
squirrel_sites_hierf <- readRDS(file = here::here("output", "data-proc", "genetic-proc", "individual_genetics",  "squirrel_sites_hierf_20251120.RDS"))
```


# Fixation index (FST)
FST = Fixation index. Genetic differentiation between subpopulations: measures how geentically different subpopulations are from the total population
F_ST < 0.05 No genetic differentiation (populations are genetically identical), high gene flow
F_ST 0.05-0.15 moderate differentiation
F_ST = 1 Complete differentiation, no gene flow, isolated populations
RST = analogous to FST, but accounts for the mutation process at microsatellite loci
```{r}
stats_sites <- basic.stats(squirrel_sites_hierf)
stats_sites
colMeans(stats_all$perloc, na.rm = TRUE)
# Ho = observed heterozygosity
# Hs = within population gene diversity (sometimes misleadingly called expected heterozygosity)
# perloc = basic statistics per loci
# overall = basic statistics averaged over loci

# Basic statistics
# Ho, Hs see above
# Ht = overall gene diversity, total heterozygosity
# Dst = amount of gene diversity (Ht - Hs)
# Htp = (maybe expected heterozygosity when sample sizes are even)
# Dstp = Proportional differentiation Dst/Ht
# Fst = Inbreeding between populations
# Fis = 1 - Ho/Hs = Inbreeding within a population
# Dest = Jost's D
```

# Pairwise FST/Gst/Josts D
```{r}
#?pairwise.neifst
pairwise.neifst(squirrel_sites_hierf)

genet.dist(squirrel_sites_hierf, method = "Nei87") # same as above
```

```{r}
pairwise.WCfst(dat = squirrel_sites_hierf,
               diploid = TRUE) %>% round(digits = 2)
pairwise.WCfst(dat = squirrel_genind_sites,
               diploid = TRUE) %>% round(digits = 2) # same as above
# > 0.05 = geringe, > 0.15 = mittlere, > 0.25 = starke Differenzierung

# genet.dist(squirrel_sites_hierf, method = "WC84") # same as above
```



```{r}
fst_boots <- boot.ppfst(dat = squirrel_genind_sites,
                        nboot = 5000,
                        quant = c(0.025,0.975),
                        diploid = TRUE)  
fst_boots
fst_boots3 <- fst_boots
fst_boots3$ll <- round(fst_boots3$ll, 2)
fst_boots3$ul <- round(fst_boots3$ul, 2)
fst_boots3

fst_boots2 <- boot.ppfst(dat = squirrel_sites_hierf,
                        nboot = 5000,
                        quant = c(0.025,0.975),
                        diploid = TRUE)  
fst_boots2
```

Pairwise genetic differentiation (Weir & Cockerham’s FST) ranged from 0.06 to 0.08 across study sites. Bootstrapped 95% confidence intervals did not include zero for any population pair (Rahnsdorf–Treptower Park: FST = 0.07, CI = 0.05–0.09; Rahnsdorf–Lilienthal: FST = 0.08, CI = 0.06–0.10; Treptower Park–Lilienthal: FST = 0.06, CI = 0.04–0.08), indicating significant differentiation among all populations.

# FIT
FIT = Total inbreeding coefficient, combines effects of FIS and FST
```{r}
#fis_mat <- as.data.frame(stats_sites$Fis) 
#fis_mat <- stats_sites$Fis %>% as.data.frame() %>% round(digits = 4)
fis_mean_pop <- colMeans(fis_mat, na.rm = TRUE)  # mean FIS per population
fis_mean_pop

# compute overall global FST (Weir & Cockerham)
wc_res <- hierfstat::wc(squirrel_sites_hierf)
global_fst <- wc_res$FST   # single scalar

# FIT per population via the identity
fit_pop <- 1 - (1 - fis_mean_pop) * (1 - global_fst)

# make a summary table
fit_summary <- data.frame(
  population = names(fis_mean_pop),
  FIS_mean = round(fis_mean_pop, 3),
  global_FST = round(global_fst, 3),
  FIT = round(fit_pop, 3)
)
fit_summary
```


# Hardy-Weinberg equilibrium
Expectations: Random mating, no selection, no migration, no inbreeding, no mutations 

If there are significant differences, it is a hint about inbreeding, selection, genetic drift etc. 

## All together
```{r}
squirrels_hwe <- squirrel_genind %>%
  hw.test(B = 1000) %>%
  as.data.frame() %>%
  rename(chi_sq = `chi^2`, 
         p_val = `Pr(chi^2 >)`) %>%
  mutate(across(.cols = where(is.numeric), .fns = round, 4))
```


```{r}
sum(squirrels_hwe$p_val < 0.05)
table(squirrels_hwe$p_val < 0.05)
```


## Study sites
```{r}
study_sites_hwtest <- lapply(study_sites_pop, FUN = function(pop) hw.test(pop, B = 1000))
```

```{r}
pop_names <- c("Rahnsdorf", "Treptower_Park", "Lilienthal")

study_sites_HWE <- study_sites_hwtest %>%
  set_names(pop_names) %>%
  imap_dfr(~ {
    as_tibble(.x, rownames = "Locus") %>%
      mutate(Pop = .y)
  }) %>%
  rename(chi_sq = `chi^2`, 
         p_val = `Pr(chi^2 >)`) %>%
  mutate(across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(across(where(is.character), as.factor))
```

```{r}
sum(study_sites_HWE$p_val < 0.05)
table(study_sites_HWE$p_val < 0.05, study_sites_HWE$Pop)
table(study_sites_HWE$p_val < 0.05)
```





***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>
