---
title: "Squirrel genetics and survival: Cluster analysis with Evanno" 
description: |
    This script evaluates STRUCTURE results using the Evanno method and
    summarises patterns of population structure among the three study sites.
author:
    - name: "Sinah Drenske"
      affiliation: "Leibniz Institute for Zoo and Wildlife Research (IZW); Technische Universität Berlin"
      affiliation_url: https://www.izw-berlin.de/en/
      orcid_id: 0000-0002-2247-6507
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
        highlight: kate      ## styling of code
        code_folding: false  ## if `true` you can expand and shrink code chunks
        toc: true            ## if `true` adds a table of content
        toc_depth: 2         ## level to be displayed in the table of content
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      dev = "ragg_png", fig.width = 9, fig.height = 6, dpi = 600, retina = 1)
Sys.setlocale("LC_TIME", "C")
```

* **Research question:** Do Eurasian red squirrel populations inhabiting sites with different levels of urbanisation in Berlin differ in genetic diversity, population structure, and connectivity?
* **Study area:** Three study sites in Berlin, Germany, representing different levels of urbanisation: a woodland site in Rahnsdorf, Lilienthal Cemetery, and Treptower Park.
  * Woodland = Rahnsdorf
  * Cemetery = Lilienthal Cemetery
  * Urban park = Treptower Park
* **Data:** Prepared data from script 6a including different subsets of individuals.


# Setup

```{r packages}
## for non-CRAN packages please keep install instruction
# install.packages("devtools")
# devtools::install_github("EcoDynIZW/d6")

## libraries used in this script
d6::simple_load(c("tidyverse", 
                  "gridExtra", 
                  "royfrancis/pophelper" 
                  ))
```


# Data

```{r data}
# Structure lists
structure_list_135ID <- readRDS(file = here::here("output", "data-proc", "genetic-proc", "cluster_analysis",
                                                  "structure_list_135ID_20250819.RData"))

structure_list_100ID <- readRDS(file = here::here("output", "data-proc", "genetic-proc", "cluster_analysis",
                                                  "structure_list_100ID_20260210.RData"))

# Gen squirrels
gen_squirrels_135_IDs <- readRDS(file = here::here("output", "data-proc", "genetic-proc", "cluster_analysis",
                                                   "gen_squirrels_135_IDs_20250719.RDS"))

gen_squirrels_100_IDs <- readRDS(file = here::here("output", "data-proc", "genetic-proc", "cluster_analysis",
                                                   "gen_squirrels_100_IDs_20260210.RDS"))

```

# Overview gen_squirrels
```{r}
table(gen_squirrels_135_IDs$Study_site)
table(gen_squirrels_135_IDs$study_site_specific)
table(gen_squirrels_135_IDs$Sex, gen_squirrels_135_IDs$Study_site)
length(unique(gen_squirrels_135_IDs$Squirrel_ID))
length(unique(gen_squirrels_135_IDs$Name))
table(gen_squirrels_135_IDs$Name, useNA = "always")
```
```{r}
structure_list_135ID$"10k_Burnin_100k_MCMC_run_30_f"
```

# Attribute table of runs
```{r}
#135 IDs
gen_tab_135_IDs <- tabulateQ(qlist = structure_list_135ID, 
                             writetable = F, 
                             exportpath = here::here("output", "data-proc", "genetic-proc", "cluster_analysis"))
gen_tab_135_IDs


#100 IDs
gen_tab_100_IDs <- tabulateQ(qlist = structure_list_100ID, 
                             writetable = TRUE, 
                             exportpath = here::here("output", "data-proc", "genetic-proc", "cluster_analysis"))
gen_tab_100_IDs
```

# Summarise data per run
Creates a summary table from a tabulated dataframe per number of cluster (k)
```{r}
#135 IDs
gen_summary_135_IDs <- summariseQ(data = gen_tab_135_IDs, 
                          writetable = F, 
                          exportpath = here::here("output", "data-proc", "genetic-proc", "cluster_analysis"))
gen_summary_135_IDs

#100 IDs
gen_summary_100_IDs <- summariseQ(data = gen_tab_100_IDs,
                          writetable = TRUE,
                          exportpath = here::here("output", "data-proc", "genetic-proc", "cluster_analysis"))
gen_summary_100_IDs

# elpdmean = mean estimated ln probability of data 
# elpdsd = stnadard deviation of estimated ln probability of data
# elpdmin = estimated ln probability of data minus standard deviation
# elpdmax = estimated ln probability of data plus standard deviation
```

# Evanno method (delta K)
```{r}
# 135 IDs
evanno_135_IDs <- evannoMethodStructure(data = gen_summary_135_IDs,
                                writetable = F, #T, 
                                exportpath = here::here("output", "data-proc", "genetic-proc", "cluster_analysis"), 
                                outputfilename = "evanno_135_IDs_20250819")
evanno_135_IDs

# 100 IDs
evanno_100_IDs <- evannoMethodStructure(data = gen_summary_100_IDs,
                                writetable = TRUE, 
                                exportpath = here::here("output", "data-proc", "genetic-proc", "cluster_analysis"), 
                                outputfilename = "evanno_100_IDs_20250819")
evanno_100_IDs

# elpdmean = mean estimated ln probability of data 
# elpdsd = stnadard deviation of estimated ln probability of data
# elpdmin = estimated ln probability of data minus standard deviation
# elpdmax = estimated ln probability of data plus standard deviation
# Ink1 = First derivate
# Ink1max = Max error of first derivate
# lnk1min = Min error of first derivate
# lnk2 = Second derivate
# lnk2max = Max error of second derivate
# lnk2min = Min error of second derivate
# deltaK = Third derivative
```


```{r}
# 135 IDs
evanno_plot_135_IDs <- evannoMethodStructure(data = gen_summary_135_IDs,
                                             exportplot = TRUE,#T,
                                             writetable = F,#T,
                                             returnplot = T,
                                             returndata = F,
                                             basesize = 20,
                                             pointsize = 8,
                                             pointcol = "grey30", #"#55847D",
                                             pointtype = 20, 
                                             linesize = 2,
                                             linecol = "grey30",
                                             ebwidth = 0.3,
                                             ebcol = "grey30", #"#084552",
                                             textcol = "grey30",
                                             xaxisbreaks = c(2,4,6,8,10),
                                             
                                             gridsize = 0.2,
                                             theme = "theme_bw", 
                                             dpi = 300,
                                             height = 25,
                                             width = 30,
                                             
                                             exportpath = here::here("output", "data-proc", "genetic-proc",
                                                                     "cluster_analysis"),
                                             outputfilename = "evanno_135_IDs_20260211") 

grid.arrange(evanno_plot_135_IDs)

# 100 IDs
evanno_plot_100_IDs <- evannoMethodStructure(data = gen_summary_100_IDs,
                                            exportplot = TRUE,#T,
                                            writetable = F,#T,
                                            returnplot = T,
                                            returndata = F,
                                            basesize = 20,
                                            pointsize = 8,
                                            pointcol = "grey30",
                                            pointtype = 20, 
                                            linesize = 2,
                                            linecol = "grey30",
                                            ebwidth = 0.3,
                                            ebcol = "grey30",
                                            textcol = "grey30",
                                            xaxisbreaks = c(2,4,6,8,10),
                                            
                                            gridsize = 0.2,
                                            theme = "theme_bw", 
                                            dpi = 300,
                                            height = 25,
                                            width = 30,
                                            
                                            exportpath = here::here("output", "data-proc", "genetic-proc",
                                                                    "cluster_analysis"),
                                            outputfilename = "evanno_100_IDs_20260211") 
grid.arrange(evanno_plot_100_IDs)
```

# Plot of best genetic cluster K = 2 (???)
```{r}
###### 136 IDs
# 10k_Burnin_100k_MCMC_run_51_f
k2_list_135_IDs <- structure_list_135ID[c("10k_Burnin_100k_MCMC_run_26_f",
                                          "10k_Burnin_100k_MCMC_run_27_f", "10k_Burnin_100k_MCMC_run_28_f",
                                          "10k_Burnin_100k_MCMC_run_29_f", "10k_Burnin_100k_MCMC_run_30_f",
                                          "10k_Burnin_100k_MCMC_run_31_f", "10k_Burnin_100k_MCMC_run_32_f",
                                          "10k_Burnin_100k_MCMC_run_33_f", "10k_Burnin_100k_MCMC_run_34_f",
                                          "10k_Burnin_100k_MCMC_run_35_f", "10k_Burnin_100k_MCMC_run_36_f",
                                          "10k_Burnin_100k_MCMC_run_37_f", "10k_Burnin_100k_MCMC_run_38_f",
                                          "10k_Burnin_100k_MCMC_run_39_f", "10k_Burnin_100k_MCMC_run_40_f",
                                          "10k_Burnin_100k_MCMC_run_41_f", "10k_Burnin_100k_MCMC_run_42_f",
                                          "10k_Burnin_100k_MCMC_run_43_f", "10k_Burnin_100k_MCMC_run_44_f",
                                          "10k_Burnin_100k_MCMC_run_45_f", "10k_Burnin_100k_MCMC_run_46_f",
                                          "10k_Burnin_100k_MCMC_run_47_f", "10k_Burnin_100k_MCMC_run_48_f",
                                          "10k_Burnin_100k_MCMC_run_49_f", "10k_Burnin_100k_MCMC_run_50_f")]
# align the clusters among the single runs of a specific K
k2_list_135_IDs_align <- alignK(k2_list_135_IDs)
# merge runs of the same cluster
k2_list_135_IDs_align_merge <- mergeQ(k2_list_135_IDs_align)
#test <- k2list_align_merge[[1]]
squirrel_IDs_k2_list_135_IDs <- as.numeric

###### 100 IDs
# 10k_Burnin_100k_MCMC_run_11_f
k2_list_100_IDs <- structure_list_100ID[c("10k_Burnin_100k_MCMC_run_26_f",
                                        "10k_Burnin_100k_MCMC_run_27_f", "10k_Burnin_100k_MCMC_run_28_f",
                                        "10k_Burnin_100k_MCMC_run_29_f", "10k_Burnin_100k_MCMC_run_30_f",
                                        "10k_Burnin_100k_MCMC_run_31_f", "10k_Burnin_100k_MCMC_run_32_f",
                                        "10k_Burnin_100k_MCMC_run_33_f", "10k_Burnin_100k_MCMC_run_34_f",
                                        "10k_Burnin_100k_MCMC_run_35_f", "10k_Burnin_100k_MCMC_run_36_f",
                                        "10k_Burnin_100k_MCMC_run_37_f", "10k_Burnin_100k_MCMC_run_38_f",
                                        "10k_Burnin_100k_MCMC_run_39_f", "10k_Burnin_100k_MCMC_run_40_f",
                                        "10k_Burnin_100k_MCMC_run_41_f", "10k_Burnin_100k_MCMC_run_42_f",
                                        "10k_Burnin_100k_MCMC_run_43_f", "10k_Burnin_100k_MCMC_run_44_f",
                                        "10k_Burnin_100k_MCMC_run_45_f", "10k_Burnin_100k_MCMC_run_46_f",
                                        "10k_Burnin_100k_MCMC_run_47_f", "10k_Burnin_100k_MCMC_run_48_f",
                                        "10k_Burnin_100k_MCMC_run_49_f", "10k_Burnin_100k_MCMC_run_50_f")]
k2_list_100_IDs_align <- alignK(k2_list_100_IDs)
k2_list_100_IDs_align_merge <- mergeQ(k2_list_100_IDs_align)
squirrel_IDs_k2_list_100_IDs <- as.numeric(rownames(k2_list_100_IDs_align_merge[[1]]))

```


## With group labels
### Prepare group labels

```{r}
squirrel_study_site3_135_IDs <- gen_squirrels_135_IDs %>% 
  dplyr::select(Study_site) %>% 
  mutate(Study_site = as.character(Study_site),
    Study_site = dplyr::recode(
      Study_site,
      "Rahnsdorf" = "Woodland area",
      "Treptower_Park" = "Urban park",
      "Lilienthal" = "Cemetery"
    )
  )

squirrel_study_site3_100_IDs <- gen_squirrels_100_IDs %>% 
  dplyr::select(Study_site) %>% 
  mutate(Study_site = as.character(Study_site),
    Study_site = dplyr::recode(
      Study_site,
      "Rahnsdorf" = "Woodland area",
      "Treptower_Park" = "Urban park",
      "Lilienthal" = "Cemetery"
    )
         )
# Just for 135 IDs
squirrel_area <- gen_squirrels_135_IDs %>% dplyr::select(area_type) %>% mutate(area_type = as.character(area_type))
squirrel_study_site4 <- gen_squirrels_135_IDs %>% dplyr::select(study_site_specific) %>% mutate(study_site_specific = as.character(study_site_specific))
squirrel_sex <- gen_squirrels_135_IDs %>% dplyr::select(Sex) %>% mutate(Sex = as.character(Sex))
squirrel_sex$Sex[is.na(squirrel_sex$Sex)] <- "unknown"
squirrel_season <- gen_squirrels_135_IDs %>% dplyr::select(Season) %>% mutate(Season = as.character(Season))
```

### Plots with one group
#### Comparison with study sites
```{r}
# 135 IDs
p_sites_135_IDs <- plotQ(k2_list_135_IDs_align_merge,
                         ### main
                          #imgoutput="join",
                          returnplot = T, 
                          exportplot = T, 
                          basesize = 5,
                          panelratio = c(7,1),
                          ### Plot panel
                          clustercol = c("#4D4D4D", "#6C91BF"), 
                          sortind = "Cluster1", 
                          ### Individual labels
                          showindlab = FALSE,
                          indlabsize = 15, # also for axis size
                          sharedindlab = F,
                          
                          ### Group label panel
                          grplab = squirrel_study_site3_135_IDs,
                          ordergrp = T,
                          
                          ### Group labels
                          grplabsize = 1, 
                          grplabalpha = 0,
                          
                          ### Label line
                          linesize = 1, 
                          linealpha = 0,
                          ### Label points
                          pointsize = 1, 
                          pointtype = 15,
                          pointalpha = 0,
                          ### Strip panel
                          showsp = FALSE, 
                          splab = "",
                          splabsize = 0,
                          ### Axis
                          showyaxis = F, 
                          showticks = F,
                          ticksize = 0.5,
                          ### Divider line
                          divsize = 2,
                          ### export settings
                          dpi = 800,
                          height = 4,
                          width = 20,
                         exportpath = here::here("plots", "genetics"),
                         outputfilename = "cluster_bar_135_IDs_20260211")
grid.arrange(p_sites_135_IDs$plot[[1]])
```


```{r}
test <- as.data.frame(sort(gen_squirrels_100_IDs$Study_site))
squirrel_study_site3_100_IDs$Study_site %in% gen_squirrels_100_IDs$Study_site
squirrel_study_site3_100_IDs$Study_site %in% test
```

```{r}
p_sites_100_IDs <- plotQ(k2_list_100_IDs_align_merge, 
                          ### main
                          #imgoutput="join",
                          returnplot = T, 
                          exportplot = T, 
                          basesize = 5,
                          panelratio = c(7,1),
                          ### Plot panel
                          clustercol = c("#4D4D4D", "#6C91BF"), 
                          sortind = "Cluster1", 
                          ### Individual labels
                          showindlab = FALSE,
                          indlabsize = 15, # also for axis size
                          sharedindlab = F,
                          
                          ### Group label panel
                          grplab = squirrel_study_site3_100_IDs,
                          ordergrp = T,
                          
                          ### Group labels
                          grplabsize = 1, 
                          grplabalpha = 0,
                          
                          ### Label line
                          linesize = 1, 
                          linealpha = 0,
                          ### Label points
                          pointsize = 1, 
                          pointtype = 15,
                          pointalpha = 0,
                          ### Strip panel
                          showsp = FALSE, 
                          splab = "",
                          splabsize = 0,
                          ### Axis
                          showyaxis = F, 
                          showticks = F,
                          ticksize = 0.5,
                          ### Divider line
                          divsize = 2,
                          ### export settings
                          dpi = 800,
                          height = 4,
                          width = 20,
                          exportpath = here::here("plots", "genetics"),
                          outputfilename = "cluster_bar_100_IDs_20260211")
#grid.arrange(p_sites_100_IDs$plot[[1]])

# Not working due to a bug 
#selgrp = "Study_site",
#subsetgrp = c("Rahnsdorf", "Lilienthal"), #c("Rahnsdorf"), # "Treptower_Park", "Lilienthal", 
#subset = c("Treptower_Park", "Lilienthal"),
```

There is a bug in the function so one cannot order individuals and groups. So if I order individuals by Cluster I cannot reorder the study sites. But with renaming the sites it is fine

```{r}
grid.arrange(p_sites_100_IDs$plot[[1]], p_sites_100_IDs2$plot[[1]])
```


comparison
```{r}
grid.arrange(p_sites_135_IDs$plot[[1]])

grid.arrange(p_sites_100_IDs$plot[[1]])

grid.arrange(p_sites_135_IDs$plot[[1]], p_sites_100_IDs$plot[[1]])
```


#### Other variables just for 135 IDs
```{r}
p_area <- plotQ(k2_list_135_IDs_align_merge,returnplot = T,exportplot = F,basesize = 11,
            grplab = squirrel_area, sortind = "Cluster1",
            subsetgrp = c("rural", "urban"), 
            ordergrp = TRUE,
            grplabsize = 4, linesize = 0.8, pointsize = 4)

p_site4 <- plotQ(k2_list_135_IDs_align_merge,returnplot = T,exportplot = F,basesize = 11,
            grplab = squirrel_study_site4, sortind = "Cluster1",
            subsetgrp = c("Rahnsdorf", "Lilienthal", "Treptow_south", "Treptow_north"), ordergrp = TRUE,
            grplabsize = 4, linesize = 0.8, pointsize = 4)

p_season <- plotQ(k2_list_135_IDs_align_merge,returnplot = T,exportplot = F,basesize = 11,
            grplab = squirrel_season, sortind = "Cluster1",
            subsetgrp = c("Spring", "Autumn"), ordergrp = TRUE,
            grplabsize = 4, linesize = 0.8, pointsize = 4)


p_sex <- plotQ(k2_list_135_IDs_align_merge,returnplot = T,exportplot = F,basesize = 11,
            grplab = squirrel_sex, sortind = "Cluster1",
            #subsetgrp = c("f", "m", "unknown"), 
            ordergrp = TRUE,
            grplabsize = 4, linesize = 0.8, pointsize = 4)


grid.arrange(p_sites_135_IDs$plot[[1]], p_area$plot[[1]], p_site4$plot[[1]], p_season$plot[[1]], p_sex$plot[[1]],
             ncol = 1, nrow = 5)


# Study site is the best representation of the clusters
```

### Plots with two groups
```{r}
# p9 <- plotQ(k6list_align_merge,returnplot = T,exportplot = F,basesize = 11,
#             grplab = squirrel_study_site_sex, sortind = "Cluster1", 
#             #subsetgrp = c("Rahnsdorf", "Lilienthal", "Treptower_Park"), 
#             ordergrp = TRUE,
#             grplabsize = 4, linesize = 0.8, pointsize = 4)
# 
# p10 <- plotQ(k6list_align_merge,returnplot = T,exportplot = F,basesize = 11,
#             grplab = squirrel_study_site_year, sortind = "Cluster1", 
#             #subsetgrp = c("Rahnsdorf", "Lilienthal", "Treptower_Park"), 
#             ordergrp = TRUE,
#             grplabsize = 4, linesize = 0.8, pointsize = 4)
# 
# #grid.arrange(p4$plot[[1]], p9$plot[[1]], ncol = 1, nrow = 2)
# grid.arrange(p4$plot[[1]], p9$plot[[1]], p10$plot[[1]], ncol = 1, nrow = 3)
```

***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>
