---
title: "Squirrel genetics and survival: Preparation of survival data" 
description: |
    This script prepares individual squirrel data, capture histories, and
    trapping effort data for Cormack–Jolly–Seber survival analyses.
author:
    - name: "Sinah Drenske"
      affiliation: "Leibniz Institute for Zoo and Wildlife Research (IZW); Technische Universität Berlin"
      affiliation_url: https://www.izw-berlin.de/en/
      orcid_id: 0000-0002-2247-6507
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
        highlight: kate      ## styling of code
        code_folding: false  ## if `true` you can expand and shrink code chunks
        toc: true            ## if `true` adds a table of content
        toc_depth: 2         ## level to be displayed in the table of content
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      dev = "ragg_png", fig.width = 9, fig.height = 6, dpi = 600, retina = 1)
Sys.setlocale("LC_TIME", "C")
```

* **Research question:** Does apparent survival of Eurasian red squirrels differ among sites with different levels of urbanisation in Berlin?
* **Study area:** Three study sites in Berlin, Germany, representing different levels of urbanisation: a woodland site in Rahnsdorf, Lilienthal Cemetery, and Treptower Park.
  * Woodland = Rahnsdorf
  * Cemetery = Lilienthal Cemetery
  * Urban park = Treptower Park
* **Data:** Prepared individual squirrel data, capture histories, and trapping effort data from the three study sites.


# Setup

```{r packages}
## libraries used in this script

library("skimr")
library("tidyverse")
```


# Data

```{r data}
# Squirrel data
squirrels <- readRDS(file = here::here("output", "data-proc", "cmr-proc", "squirrels_20251127.RDS"))

# Human activity
huac_means_specific <- readRDS(file = here::here("output", "data-proc", "cmr-proc", "human_activity_means_20251127.RDS"))

# Genetic indices
genetic_indices_ID <- readRDS(file = here::here("output", "data-proc", "genetic-proc", "individual_genetics", 
                                                "squirrel_genetic_indices_20251127.RDS"))
# Trapping effort
trapping_effort <- readRDS(file = here::here("output", "data-proc", "cmr-proc", "trapping_effort_20251122.RDS"))
```

```{r}
table(squirrels$age_class, squirrels$Squirrel_ID)

# more than 1 age class: 8041A2E, 80428EA, 8042AB3, 8047A08, 8049A4F, 804BCFF, 8061F66, 808F334, 817F48A
```

# Data preparation
```{r}
## remove entries were there is no Sex and no Squirrel_ID
squirrels <- squirrels %>%
  dplyr::filter(!is.na(Sex)) %>% 
  dplyr::filter(!is.na(Squirrel_ID)) %>%
  #dplyr::filter(age_class == "adult") %>%
  droplevels()

# calculate the number of total recaptures per ID
squirrels <- squirrels %>%
  group_by(Squirrel_ID) %>%
  mutate(n_recap = n()) %>%
  ungroup()
table(squirrels$n_recap, useNA = "always")
squirrels$detect <- 1

# Only adults
## remove entries were there is no Sex and no Squirrel_ID
squirrels_adu <- squirrels %>%
  dplyr::filter(!is.na(Sex)) %>% 
  dplyr::filter(!is.na(Squirrel_ID)) %>%
  dplyr::filter(age_class == "adult") %>%
  droplevels()

# calculate the number of total recaptures per ID
squirrels_adu <- squirrels_adu %>%
  group_by(Squirrel_ID) %>%
  mutate(n_recap = n()) %>%
  ungroup()
table(squirrels_adu$n_recap, useNA = "always")
squirrels_adu$detect <- 1
```

```{r}
table(squirrels_once$age_class)
```


```{r}
squirrels_once <- squirrels %>%
  group_by(Squirrel_ID) %>%
  #mutate(n_recaps_season = )
  filter(row_number()== 1) %>%
  droplevels() %>%
  ungroup()
table(squirrels_once$Squirrel_ID, useNA = "always")

squirrels_once_adu <- squirrels_adu %>%
  group_by(Squirrel_ID) %>%
  #mutate(n_recaps_season = )
  filter(row_number()== 1) %>%
  droplevels() %>%
  ungroup()
table(squirrels_once_adu$Squirrel_ID, useNA = "always")
```
add genetic indices
```{r}
squirrels_gen <- squirrels_once %>%
  left_join(y = genetic_indices_ID[, c(2,4,13)], by = "Squirrel_ID")
# without values: 80426FF, 808F586, 805F9EA, 815E575, 805BECE, 7F25184
# IDs with geentic samples but not included in survival because no sex was determined: 806214A (Noah), 8042D7C (Jona)

squirrels_gen_adu <- squirrels_once_adu %>%
  left_join(y = genetic_indices_ID[, c(2,4,13)], by = "Squirrel_ID")
```

Add mean values per study site for individuals without genetic indices


```{r}
squirrels_gen_fill <- squirrels_gen %>%
  group_by(Study_site) %>%
  mutate(
    MLH_sq = case_when(
      is.na(MLH_sq) ~ mean(MLH_sq, na.rm = TRUE),
      TRUE        ~ MLH_sq
    ),
    F = case_when(
      is.na(F)    ~ mean(F, na.rm = TRUE),
      TRUE        ~ F
    )
  ) %>%
  ungroup()

summary(squirrels_gen_fill$MLH_sq)
summary(squirrels_gen_fill$F)

# Check if there are any NAs left
colSums(is.na(squirrels_gen_fill[, c("MLH_sq", "F")]))

########################
# Only adults
squirrels_gen_fill_adu <- squirrels_gen_adu %>%
  group_by(Study_site) %>%
  mutate(
    MLH_sq = case_when(
      is.na(MLH_sq) ~ mean(MLH_sq, na.rm = TRUE),
      TRUE        ~ MLH_sq
    ),
    F = case_when(
      is.na(F)    ~ mean(F, na.rm = TRUE),
      TRUE        ~ F
    )
  ) %>%
  ungroup()

summary(squirrels_gen_fill_adu$MLH_sq)
summary(squirrels_gen_fill_adu$F)

# Check if there are any NAs left
colSums(is.na(squirrels_gen_fill_adu[, c("MLH_sq", "F")]))
```

add data about the human activity
```{r}
squirrels_complete <- squirrels_gen_fill %>%
  left_join(y = huac_means_specific[, c(2,6:9)], by = "study_site_specific")

squirrels_complete_adu <- squirrels_gen_fill_adu %>%
  left_join(y = huac_means_specific[, c(2,6:9)], by = "study_site_specific")
```

# Create capture histories
adapted from https://jamesepaterson.github.io/jamespatersonblog/07_creatingcapturehistories 
```{r}
#length(unique(squirrels$Squirrel_ID)) # 115 - table should have 115 rows

capt_hist <- squirrels %>%
  select(Squirrel_ID, Project_phase, detect) %>%
  # remove duplicates, which may occur when individuals are caught multiple times in an event
  # For example, your event may be a year and an individual may be caught multiple times in a year.
  distinct(Project_phase, Squirrel_ID, .keep_all = TRUE) %>%
  # spread out data. The fill = 0 adds rows for combinations of id and event where individuals were not observerd
  pivot_wider(names_from = Project_phase, values_from = detect, values_fill = 0) %>%
  #spread(event, detect, fill = 0) %>% 
  # For every individual....
  group_by(Squirrel_ID) %>%
  # Paste together 0's and 1's
  # Unite is similar to paste. Here we are pasting the strings together from the second column (first capture event)
  # to the last capture event ("tail(names(.),1)").
  # we don't want any characters separating 0's and 1's, so we use: sep = ""
  unite("ch", 2:tail(names(.),1), sep = "")

##################
# Only adults
capt_hist_adu <- squirrels_adu %>%
  select(Squirrel_ID, Project_phase, detect) %>%
  # remove duplicates, which may occur when individuals are caught multiple times in an event
  # For example, your event may be a year and an individual may be caught multiple times in a year.
  distinct(Project_phase, Squirrel_ID, .keep_all = TRUE) %>%
  # spread out data. The fill = 0 adds rows for combinations of id and event where individuals were not observerd
  pivot_wider(names_from = Project_phase, values_from = detect, values_fill = 0) %>%
  #spread(event, detect, fill = 0) %>% 
  # For every individual....
  group_by(Squirrel_ID) %>%
  # Paste together 0's and 1's
  # Unite is similar to paste. Here we are pasting the strings together from the second column (first capture event)
  # to the last capture event ("tail(names(.),1)").
  # we don't want any characters separating 0's and 1's, so we use: sep = ""
  unite("ch", 2:tail(names(.),1), sep = "")
```

```{r}
table(capt_hist$ch)
```
# Add individual information
```{r}
capt_hist <- capt_hist %>%
  left_join(squirrels_complete, by = "Squirrel_ID") %>%
  select(Squirrel_ID, Name, ch, Sex, n_recap, # Squirrel specific
         area_type, Study_site, study_site_specific,  # Study site
         F, MLH_sq, # genetics
         spec_mean_30, spec_mean_60, mean_30, mean_60) # human activity

####
# Only adults
capt_hist_adu <- capt_hist_adu %>%
  left_join(squirrels_complete_adu, by = "Squirrel_ID") %>%
  select(Squirrel_ID, Name, ch, Sex, n_recap, # Squirrel specific
         area_type, Study_site, study_site_specific,  # Study site
         F, MLH_sq, # genetics
         spec_mean_30, spec_mean_60, mean_30, mean_60) # human activity
```

Change study_site_specific, sex and area type to numeric
```{r}
capt_hist <- capt_hist %>%
  mutate(sex_num = case_when(Sex == "f" ~ 1,
                             Sex == "m" ~ 0),
         area_num = case_when(area_type == "urban" ~ 1,
                             area_type == "rural" ~ 0), 
         site_spec = case_when(study_site_specific == "Rahnsdorf" ~ 1,
                                         study_site_specific == "Lilienthal" ~ 2,
                                         study_site_specific == "Treptow_south" ~ 3,
                                         study_site_specific == "Treptow_north" ~ 4))
```

Rename columns because the names are too long for RMark
```{r}
capt_hist <- capt_hist %>%
  rename(spec_30 = spec_mean_30,
         spec_60 = spec_mean_60)

# Only adults
capt_hist_adu <- capt_hist_adu %>%
  rename(spec_30 = spec_mean_30,
         spec_60 = spec_mean_60)
```

Remove columns
```{r}
capt_hist <- 
  capt_hist %>%
  #select(!c(Sex, Name, n_recap, area_type, study_site_specific)) %>%
  mutate(across(where(is.numeric), round, 5))

capt_hist_adu <- 
  capt_hist_adu %>%
  #select(!c(Sex, Name, n_recap, area_type, study_site_specific)) %>%
  mutate(across(where(is.numeric), round, 5))
```

# Prepare time-varying covariates

## Season
```{r}
season_phase <- squirrels %>%
  group_by(Project_phase, Season) %>%
  summarise(n = n()) %>%
  mutate(n = NULL)
```


## Effort per phase per study site
```{r}
effort_site_phase <- trapping_effort %>%
  group_by(Project_phase, Study_site) %>%
  summarise(trap_hours = sum(hours_day)) %>%
  ungroup() %>%
  complete(Study_site, Project_phase, fill = list(trap_hours = 0))
```



# Save data
```{r}
# # Complete squirrel data
# saveRDS(object = squirrels_complete, file = here::here("output", "data-proc", "cmr-proc", "squirrels_complete_20251128.RDS"))
# 
# # Capture histories
# saveRDS(object = capt_hist, file = here::here("output", "data-proc", "cmr-proc", "capture_histories_20251128.RDS"))
# 
# ## Only adults
# saveRDS(object = capt_hist_adu, file = here::here("output", "data-proc", "cmr-proc", "capture_histories_adults_20251212.RDS"))
# # Nr recaptures
# # saveRDS(object = N_recaps_phase, file = here::here("output", "data-proc", "cmr-proc", "N_recaps_phase_20251128.RDS"))
# 
# # Season
# saveRDS(object = season_phase, file = here::here("output", "data-proc", "cmr-proc", "season_phase_20251128.RDS"))
# 
# # Effort
# saveRDS(object = effort_site_phase, file = here::here("output", "data-proc", "cmr-proc", "effort_site_phase_20251128.RDS"))
```


***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>
