---
title: "Squirrel genetics and survival: Survival analysis" 
description: |
    This script fits, compares, and evaluates Cormack–Jolly–Seber models to
    estimate apparent survival and recapture probabilities.
author:
    - name: "Sinah Drenske"
      affiliation: "Leibniz Institute for Zoo and Wildlife Research (IZW); Technische Universität Berlin"
      affiliation_url: https://www.izw-berlin.de/en/
      orcid_id: 0000-0002-2247-6507

    - name: "Alina Theresa Berger"
      affiliation: "Leibniz Institute for Zoo and Wildlife Research (IZW); Technische Universität Berlin"

    - name: "Ashlee Jean Mikkelsen"
      affiliation: "Leibniz Institute for Zoo and Wildlife Research (IZW)"
      affiliation_url: https://www.izw-berlin.de/en/
      orcid_id: 0000-0003-1620-7711
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
        highlight: kate      ## styling of code
        code_folding: false  ## if `true` you can expand and shrink code chunks
        toc: true            ## if `true` adds a table of content
        toc_depth: 2         ## level to be displayed in the table of content
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      dev = "ragg_png", fig.width = 9, fig.height = 6, dpi = 600, retina = 1)
Sys.setlocale("LC_TIME", "C")
```

* **Research question:** Does apparent survival of Eurasian red squirrels differ among sites with different levels of urbanisation in Berlin?
* **Study area:** Three study sites in Berlin, Germany, representing different levels of urbanisation: a woodland site in Rahnsdorf, Lilienthal Cemetery, and Treptower Park.
  * Woodland = Rahnsdorf
  * Cemetery = Lilienthal Cemetery
  * Urban park = Treptower Park
* **Data:** Prepared capture histories, individual covariates, and trapping effort data for Eurasian red squirrels from the three study sites.


# Setup

```{r packages}
## libraries used in this script
library("tidyverse")
library("RMark")
```

## Plot setup

Set theme for ggplots
```{r theme}
theme_set(
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 2),
    plot.margin = margin(6, 10, 4, 4),
    axis.ticks = element_line(colour = "black", linewidth = 1.5),
    axis.ticks.length = unit(0.2, "cm"),
    title = element_text(size = 20), # axis family = "Zilla Slab", family = "Roboto", 
    axis.text = element_text(size = rel(1.4)))) # axis ticks, family = "Roboto", 
```

# Data
```{r data}
# Capture histories
capt_hist <- readRDS(file = here::here("output", "data-proc", "cmr-proc", "capture_histories_20251128.RDS"))
## Only adults
capt_hist_adu <- readRDS(file = here::here("output", "data-proc", "cmr-proc", "capture_histories_adults_20251212.RDS"))

# Season
season_phase <- readRDS(file = here::here("output", "data-proc", "cmr-proc", "season_phase_20251128.RDS"))

# Effort
effort_site_phase <- readRDS(file = here::here("output", "data-proc", "cmr-proc", "effort_site_phase_20251128.RDS"))
```


# Dipper example
```{r}
# data("dipper")
# dipper.proc <- process.data(data = dipper,
#                             group = "sex",
#                             model = "CJS") # group variable has to be in the data
# 
# # Make design data (from processed data)
# dipper.ddl <- make.design.data(dipper.proc)
```

```{r}
# fit.dipper.cjs.models <- function(){
# 
#   # Apparent survival (Phi) formula
#   Phi.sex.time <- list(formula = ~sex*time)  # Just like in other linear models "*" includes main effects and an interaction
#   Phi.time <- list(formula = ~time) # differs between discrete times
#   Phi.sex <- list(formula = ~sex) # differs between males and females
#   Phi.dot <- list(formula = ~1) # constant survival
# 
#   # Detection probability (p) formula
#   p.sex <- list(formula = ~sex)  # differs between males and females
#   p.time <- list(formula = ~time)  # one discrete estimate of p per capture event
#   p.dot <- list(formula = ~1) # constant detection
# 
#   # Construct all combinations and put into one model table
#   cml <- create.model.list(model = "CJS") # makes all possibile combinations of those parameter formulas
#   results <- mark.wrapper(model.list = cml,
#                          data = dipper.proc,
#                          ddl = dipper.ddl,
#                          external = FALSE,
#                          accumulate = FALSE,
#                          hessian = TRUE)
#   return(results)
# }
```


```{r}
# # Run function
# dipper.cjs.models <- fit.dipper.cjs.models()
```

# Study site + Sex as group
Process data and set grouping variables and afterwards create the design data for the MARK model specification
```{r}
# Process data
sitesex_proc <- process.data(data = capt_hist, 
                              model = "CJS", 
                              begin.time = 1,
                              groups = c("Study_site", "Sex")
                              )
# Design data for MARK
sitesex_ddl <- make.design.data(sitesex_proc)
```

Look at the data
```{r}
sitesex_proc$data

sitesex_ddl$Phi
```



## First trial

```{r}
sitesex_models <- function() {
# Phi
## basics
Phi.dot <- list(formula = ~1)
Phi.sex <- list(formula = ~Sex)
Phi.site <- list(formula = ~Study_site)


# p
p.dot <- list(formula = ~1)
p.site <- list(formula = ~Study_site)


cml <- create.model.list(model = "CJS")
results <- mark.wrapper(model.list = cml, 
                        data = sitesex_proc, 
                        ddl = sitesex_ddl, 
                        adjust = TRUE, 
                        output = FALSE)
return(results)
}
```

```{r}
sitesex_results <- sitesex_models()
sitesex_results
sitesex_results$model.table
```

## Adding time-dependent variables
### Season 
```{r}
sitesex_ddl$Phi$Season <- "Autumn" # New variable set as 0
sitesex_ddl$p$Season <- "Autumn"

# #dipper.ddl$Phi$Flood <- "noflood" 
# for (i in 1:8) {
#   
#   # Season für Phi und p
#   season_value <- season_phase$Season[season_phase$Project_phase == i]
#   sitesex_ddl$Phi$Season[sitesex_ddl$Phi$time == i] <- season_value
#   sitesex_ddl$p$Season[sitesex_ddl$p$time == i] <- season_value
# }

season_lookup <- setNames(
  as.character(season_phase$Season),
  as.character(season_phase$Project_phase)
)

# Füge Season hinzu
sitesex_ddl$Phi$Season <- factor(
  season_lookup[as.character(sitesex_ddl$Phi$time)],
  levels = c("Autumn", "Spring")
)

sitesex_ddl$p$Season <- factor(
  season_lookup[as.character(sitesex_ddl$p$time)],
  levels = c("Autumn", "Spring")
)
```


Look at the data
```{r}
sitesex_ddl$Phi
sitesex_ddl$p
```

Why does the dataframe have 168 rows? 
The number of rows results from the combination of groups and cohorts for each time interval (column 'time'). Each time interval may introduce a new cohort, representing captures and recaptures individuals. For each cohort, there are 6 groups (Sex: f, m × Study site: Rahnsdorf, Lilienthal, Treptower Park). Therefore, each new time step with a new cohort adds 6 rows corresponding to the 6 groups of that cohort.

### Effort 
```{r}
effort_lookup <- setNames(
  effort_site_phase$trap_hours,
  paste(effort_site_phase$Study_site, effort_site_phase$Project_phase)
)

sitesex_ddl$p$trap_hours <- effort_lookup[
  paste(sitesex_ddl$p$Study_site, sitesex_ddl$p$time)
]
```

Check if it works
```{r}
# First look
head(sitesex_ddl$p[, c("group", "time", "Study_site", "Season", "trap_hours")], 20)

# Does the site has the right time? Should be zero as we were not there
subset(sitesex_ddl$p, Study_site == "Lilienthal" & time %in% c(1, 2))

# Are there any NAs?
table(is.na(sitesex_ddl$p$trap_hours))
```



## Model formulation
### p combination
```{r}
sitesex_models_p <- function() {
# Phi
## basics
Phi.dot <- list(formula = ~ 1)


# p
## basics
p.dot <- list(formula = ~ 1)
p.site <- list(formula = ~ Study_site)
p.sex <- list(formula = ~ Sex)
p.effort <- list(formula = ~ trap_hours)
p.season <- list(formula = ~ Season)

## additive effects
p.effort.season <- list(formula = ~ trap_hours + Season)
p.effort.site <- list(formula = ~ trap_hours + Study_site)
p.effort.sex <- list(formula = ~ trap_hours + Sex)
p.season.sex <- list(formula = ~ Season + Sex)

## interactions
p.effort.season.int <- list(formula = ~ trap_hours * Season)
p.effort.site.int <- list(formula = ~ trap_hours * Study_site)
p.sex.season.int <- list(formula = ~ Sex * Season)
p.sex.site.int <- list(formula = ~ Sex * Study_site)
p.sex.season.effort.int <- list(formula = ~ Season*(Sex + trap_hours))

# Model list
cml <- create.model.list(model = "CJS")
results <- mark.wrapper(model.list = cml, 
                        data = sitesex_proc, 
                        ddl = sitesex_ddl, 
                        adjust = TRUE, 
                        output = FALSE)
# adjust = TRUE since some known S parameters could be at boundaries
return(results)
}
```

```{r}
sitesex_results_p <- sitesex_models_p()
sitesex_results_p
```


```{r}
Mod_tab_sitesex_p <- sitesex_results_p$model.table
Mod_tab_sitesex_p
```

Best p combination is p(~Season + sex_num). We will use this combination for the survival models

### Phi combination
```{r}
sitesex_models <- function() {
# Phi
## basics
Phi.dot <- list(formula = ~ 1)
Phi.site <- list(formula = ~ Study_site)
#Phi.spec <- list(formula = ~ site_spec)
#Phi.area <- list(formula = ~ area_num)
Phi.season <- list(formula = ~ Season)

## individual covariates
Phi.sex <- list(formula = ~ Sex)
Phi.MLH <- list(formula = ~ MLH_sq)
Phi.F <- list(formula = ~ F)

## human activity
Phi.human30 <- list(formula = ~ mean_30)


### additive effects
#Phi.human.site <- list(formula = ~ mean_30 + Study_site)
#Phi.human.spec30 <- list(formula = ~ spec_30 + site_spec)
#Phi.site.season <- list(formula = ~ Study_site + Season)
Phi.sex.season <- list(formula = ~ Sex + Season)
#Phi.sex.site <- list(formula = ~ Sex + Study_site)
#Phi.sex.area <- list(formula = ~ Sex + area_num)
Phi.sex.season.site <- list(formula = ~ Sex + Season + Study_site)
Phi.MLH.sex <- list(formula = ~ MLH_sq + Sex)
Phi.MLH.site <- list(formula = ~ MLH_sq + Study_site)

### interactions
Phi.sex.season.int <- list(formula = ~ Sex * Season)
#Phi.sex.site.int <- list(formula = ~ Sex * Study_site)
#Phi.site.season.int <- list(formula = ~ Study_site * Season)
Phi.sex.season.site.int <- list(formula = ~ Study_site*(Sex + Season))
Phi.MLH.sex.int <- list(formula = ~ MLH_sq * Sex)
Phi.MLH.site.int <- list(formula = ~ MLH_sq * Study_site)
Phi.human30.int <- list(formula = ~ mean_30 * Study_site)
#Phi.human.spec30.int <- list(formula = ~ spec_30 * site_spec)

## Full model
Phi.full.F <- list(formula = ~ Sex + F + Study_site + mean_30 + Season)
Phi.full.MLH <- list(formula = ~ Sex + MLH_sq + Study_site + mean_30 + Season)
Phi.full.MLH.int <- list(formula = ~ Season*(Sex + MLH_sq + Study_site + mean_30))
Phi.full.F.int <- list(formula = ~ Season*(Sex + F + Study_site + mean_30))

# p
#p.dot <- list(formula = ~ 1)
p.season.sex <- list(formula = ~ Season + Sex)


# Model list
cml <- create.model.list(model = "CJS")
results <- mark.wrapper(model.list = cml, 
                        data = sitesex_proc, 
                        ddl = sitesex_ddl, 
                        adjust = TRUE, 
                        output = FALSE)
# adjust = TRUE since some known S parameters could be at boundaries
return(results)
}
```

```{r}
?mark
sitesex_results <- sitesex_models()
sitesex_results
```

#### Results

```{r}
Mod_tab_sitesex <- sitesex_results$model.table
best_models_sitesex <- Mod_tab_sitesex %>%
  filter(DeltaAICc < 7)
```

Compare results from sex as numeric and sex as factor
```{r}
#Mod_tab$AICc == Mod_tab_sitesex$AICc
```

small overview over the best models
```{r}
table(best_models_sitesex$Phi)
```



```{r}
summary(sitesex_results$Phi.F.p.season.sex) # NS
summary(sitesex_results$Phi.sex.p.season.sex) # NS males lower survival

```

```{r}
sitesex_ddl$Phi
```

##### Plot it
```{r}
model.top=mark(sitesex_proc,sitesex_ddl,
             model.parameters=list(Phi=list(formula= ~F), 
                                   p=list(formula = ~Season + Sex)))


dat_pred_top = data.frame(index = rep(c(1, 85), each = 90), # 2 categories, length 45. 1 = sex one starts, 67 = sx 2 starts. 85comes from the ddl (sitesex_ddl$Phi) and is the first row with males as sex
                      Sex = rep(c('f', 'm'), each = 90),
                      Season = rep(c('Autumn', 'Spring'), length.out = 45,2), # sex dummy, same length. Dummy database
                      F = rep(seq(from = min(capt_hist$F), # numeric var, from min to max, 2 times for each sex (length 45 *2)
                                      to = max(capt_hist$F), length.out = 90),2))
Phi.pred_top = covariate.predictions(model = model.top, data = dat_pred_top)  # use the beste model, with the dummy data frame, predict survival based on sex and stadtardized body mass (ZBmass)

# Phi.Mass.Eqn <- function(x){ # probably I can ignore this, maybe a test (intercept + beta estimates)
#   1.51+(0.582*x)
# }
#
# Dummy.Zmass <- Phi.pred_top$estimates$Zbmass # take list of the body mass from the survival model
# Phi.pred.logit <- Phi.Mass.Eqn(Dummy.Zmass) # test function (for line 247)
# Phi.Mass.Pred <- logit2prob(Phi.pred.logit) # from logit 2 probability scale
# Phi.Mass.Pred
# 
# Phi.pred_top$estimates$TestPhi <- Phi.Mass.Pred # add TestPhi to the dataframe
# Phi.pred_top$estimates$Abmass <- (Phi.pred_top$estimates$Zbmass*sd(dat$BMass))+ # transform weight back + sd + mean. ABmass = opposite from Z
#   mean(dat$BMass)
# 

Phi.F.sex.season.fig <- 
  ggplot(data = Phi.pred_top$estimates, aes(F, estimate))+ # estimate = survival, you get it from covariate.predictions
  geom_ribbon(aes(ymin=lcl,
                  ymax=ucl
                  ),
              alpha=0.1)+
  geom_line(lwd=1)+
  ylab("Estimated survival probability")+
  xlab("F")+
  scale_x_continuous(expand = c(0.03,0.01)
                     #breaks = seq(1000,2500,250)
                     )+
  scale_y_continuous(expand = c(0.01,0.01))#+
  # geom_rug(data=dat, aes(x=BMass), inherit.aes = FALSE,
  #          lwd=1.1,
  #          alpha = 1/2)+
  # scale_color_manual(values = c("#D34643","#2C0B58"))+
  # scale_fill_manual(values = c("#D34643","#2C0B58"))

Phi.F.sex.season.fig

# ggsave(plot = Phi.MassSex.fig,
#        "Figures/Phi_MassSex.png",
#        height = 4, width = 4.7, units = "in",
#        dpi=800)

```

## Modify p 
```{r}
View(sitesex_ddl$p)
```

We were not on the cemetery in Project Phae 1 + 2 and not in the forest (Rahnsdorf) in Project Phase 4
```{r}
# pLilie.1 <- as.numeric(row.names(sitesex_ddl$p[sitesex_ddl$p$time == 1 & 
#                                                  sitesex_ddl$p$Study_site == "Lilienthal",]))
pLilie.2 <- as.numeric(row.names(sitesex_ddl$p[sitesex_ddl$p$time == 2 & 
                                                 sitesex_ddl$p$Study_site == "Lilienthal",]))
pRahn.4 <- as.numeric(row.names(sitesex_ddl$p[sitesex_ddl$p$time == 4 & 
                                                 sitesex_ddl$p$Study_site == "Rahnsdorf",]))

# p.time.fixed = list(formula = ~time, 
#                     fixed = list(index = c(pLilie.2, pRahn.4), 
#                                  value = 0))
```

```{r}
#test <- mark(data = sitesex_proc, ddl = sitesex_ddl, model.parameters = list(p = p.time.fixed))
```



### p combination
```{r}
sitesex2_models_p <- function() {
  # Phi
  ## basics
  Phi.dot <- list(formula = ~ 1)
  
  
  # p
  ## basics
  p.dot <- list(formula = ~ 1, 
                fixed = list(index = c(pLilie.2, pRahn.4), 
                             value = 0))
  p.site <- list(formula = ~ Study_site, 
                 fixed = list(index = c(pLilie.2, pRahn.4), 
                              value = 0))
  p.sex <- list(formula = ~ Sex, 
                fixed = list(index = c(pLilie.2, pRahn.4), 
                             value = 0))
  p.effort <- list(formula = ~ trap_hours, 
                   fixed = list(index = c(pLilie.2, pRahn.4), 
                                value = 0))
  p.season <- list(formula = ~ Season, 
                   fixed = list(index = c(pLilie.2, pRahn.4), 
                                value = 0))
  
  ## additive effects
  p.effort.season <- list(formula = ~ trap_hours + Season, 
                          fixed = list(index = c(pLilie.2, pRahn.4), 
                                       value = 0))
  p.effort.site <- list(formula = ~ trap_hours + Study_site, 
                        fixed = list(index = c(pLilie.2, pRahn.4), 
                                     value = 0))
  p.effort.sex <- list(formula = ~ trap_hours + Sex, 
                       fixed = list(index = c(pLilie.2, pRahn.4), 
                                    value = 0))
  p.season.sex <- list(formula = ~ Season + Sex, 
                       fixed = list(index = c(pLilie.2, pRahn.4), 
                                    value = 0))
  
  ## interactions
  p.effort.season.int <- list(formula = ~ trap_hours * Season, 
                              fixed = list(index = c(pLilie.2, pRahn.4), 
                                           value = 0))
  p.effort.site.int <- list(formula = ~ trap_hours * Study_site, 
                            fixed = list(index = c(pLilie.2, pRahn.4), 
                                         value = 0))
  p.sex.season.int <- list(formula = ~ Sex * Season, 
                           fixed = list(index = c(pLilie.2, pRahn.4), 
                                        value = 0))
  p.sex.site.int <- list(formula = ~ Sex * Study_site, 
                         fixed = list(index = c(pLilie.2, pRahn.4), 
                                      value = 0))
  p.sex.season.effort.int <- list(formula = ~ Season*(Sex + trap_hours), 
                                  fixed = list(index = c(pLilie.2, pRahn.4), 
                                               value = 0))
  
# Model list
cml <- create.model.list(model = "CJS")
results <- mark.wrapper(model.list = cml, 
                        data = sitesex_proc, 
                        ddl = sitesex_ddl, 
                        adjust = TRUE, 
                        output = FALSE)
# adjust = TRUE since some known S parameters could be at boundaries
return(results)
}
```

```{r}
sitesex2_results_p <- sitesex2_models_p()
sitesex2_results_p
```

```{r}
Mod_tab_sitesex2_p <- sitesex2_results_p$model.table %>%
  mutate(across(where(is.numeric), ~ round(., 2)))
Mod_tab_sitesex2_p
```

Best p combination is p(~Season + sex_num). We will use this combination for the survival models

### Phi combination
```{r}
sitesex2_models <- function() {
  # Phi
  ## basics
  Phi.dot <- list(formula = ~ 1)
  Phi.site <- list(formula = ~ Study_site)
  Phi.season <- list(formula = ~ Season)
  
  ## individual covariates
  Phi.sex <- list(formula = ~ Sex)
  Phi.MLH <- list(formula = ~ MLH_sq)
  Phi.F <- list(formula = ~ F)
  
  ## human activity
  Phi.human30 <- list(formula = ~ mean_30)
  
  
  ### additive effects
  Phi.sex.season <- list(formula = ~ Sex + Season)
  Phi.sex.season.site <- list(formula = ~ Sex + Season + Study_site)
  Phi.MLH.sex <- list(formula = ~ MLH_sq + Sex)
  Phi.MLH.site <- list(formula = ~ MLH_sq + Study_site)
  #Phi.F.sex <- list(formula = ~ F + Sex)
  Phi.F.site <- list(formula = ~ F + Study_site)
  
  ### interactions
  Phi.sex.season.int <- list(formula = ~ Sex * Season)
  Phi.sex.season.site.int <- list(formula = ~ Study_site*(Sex + Season))
  Phi.MLH.sex.int <- list(formula = ~ MLH_sq * Sex)
  Phi.MLH.site.int <- list(formula = ~ MLH_sq * Study_site)
  Phi.F.sex.int <- list(formula = ~ F * Sex)
  Phi.F.site.int <- list(formula = ~ F * Study_site)
  Phi.human30.int <- list(formula = ~ mean_30 * Study_site)
  
  ## Full model
  Phi.full.F <- list(formula = ~ Sex + F + Study_site + mean_30 + Season)
  Phi.full.MLH <- list(formula = ~ Sex + MLH_sq + Study_site + mean_30 + Season)
  Phi.full.MLH.int <- list(formula = ~ Season*(Sex + MLH_sq + Study_site + mean_30))
  Phi.full.F.int <- list(formula = ~ Season*(Sex + F + Study_site + mean_30))
  
  # p
  # p.dot <- list(formula = ~ 1, 
  #                      fixed = list(index = c(pLilie.2, pRahn.4),
  #                                   value = 0))
  p.season.sex <- list(formula = ~ Season + Sex,
                       fixed = list(index = c(pLilie.2, pRahn.4),
                                    value = 0))

  # Model list
  cml <- create.model.list(model = "CJS")
  results <- mark.wrapper(model.list = cml, 
                          data = sitesex_proc, 
                          ddl = sitesex_ddl, 
                          adjust = TRUE, 
                          output = FALSE)
  # adjust = TRUE since some known S parameters could be at boundaries
  return(results)
}
```

```{r}
#?mark
sitesex2_results <- sitesex2_models()
sitesex2_results
```

#### Results

```{r}
Mod_tab_sitesex2 <- sitesex2_results$model.table %>%
  mutate(across(where(is.numeric), ~ round(., 2)))
best_models_sitesex2 <- Mod_tab_sitesex2 %>%
  filter(DeltaAICc < 2)
Mod_tab_sitesex2
best_models_sitesex2
```
Best model
```{r}
summary(sitesex2_results$Phi.F.p.season.sex) # NS
# lowest survival
plogis(0.4715957 + (2.1898015*0.08857000)) # 0.6605052
```

Other models with delta AICc < 2
```{r}
summary(sitesex2_results$Phi.dot.p.season.sex) # 
summary(sitesex2_results$Phi.sex.p.season.sex) # NS
summary(sitesex2_results$Phi.MLH.p.season.sex) # NS
summary(sitesex2_results$Phi.human30.p.season.sex) # NS
# Null model
plogis(0.9423344)
# males
plogis(1.1126103 + (-0.3033973))
# females
plogis(1.1126103)
# MLH
plogis(1.4393898	 + (-0.4911230*0.34955))
# human activity
# Rahn
plogis(1.1305326 + (-0.0740061	*0.26316))
# Lilie
plogis(1.1305326 + (-0.0740061	*1.14423))
# Treptow
plogis(1.1305326 + (-0.0740061	*4.24936))
```



## Model predictions
Create a dummy data set
```{r}
# Calculate different F from min - max
F_grid <- seq(
  from = min(capt_hist$F, na.rm = TRUE),
  to   = max(capt_hist$F, na.rm = TRUE),
  length.out = 50
)

MLH_grid <- seq(
  from = min(capt_hist$MLH_sq, na.rm = TRUE),
  to   = max(capt_hist$MLH_sq, na.rm = TRUE),
  length.out = 50
)

# create data set
data_pred <- expand.grid(Sex = c('f', 'm'), 
                         Season = c('Autumn', 'Spring'),
                         Study_site = c('Rahnsdorf', 'Lilienthal', 'Treptower_Park'),
                         F = F_grid)

data_pred2 <- expand.grid(Sex = c('f', 'm'), 
                         Season = c('Autumn', 'Spring'),
                         Study_site = c('Rahnsdorf', 'Lilienthal', 'Treptower_Park'),
                         #F = F_grid,
                         MLH_sq = MLH_grid)
# add index
data_pred <- data_pred %>%
  mutate(index = case_when(
    Sex == "f" ~ 1,
    Sex == "m" ~ 85
  ))

data_pred2 <- data_pred2 %>%
  mutate(index = case_when(
    Sex == "f" ~ 1,
    Sex == "m" ~ 85
  ))

# add values for human activity
data_pred <- data_pred %>%
  mutate(mean_30 = case_when(
    Study_site == "Rahnsdorf" ~ 0.26316,
    Study_site == "Lilienthal" ~ 1.14423,
    Study_site == "Treptower_Park" ~ 4.24936
  ))

data_pred2 <- data_pred2 %>%
  mutate(mean_30 = case_when(
    Study_site == "Rahnsdorf" ~ 0.26316,
    Study_site == "Lilienthal" ~ 1.14423,
    Study_site == "Treptower_Park" ~ 4.24936
  ))
```

Check data
```{r}
nrow(data_pred)          # 600
table(data_pred$Sex)     # 300 f / 300 m
table(data_pred$index)   # 300 × 1, 300 × 85
table(data_pred$Study_site) # each 200
table(data_pred$mean_30)
table(data_pred$F)
table(data_pred2$MLH)
```




save the models
```{r}
model_1_F <- sitesex2_results$Phi.F.p.season.sex
model_2_null <- sitesex2_results$Phi.dot.p.season.sex # 
model_3_sex <- sitesex2_results$Phi.sex.p.season.sex # 
model_4_MLH <- sitesex2_results$Phi.MLH.p.season.sex # 
model_5_human <- sitesex2_results$Phi.human30.p.season.sex # 
```

Predict covariates
```{r}
Phi.pred.1.F <- covariate.predictions(model = model_1_F, data = data_pred)
Phi.pred.2.null <- covariate.predictions(model = model_2_null, data = data_pred)
Phi.pred.3.sex <- covariate.predictions(model = model_3_sex, data = data_pred)
Phi.pred.4.MLH <- covariate.predictions(model = model_4_MLH, data = data_pred2)
Phi.pred.5.human <- covariate.predictions(model = model_5_human, data = data_pred)
```

```{r}
View(Phi.pred.1.F$estimates)
View(Phi.pred_site_top2$estimates)
View(Phi.pred.2.null$estimates)
View(Phi.pred.3.sex$estimates)
View(Phi.pred.4.MLH$estimates)
View(Phi.pred.5.human$estimates)
# View(Phi.pred_top2$estimates)
# View(Phi.pred_top3$estimates)
```

### Results
Calculate mean survival
```{r}
avg_1_F <- Phi.pred.1.F$estimates %>%
  #group_by(Season, Sex) %>%
  summarise(avg_F = mean(F),
            avg_surv = mean(estimate),
            min_surv = min(estimate),
            max_surv = max(estimate))
avg_1_F

avg_2_null <- Phi.pred.2.null$estimates %>%
  #group_by(Season, Sex) %>%
  summarise(avg_F = mean(F),
            avg_surv = mean(estimate),
            min_surv = min(estimate),
            max_surv = max(estimate))
avg_2_null

avg_3_sex <- Phi.pred.3.sex$estimates %>%
  group_by(Sex) %>%
  summarise(avg_F = mean(F),
            avg_surv = mean(estimate),
            min_surv = min(estimate),
            max_surv = max(estimate))
avg_3_sex

avg_4_MLH <- Phi.pred.4.MLH$estimates %>%
  #group_by(Season, Sex) %>%
  summarise(avg_MLH = mean(MLH_sq),
            avg_surv = mean(estimate),
            min_surv = min(estimate),
            max_surv = max(estimate))
avg_4_MLH

avg_5 <- Phi.pred.5.human$estimates %>%
  group_by(Study_site) %>%
  summarise(avg_F = mean(F),
            avg_surv = mean(estimate),
            min_surv = min(estimate),
            max_surv = max(estimate))
avg_5

avg_52 <- Phi.pred.5.human$estimates %>%
  group_by(mean_30) %>%
  summarise(avg_F = mean(F),
            avg_surv = mean(estimate),
            min_surv = min(estimate),
            max_surv = max(estimate))
avg_52

# sum_phi <- Phi.pred_site_top2$estimates %>%
#   group_by(Season, Sex) %>%
#   summarise(avg_F = mean(F), 
#             avg_surv = mean(estimate),
#             min_surv = min(estimate),
#             max_surv = max(estimate))
# sum_phi
```
### Plot
```{r}
survival_plot_F <- 
  ggplot(data = Phi.pred.1.F$estimates, aes(F, estimate)) + 
  # estimate = survival, you get it from covariate.predictions
  geom_ribbon(aes(ymin = lcl,
                  ymax = ucl
  ),
  alpha = 0.1) +
  geom_line(lwd = 1) +
  ylab("Est. survival probability") +
  xlab("F") +
  scale_x_continuous(
    limits = c(0.05,0.65),
    breaks = seq(0.0,0.6, by = 0.2), 
    expand = c(0.03,0.01)
  ) +
  scale_y_continuous(
    limits = c(0.5,1),
    breaks = seq(0.5,1, by = 0.25),
    expand = c(0.01,0.01)) +
  geom_rug(data = capt_hist, aes(x = F), 
           inherit.aes = FALSE,
           length = unit(0.05, "npc"),
           lwd = 1.1,
           alpha = 1/2) 
survival_plot_F

ggsave(plot = survival_plot_F,
       filename = here::here("plots", "survival", "survival_F_20260122.png"),
       height = 4, width = 4.7, units = "in",
       dpi=800)
```


```{r}
Phi.F.sex.season.fig2 <- 
  ggplot(data = Phi.pred_top2$estimates, aes(F, estimate)) + # estimate = survival, you get it from covariate.predictions
  geom_ribbon(aes(ymin = lcl,
                  ymax = ucl
                  ),
              alpha = 0.1)+
  geom_line(lwd=1)+
  ylab("Est. survival probability")+
  xlab("F")+
  scale_x_continuous(expand = c(0.03,0.01)
                     #breaks = seq(1000,2500,250)
                     )+
  scale_y_continuous(expand = c(0.01,0.01))#+
  # geom_rug(data=dat, aes(x=BMass), inherit.aes = FALSE,
  #          lwd=1.1,
  #          alpha = 1/2)+
  # scale_color_manual(values = c("#D34643","#2C0B58"))+
  # scale_fill_manual(values = c("#D34643","#2C0B58"))
#Phi.F.sex.season.fig
Phi.F.sex.season.fig2

# ggsave(plot = Phi.F.sex.season.fig2,
#        filename = here::here("plots", "survival", "survival_F.png"),
#        height = 4, width = 4.7, units = "in",
#        dpi=800)

##############################################################################
# With study site
Phi.F.sex.season.site.fig2 <- 
  ggplot(data = Phi.pred_site_top2$estimates, aes(F, estimate, group = Study_site, colour = Study_site)) + # estimate = survival, you get it from covariate.predictions
  geom_ribbon(aes(ymin = lcl,
                  ymax = ucl
                  ),
              alpha = 0.1)+
  geom_line(lwd=1)+
  ylab("Est. survival probability")+
  xlab("F")+
  scale_x_continuous(expand = c(0.03,0.01)
                     #breaks = seq(1000,2500,250)
                     )+
  scale_y_continuous(expand = c(0.01,0.01))+
  geom_rug(data=dat2_pred_site_top, aes(x=F), inherit.aes = FALSE,
           lwd=1.1,
           alpha = 1/2) +
  facet_wrap(~Study_site)
  # scale_color_manual(values = c("#D34643","#2C0B58"))+
  # scale_fill_manual(values = c("#D34643","#2C0B58"))
#Phi.F.sex.season.fig
Phi.F.sex.season.site.fig2

```



# Save data
```{r}
# # Study site + Sex as group, modified p
# saveRDS(object = Mod_tab_sitesex2_p, file = here::here("output", "data-proc", "cmr-proc", "squirrels_detprob_20251201.RDS"))
# vroom::vroom_write(x = Mod_tab_sitesex2_p, file = here::here("output", "data-proc", "cmr-proc", "squirrels_detprob_20251201.csv"), delim = "\t", append = FALSE, bom = TRUE) # , quote = "all"
# 
# saveRDS(object = Mod_tab_sitesex2, file = here::here("output", "data-proc", "cmr-proc", "squirrels_survival_20251201.RDS"))
# vroom::vroom_write(x = Mod_tab_sitesex2, file = here::here("output", "data-proc", "cmr-proc", "squirrels_survival_20251201.csv"), delim = "\t", append = FALSE, bom = TRUE)
```



***

<details><summary>Session Info</summary>

```{r sessionInfo}
## DO NOT REMOVE!
## We store the settings of your computer and the current versions of the
## packages used to allow for reproducibility
Sys.time()
#git2r::repository() ## uncomment if you are using GitHub
sessionInfo()
```

</details>
